---
title: "Temp_Data_All_Years"
output: html_notebook
---

```{r setup}
library(dplyr)
library(data.table)
library(here)
library(raster)
library(ncdf4)
library(lubridate)
load("/Users/zoekitchel/Documents/grad school/Rutgers/Coursework/Fall 2018/SDM/hauls_catch_Dec2017 1.RData")
```

First, I will make list of GPS points and days/year ranges for which I need temperatures

```{r}
latlongyear <- hauls %>%
  dplyr::select(lat, lon, year, month)
  

```

Okay, now getting SODA data
Where I got the data to match soda in trawlData (following Ryan's readme)
Soda Data Files

 1. Go Here: http://iridl.ldeo.columbia.edu/SOURCES/.CARTON-GIESE/.SODA/.v2p1p6/.temp/time/%28Jan%202000%29%28Dec%202008%29RANGEEDGES/lat/%2810.25N%29%2884.75N%29RANGEEDGES/lon/%28179.75W%29%2820.25W%29RANGEEDGES/dataselection.html

 2. Restrict ranges as follows for SST to match SODA in trawlData

 	-200E to 20E --> 440 pts

 	0N to 89.5N --> 179 pts

 	5.01 to 5.01

 	Jan 1958 to Dec 2008 --> 612 pts


 3. Stop Selecting

 4. Click Data Files

 5. Download NetCDF Format

```{r}
cp <- nc_open("/Users/zoekitchel/Documents/grad school/Rutgers/LabWork/Colonization_Extinction/CARTON-GIESE_SODA_v2p1p6_sstemp.nc")
print(cp)

# =========================================
# = Function to Read in SODA, Grab Surface =
# =========================================
get.soda.sst <- function(file){

	soda.info <- nc_open(file)
	name.soda.sizes <- sapply(soda.info$var$temp$dim, function(x)x$name)
	soda.sizes <- soda.info$var$temp$size
	dim.units <- sapply(soda.info$var$temp$dim, function(x)x$units)
	print(dim.units)
	stopifnot(grepl("months since ", dim.units[4])) # make sure time is in correct units and in right place
	names(soda.sizes) <- name.soda.sizes
	ntime <- soda.sizes["time"]
	ndepth <- soda.sizes["depth"]

	soda.time0 <- soda.info$var$temp$dim[[4]]$vals
	ref.date <- as.Date(gsub("months since ", "", dim.units[4]))
	start.before.ref <- grepl("-", soda.time0[1]) # is the first date before ref.date?
	n.month.before <- ceiling(abs(soda.time0[1])) + as.integer(start.before.ref)
	start.increment <- ifelse(start.before.ref, "-1 month", "1 month")
	time.start <- rev(seq.Date(ref.date, by=start.increment, length.out=n.month.before))[1]
	soda.time <- seq.Date(time.start, by="1 month", length.out=ntime)
	
	soda.sst <- brick(file)
	names(soda.sst) <- soda.time

		
	return(soda.sst)
	
}


## Using Ryan's code
soda_sst <- get.soda.sst("/Users/zoekitchel/Documents/grad school/Rutgers/LabWork/Colonization_Extinction/CARTON-GIESE_SODA_v2p1p6_sstemp.nc")
soda_sst

plot(soda_sst[[9]])

saveRDS(soda_sst, "SODA2.1.6_SST.rds")
```

New column in latlongyear with "X1958.01.01" format, looks like I may not need to do this after all...

```{r}
latlongyear$month_digits <- NA
for (i in 1:nrow(latlongyear)) {
  if (latlongyear$month[[i]] < 10) {
    latlongyear$month_digits[[i]] <- paste0("0", month(latlongyear$month[[i]]))
  } else {
    latlongyear$month_digits[[i]] <- paste0(month(latlongyear$month[[i]]))
  }
}
latlongyear$day <- paste0("0", c(1))
latlongyear$year_digits <- paste0("X", latlongyear$year)
latlongyear$ID <- paste(latlongyear$year_digits, latlongyear$month_digits, latlongyear$day, sep=".")
latlongyearIDonly <- latlongyear %>%
  dplyr::select(ID, lat, lon)
justlatlong <- latlongyear %>%
  dplyr::select(lat, lon)
```
Now, let's extract values
```{r}
rasValue <- extract(soda_sst, justlatlong)
```

