---
title: "Temp_Data_All_Years"
output: html_notebook
---
This will all be done in terminal R window. 

```{bash terminal setup}
ssh zoek@amphiprion.deenr.rutgers.edu
conda activate /local/home/zoek/enter/envs/myRenv3_5

```


```{r setup}
library(dplyr)
library(data.table)
library(raster)
library(ncdf4)
library(lubridate)
library(maptools)
library(rgdal)
library(tidyr)
library(ggplot2)

```

I have already converted NetCDF files to grd raster files. 
`col_ext/soda_bottom_sep2019.gri`
`col_ext/soda_bottom_sep2019.grd`
`col_ext/soda_surface_sep2019.gri`
`col_ext/soda_surface_sep2019.grd`

At this point, don't know difference between gri and grd. But, gri is 68.5 MB and .grd is 35.7 KB

```{r load netcdf now rasterbrick files}
soda_bottom <- brick("col_ext/soda_bottom_sep2019.gri")
soda_surface <- brick("col_ext/soda_surface_sep2019.gri")

```

Generating coordinates I need to extract, and then average
```{r generating coordinates to extract from hauls}
load("col_ext/hauls_catch_Dec2017 1.RData")
hauls <- data.table(hauls)
cols <- c("lat", "lon", "region")
latlongyear <- hauls[, cols, with = F]

# set the key to all columns
setkey(latlongyear)

# Get unique lat lon combos
latlongyear.u <- unique(latlongyear[list(lat, lon), nomatch = 0])
```

SET COORDINATE SYSTEM
```{r set coordinate system}
locations <- SpatialPoints(cbind(latlongyear.u$lon, latlongyear.u$lat), 
                            proj4string=CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
jpeg("plots.jpg")
plot(locations)
dev.off()
```

Now, let's extract values 

```{r extracting temperature values at correct GPS points}
system.time(temp_values_sbt <- raster::extract(soda_bottom, locations))
system.time(temp_values_sst <- raster::extract(soda_surface, locations))

#column bind
latlon_temp_values_sbt <- cbind(latlongyear.u, temp_values_sbt)
latlon_temp_values_sst <- cbind(latlongyear.u, temp_values_sst)
```

TROUBLE SHOOTING
#which rows are NA?

latlon_temp_values_sbt[is.na(X2002.07.01),]
#no data from 1980 ONWARDS for DFO_SoGulf

##############I'll fix this at some point, but for now, I will exclude them from avg

1: 47.80000 -65.58333 DFO_SoGulf
2: 47.91667 -65.61667 DFO_SoGulf
3: 47.91667 -65.51667 DFO_SoGulf
4: 47.93333 -65.63333 DFO_SoGulf
5: 47.95000 -65.80000 DFO_SoGulf
6: 47.98333 -65.88333 DFO_SoGulf
7: 48.00000 -65.60000 DFO_SoGulf

#no data before 1980 for DFO_SoGulf and a few AFSC_GOA sites

1: 45.92100  -63.50167 DFO_SoGulf         
2: 45.93683  -63.57550 DFO_SoGulf         
3: 45.95000  -63.53333 DFO_SoGulf          
4: 45.99750  -63.58583 DFO_SoGulf          
5: 55.69430 -133.45270   AFSC_GOA       
6: 55.80966 -133.42730   AFSC_GOA        
7: 57.53769 -156.01590   AFSC_GOA     

-----

I need a data.table in long format. This (sometimes) takes a while.

```{r melt temp data}
#surface
system.time(df_long_sst <- melt(latlon_temp_values_sst, measure.vars = 4:ncol(latlon_temp_values_sst),
               variable.name = "dateID", value.name = "temp")
)

#bottom
system.time(df_long_sbt <- melt(latlon_temp_values_sbt, measure.vars = 4:ncol(latlon_temp_values_sbt),
               variable.name = "dateID", value.name = "temp")
)

```

We now have to split the date, and get rid of X

```{r split dateID}
#surface
df_long_sst[, c("year", "month", "day") := tstrsplit(dateID, ".", fixed=TRUE)]

df_long_sst$year <- substr(as.character(df_long_sst$year), 2, 5)

#bottom
df_long_sbt[, c("year", "month", "day") := tstrsplit(dateID, ".", fixed=TRUE)]

df_long_sbt$year <- substr(as.character(df_long_sbt$year), 2, 5)

```

____

First is to average over all points in a region for every month/year combo. This takes a while.
```{r avg over GPS points}
#make sure region, year, and month are all factors
  #surface
  nms <- c("region", "year", "month")
  df_long_sst[, (nms) := lapply(.SD, factor), .SDcols = nms]
  
  #bottom
  df_long_sbt[, (nms) := lapply(.SD, factor), .SDcols = nms]

#average over year
  #surface
  soda_sst_monthavg <- df_long_sst[,list(sst_mean=mean(temp, na.rm=TRUE)),by=c("region", "year", "month")]
  #bottom
  soda_sbt_monthavg <- df_long_sbt[,list(sbt_mean=mean(temp,na.rm=TRUE)),by=c("region", "year", "month")]
  
  #good time to save

save(soda_sst_monthavg, soda_sbt_monthavg, file = "col_ext/soda_avg_monthly_temps.Rdata")

#for some reason, we're getting NA's for DFO_SoGulf and AFSC_GOA
```

It seems as though the best way to speak to the trawls data set and ask it to look for a specific year/month and then go back from that date by a specified interval is to turn all months into julian months, to eliminate year variable.
```{r convert month year to julian month}
#how many months?
year_month_key <- soda_sst_monthavg[, .N, by = c('year', 'month')][,1:2]
max_julian_month <- nrow(year_month_key)

#sequence of 1:756
julian_months <- seq(1, max_julian_month, by=1)

#makekey
year_month_key <- cbind(year_month_key, "month_julian" = julian_months)


```
I need to use this key to add julian month to data tables for sst and sbt. 
```{r add julian month}
#surface
soda_sst_monthavg <- soda_sst_monthavg[year_month_key, on = c(year = 'year', month = 'month')]

#surface
soda_sbt_monthavg <- soda_sbt_monthavg[year_month_key, on = c(year = 'year', month = 'month')]


```

Make sure that month and year match in characters and class in `hauls` and `soda_s?t_monthavg`

```{r match year month in hauls}
hauls$year <- as.factor(hauls$year)
hauls$month <- as.factor(hauls$month)
setattr(hauls$month,"levels",c('01','02','03','04','05','06','07','08','09','10','11','12'))

#now, add julian day to hauls data table as well

hauls <- hauls[year_month_key, nomatch = 0, on = c(year = 'year', month = 'month')]
```



Now, I have to start taking averages.
First, average over all GPS points in a region for a given year and month (this gives us avg temp per region per month)
Second, group by year, find mean in a year
Second, group by year, max in each year
Second, group by year, min in each year
Third, find difference between min and max for each year

This being said, I think I should think more carefully than this. I need to look a mean, min, max for a year BEFORE the month of the trawl.. this will be a little more complicated than how I've been doing it for sure. 

It also means I have to keep this mega file, and look on demand for temp data, which may not be possible. 

For each haul, look at haul month, and then take data from 12 months before that date. 

```{r averaging temperature values}
#hauls to have ready to go for this step
save(hauls, file="col_ext/hauls_10-01.Rdata")
#turn region into factor in hauls datatable
hauls$region <- as.factor(hauls$region)

#new columns for loop to populate
  #surface
    #raw
    temp.vars <- c("mean", "max", "min", "seas")
      hauls[,mean_sst_temp :=numeric(.N)]
      hauls[,max_sst_temp :=numeric(.N)]
      hauls[,min_sst_temp :=numeric(.N)]
      hauls[,seas_sst_temp :=numeric(.N)]
    
    
    #lag raw
      #new columns
      raw.lag.sst.col.names <- CJ(temp.vars, 1:10)[, paste(temp.vars, 1:10, sep ="_sst_temp_lag")]
      
      #loop to make new column names
      hauls[ ,raw.lag.sst.col.names[1:40]:=numeric(.N) ]
      #ended here
     
      
    #change
    hauls[, mean_sst_temp_change := numeric(.N)]
    hauls[, max_sst_temp_change := numeric(.N)]
    hauls[, min_sst_temp_change := numeric(.N)]
    hauls[, seas_sst_temp_change := numeric(.N)]

    
    #lag change
      #new columns
      change.lag.sst.col.names <- CJ(temp.vars, 1:9)[, paste(temp.vars, 1:9, sep ="_sst_temp_change_lag")]
      
      #loop to make new column names
      hauls[ ,change.lag.sst.col.names[1:36]:=numeric(.N) ]
      
    #abs
    hauls[, mean_sst_temp_change_abs := numeric(.N)]
    hauls[, max_sst_temp_change_abs := numeric(.N)]
    hauls[, min_sst_temp_change_abs := numeric(.N)]
    hauls[, seas_sst_temp_change_abs := numeric(.N)]
      
    #lag abs
      abs.change.lag.sst.col.names <- CJ(temp.vars, 1:9)[, paste(temp.vars, 1:9, sep ="_sst_temp_change_abs_lag")]
      
      #loop to make new column names
      hauls[ ,abs.change.lag.sst.col.names[1:36]:=numeric(.N) ]
  
  #bottom
    #raw
    temp.vars <- c("mean", "max", "min", "seas")
      hauls[,mean_sbt_temp :=numeric(.N)]
      hauls[,max_sbt_temp :=numeric(.N)]
      hauls[,min_sbt_temp :=numeric(.N)]
      hauls[,seas_sbt_temp :=numeric(.N)]
    
    
    #lag raw
      #new columns
      raw.lag.sbt.col.names <- CJ(temp.vars, 1:10)[, paste(temp.vars, 1:10, sep ="_sbt_temp_lag")]
      
      #loop to make new column names
      hauls[ ,raw.lag.sbt.col.names[1:40]:=numeric(.N) ]
      #ended here
     
      
    #change
    hauls[, mean_sbt_temp_change := numeric(.N)]
    hauls[, max_sbt_temp_change := numeric(.N)]
    hauls[, min_sbt_temp_change := numeric(.N)]
    hauls[, seas_sbt_temp_change := numeric(.N)]

    
    #lag change
      #new columns
      change.lag.sbt.col.names <- CJ(temp.vars, 1:9)[, paste(temp.vars, 1:9, sep ="_sbt_temp_change_lag")]
      
      #loop to make new column names
      hauls[ ,change.lag.sbt.col.names[1:36]:=numeric(.N) ]
      
    #abs
    hauls[, mean_sbt_temp_change_abs := numeric(.N)]
    hauls[, max_sbt_temp_change_abs := numeric(.N)]
    hauls[, min_sbt_temp_change_abs := numeric(.N)]
    hauls[, seas_sbt_temp_change_abs := numeric(.N)]
      
    #lag abs
      abs.change.lag.sbt.col.names <- CJ(temp.vars, 1:9)[, paste(temp.vars, 1:9, sep ="_sbt_temp_change_abs_lag")]
      
      #loop to make new column names
      hauls[ ,abs.change.lag.sbt.col.names[1:36]:=numeric(.N) ]
  
#now, loop to fill these new columns
for (i in 1:nrow(hauls)) {
    this_month_julian <- hauls$month_julian[i] #identify month/year
    this_region <- hauls$region[i] #identify region
    monthstosubtract <- seq(from = 11, by = 12, length.out = 11) 
    
    this_month_julian_reverse <- NA #month/year combo back in time
    
    for (j in 1:length(monthstosubtract)){
      this_month_julian_reverse[j] <- this_month_julian-monthstosubtract[j]
    }
    
  #surface raw
    subset_sst <- soda_sst_monthavg[region == this_region & month_julian >= this_month_julian_reverse[1] & soda_sst_monthavg$month_julian <= this_month_julian]
    
    hauls[(i),mean_sst_temp :=mean(subset_sst$sst_mean)] #mean temperature over the past 12 months
    hauls[(i),max_sst_temp :=max(subset_sst$sst_mean)] #max temp over the past 12 months
    hauls[(i),min_sst_temp :=min(subset_sst$sst_mean)] #min temp over the past 12 months
    hauls[(i),seas_sst_temp := max_sst_temp - min_sst_temp] #seasonality over the past 12 months
    
  #surface raw lags 
      for (j in 2:length(this_month_julian_reverse)) {
        
        subset_sst <- soda_sst_monthavg[region == this_region & month_julian >= this_month_julian_reverse[j] & month_julian < this_month_julian_reverse[j-1]] #subsets to correct year
        
        indx <- which(colnames(hauls) == paste0("mean_sst_temp_lag", j-1)) #finds column number with right name
        hauls[(i),(indx) := mean(subset_sst$sst_mean)] #assigns value to that cell
        indx <- which(colnames(hauls) == paste0("max_sst_temp_lag", j-1))
        hauls[(i),(indx) := max(subset_sst$sst_mean)]
        indx <- which(colnames(hauls) == paste0("min_sst_temp_lag", j-1))
        hauls[(i),(indx) := min(subset_sst$sst_mean)]
        indx <- which(colnames(hauls) == paste0("seas_sst_temp_lag", j-1))
        hauls[(i),(indx) := (get(paste0("max_sst_temp_lag", j-1))-get(paste0("min_sst_temp_lag", j-1)))]
      }
    
    #surface change
    hauls[(i),mean_sst_temp_change :=  mean_sst_temp-mean_sst_temp_lag1]
    hauls[(i),max_sst_temp_change :=  max_sst_temp-max_sst_temp_lag1]
    hauls[(i),min_sst_temp_change :=  min_sst_temp-min_sst_temp_lag1]
    hauls[(i),seas_sst_temp_change :=  seas_sst_temp-seas_sst_temp_lag1]
      
    #surface lag change
    for (j in 1:9) {
  #--mean
      indx_raw <- which(colnames(hauls) == paste0("mean_sst_temp_lag", j))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- hauls[(i), ..indx_raw]-hauls[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(hauls) == paste0("mean_sst_temp_change_lag", j)) #finds column number with right name to fill now
      hauls[(i),(indx_change) := temp_change] #assigns value to that cell
  #--max
      indx_raw <- which(colnames(hauls) == paste0("max_sst_temp_lag", j))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- hauls[(i), ..indx_raw]-hauls[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(hauls) == paste0("max_sst_temp_change_lag", j)) #finds column number with right name to fill now
      hauls[(i),(indx_change) := temp_change] #assigns value to that cell
  #--min
      indx_raw <- which(colnames(hauls) == paste0("min_sst_temp_lag", j))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- hauls[(i), ..indx_raw]-hauls[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(hauls) == paste0("min_sst_temp_change_lag", j)) #finds column number with right name to fill now
      hauls[(i),(indx_change) := temp_change] #assigns value to that cell
  #--seas
      indx_raw <- which(colnames(hauls) == paste0("seas_sst_temp_lag", j))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- hauls[(i), ..indx_raw]-hauls[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(hauls) == paste0("seas_sst_temp_change_lag", j)) #finds column number with right name to fill now
      hauls[(i),(indx_change) := temp_change] #assigns value to that cell
    }
    
    #absolute value of change in sst
    hauls[(i), mean_sst_temp_change_abs := abs(mean_sst_temp_change)]
    hauls[(i), max_sst_temp_change_abs := abs(max_sst_temp_change)]
    hauls[(i), min_sst_temp_change_abs := abs(min_sst_temp_change)]
    hauls[(i), seas_sst_temp_change_abs := abs(seas_sst_temp_change)]
  
    #absolute value of change in sst LAGS
    
    for (j in 1:9) {
  #--mean
      indx_raw <- which(colnames(hauls) == paste0("mean_sst_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(hauls[(i),..indx_raw])
      indx_abs <- which(colnames(hauls) == paste0("mean_sst_temp_change_abs_lag", j)) #find the right column to input new value
      hauls[(i), (indx_abs) := temp_change_abs]
  #--max
      indx_raw <- which(colnames(hauls) == paste0("max_sst_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(hauls[(i),..indx_raw])
      indx_abs <- which(colnames(hauls) == paste0("max_sst_temp_change_abs_lag", j)) #find the right column to input new value
      hauls[(i), (indx_abs) := temp_change_abs]
  #--min
      indx_raw <- which(colnames(hauls) == paste0("min_sst_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(hauls[(i),..indx_raw])
      indx_abs <- which(colnames(hauls) == paste0("min_sst_temp_change_abs_lag", j)) #find the right column to input new value
      hauls[(i), (indx_abs) := temp_change_abs]
  #--seas
      indx_raw <- which(colnames(hauls) == paste0("seas_sst_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(hauls[(i),..indx_raw])
      indx_abs <- which(colnames(hauls) == paste0("seas_sst_temp_change_abs_lag", j)) #find the right column to input new value
      hauls[(i), (indx_abs) := temp_change_abs]
    }
    
  #---------BOTTOM
      #bottom raw
    subset_sbt <- soda_sbt_monthavg[region == this_region & month_julian >= this_month_julian_reverse[1] & soda_sbt_monthavg$month_julian <= this_month_julian]
    
    hauls[(i),mean_sbt_temp :=mean(subset_sbt$sbt_mean)] #mean temperature over the past 12 months
    hauls[(i),max_sbt_temp :=max(subset_sbt$sbt_mean)] #max temp over the past 12 months
    hauls[(i),min_sbt_temp :=min(subset_sbt$sbt_mean)] #min temp over the past 12 months
    hauls[(i),seas_sbt_temp := max_sbt_temp - min_sbt_temp] #seasonality over the past 12 months
    
  #bottom raw lags 
      for (j in 2:length(this_month_julian_reverse)) {
        
        subset_sbt <- soda_sbt_monthavg[region == this_region & month_julian >= this_month_julian_reverse[j] & month_julian < this_month_julian_reverse[j-1]] #subsets to correct year
        
        indx <- which(colnames(hauls) == paste0("mean_sbt_temp_lag", j-1)) #finds column number with right name
        hauls[(i),(indx) := mean(subset_sbt$sbt_mean)] #assigns value to that cell
        indx <- which(colnames(hauls) == paste0("max_sbt_temp_lag", j-1))
        hauls[(i),(indx) := max(subset_sbt$sbt_mean)]
        indx <- which(colnames(hauls) == paste0("min_sbt_temp_lag", j-1))
        hauls[(i),(indx) := min(subset_sbt$sbt_mean)]
        indx <- which(colnames(hauls) == paste0("seas_sbt_temp_lag", j-1))
        hauls[(i),(indx) := (get(paste0("max_sbt_temp_lag", j-1))-get(paste0("min_sbt_temp_lag", j-1)))]
      }
    
    #bottom change
    hauls[(i),mean_sbt_temp_change :=  mean_sbt_temp-mean_sbt_temp_lag1]
    hauls[(i),max_sbt_temp_change :=  max_sbt_temp-max_sbt_temp_lag1]
    hauls[(i),min_sbt_temp_change :=  min_sbt_temp-min_sbt_temp_lag1]
    hauls[(i),seas_sbt_temp_change :=  seas_sbt_temp-seas_sbt_temp_lag1]
      
    #bottom lag change
    for (j in 1:9) {
  #--mean
      indx_raw <- which(colnames(hauls) == paste0("mean_sbt_temp_lag", j))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- hauls[(i), ..indx_raw]-hauls[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(hauls) == paste0("mean_sbt_temp_change_lag", j)) #finds column number with right name to fill now
      hauls[(i),(indx_change) := temp_change] #assigns value to that cell
  #--max
      indx_raw <- which(colnames(hauls) == paste0("max_sbt_temp_lag", j))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- hauls[(i), ..indx_raw]-hauls[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(hauls) == paste0("max_sbt_temp_change_lag", j)) #finds column number with right name to fill now
      hauls[(i),(indx_change) := temp_change] #assigns value to that cell
  #--min
      indx_raw <- which(colnames(hauls) == paste0("min_sbt_temp_lag", j))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- hauls[(i), ..indx_raw]-hauls[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(hauls) == paste0("min_sbt_temp_change_lag", j)) #finds column number with right name to fill now
      hauls[(i),(indx_change) := temp_change] #assigns value to that cell
  #--seas
      indx_raw <- which(colnames(hauls) == paste0("seas_sbt_temp_lag", j))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- hauls[(i), ..indx_raw]-hauls[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(hauls) == paste0("seas_sbt_temp_change_lag", j)) #finds column number with right name to fill now
      hauls[(i),(indx_change) := temp_change] #assigns value to that cell
    }
    
    #absolute value of change in sbt
    hauls[(i), mean_sbt_temp_change_abs := abs(mean_sbt_temp_change)]
    hauls[(i), max_sbt_temp_change_abs := abs(max_sbt_temp_change)]
    hauls[(i), min_sbt_temp_change_abs := abs(min_sbt_temp_change)]
    hauls[(i), seas_sbt_temp_change_abs := abs(seas_sbt_temp_change)]
  
    #absolute value of change in sbt LAGS
    
    for (j in 1:9) {
  #--mean
      indx_raw <- which(colnames(hauls) == paste0("mean_sbt_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(hauls[(i),..indx_raw])
      indx_abs <- which(colnames(hauls) == paste0("mean_sbt_temp_change_abs_lag", j)) #find the right column to input new value
      hauls[(i), (indx_abs) := temp_change_abs]
  #--max
      indx_raw <- which(colnames(hauls) == paste0("max_sbt_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(hauls[(i),..indx_raw])
      indx_abs <- which(colnames(hauls) == paste0("max_sbt_temp_change_abs_lag", j)) #find the right column to input new value
      hauls[(i), (indx_abs) := temp_change_abs]
  #--min
      indx_raw <- which(colnames(hauls) == paste0("min_sbt_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(hauls[(i),..indx_raw])
      indx_abs <- which(colnames(hauls) == paste0("min_sbt_temp_change_abs_lag", j)) #find the right column to input new value
      hauls[(i), (indx_abs) := temp_change_abs]
  #--seas
      indx_raw <- which(colnames(hauls) == paste0("seas_sbt_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(hauls[(i),..indx_raw])
      indx_abs <- which(colnames(hauls) == paste0("seas_sbt_temp_change_abs_lag", j)) #find the right column to input new value
      hauls[(i), (indx_abs) := temp_change_abs]
    }
    
    print(paste(i, nrow(hauls), sep = "/"))

}
      
#this takes 2 days to run
save(hauls, file = "hauls_fulltempdata.Rdata")
      
```

Another way to look at lags, but this isn't appropriate for the binomials I'm currently looking at

```{r}
ccf()
```