---
title: "Modeling individual instances of colonizations and extinctions usings traits"
output: html_notebook
---
Question: do species traits intereact with changes in temperature (current or past) to determine whether or not a given species colonizes a new region in a given year?

binomial regressions:

colonization in a given year? (yes/no) ~ mean temp of year of trawl x traits + max temp of year of trawl x traits + min temp of year of trawl x traits + change in mean temp from last year x traits + lags + traits

extinction in a given year? (yes/no) ~ mean temp of year of trawl x traits + max temp of year of trawl x traits + min temp of year of trawl x traits + change in mean temp from last year x traits + lags + traits

Traits to use?
  Ideal:
  generalism
  dispersal syndrome
  longevity
  body size
  generation time
  trophic level
  current range size
  mobility (fish, yes/no)

What we actually have from Tim: (** = ideal)
  trophic level:
    feeding.mode (generalism?)**
    tl (trophic level)**
  Reproduction: (probably whichever we have more data for)
    spawning.type (open_water/substratum, etc. )**
    egg.size**
  Growth:
    Linf (body size)
    Lmax (body size)*
    K (growth rate, 1/y)
    Lm (length at maturity)**
    Tmax (maximum age)
    tm (age maturity)**

*NB: I'm noticing that some species have multiple records in Tim's database, and that some stats are not quite matching up with what I find on fishbase.org (ie BSB max size and such)*

```{r setup}

```

For now, we only care about fish so let's go back to earlier code to extract fish only

```{r}

```

