---
title: "Modeling individual instances of colonizations and extinctions"
output: html_notebook
---
Question: do species traits intereact with changes in temperature (current or past) to determine whether or not a given species colonizes a new region in a given year?

binomial regressions:

colonization in a given year? (yes/no) ~ mean temp of year of trawl x traits + max temp of year of trawl x traits + min temp of year of trawl x traits + change in mean temp from last year x traits + lags + traits

extinction in a given year? (yes/no) ~ mean temp of year of trawl x traits + max temp of year of trawl x traits + min temp of year of trawl x traits + change in mean temp from last year x traits + lags + traits

Traits to use?
  Ideal:
  generalism
  dispersal syndrome
  longevity
  body size
  generation time
  trophic level
  current range size
  mobility (fish, yes/no)

What we actually have from Tim: (** = ideal)
  trophic level:
    feeding.mode (generalism?)**
    tl (trophic level)**
  Reproduction: (probably whichever we have more data for)
    spawning.type (open_water/substratum, etc. )**
    egg.size**
  Growth:
    Linf (body size)
    Lmax (body size)*
    K (growth rate, 1/y)
    Lm (length at maturity)**
    Tmax (maximum age)
    tm (age maturity)**

*NB: I'm noticing that some species have multiple records in Tim's database, and that some stats are not quite matching up with what I find on fishbase.org (ie BSB max size and such)*

```{r setup}
library(MuMIn)
library(lme4)
library(dplyr)
library(tidyr)
library(faraway)
library(data.table)
traits <- read.csv("traits_formanip.csv")
load("col_ext/hauls.Rdata")
load("col_ext/spp_master.RData")

```

Let's link species data from spp_master to temp data with lags and changes

```{r link species data to temp data}
#match region names between two data tables
reg <- levels(as.factor(spp_master$reg))
region <- c("AFSC_Aleutians","AFSC_EBS", "SEFSC_GOMex", "AFSC_GOA", "NEFSC_NEUS", "DFO_Newfoundland", "SCDNR_SEUS", "DFO_ScotianShelf", "AFSC_WCTri")
regions.dt <- data.table(reg, region)

hauls <- hauls[regions.dt, nomatch = 0, on = c("region")]

hauls[, reg := as.factor(reg)] #make region a factor in hauls

```
Just realizing now that I currently have data from 12 months before each trawl event. This isn't appropriate, as my colonization/extinction data is per year. FOR NOW (before I speak with Malin), I will just use earliest month in each year. 

```{r only keep earlier year temp data}
#keep one line of data for each year/region combo, and pick it based on smaller month value
hauls2 <- hauls[hauls[, .I[1], by = c("year", "region")]$V1]
cols <- c(5,38:286)
hauls2.r <- hauls2[, ..cols] #.. pulls in variable vs column name

```


```{r combine species data with temp data}
#make sure data type for reg and year in both are consistent
hauls2.r$year <- as.numeric(as.character(hauls2.r$year))
spp_master[, reg := as.factor(reg)]
#r combine species data with temp data
spp_master_ztemp <- spp_master[hauls2.r, nomatch = 0, on = c("reg", "year")]
head(spp_master_ztemp)
save(spp_master_ztemp, file = "spp_master_ztemp.Rdata")
```


First, let's ignore traits, and just look at temperature as predictors of yes no colonization and yes no extinction
Colonization in this year: col
Goes extinct in this year (here last year, gone now) : now_ext
Mean temp: mean_sbt_temp
Max temp: max_sbt_temp
Min temp: min_sbt_temp
Seasonality: seas_sbt_temp

```{r get to know temperature data better}
library(ggplot2)
#temperature through time
ggplot(data=spp_master_ztemp, aes(x=year, y=mean_sst_temp, col = reg)) +
  geom_point() +
  labs(x="Year", y="Mean Annual Surface Temperature (˚C)") +
  ylim(0,25) +
  geom_smooth(aes(col = reg), se = F, method = "lm") +
  theme_bw()

#temperature through time
ggplot(data=spp_master_ztemp, aes(x=year, y=mean_sbt_temp, col = reg)) +
  geom_point() +
  labs(x="Year", y="Mean Annual Bottom Temperature (˚C)") +
  ylim(0,25) +
  geom_smooth(aes(col = reg), se = F, method = "lm") +
  theme_bw()

```

What about correlations among temperature variables?
```{r r-correlations between temperature variables }
#surface
res <- cor(spp_master_ztemp[,45:48])
res
#bottom
res2 <- cor(spp_master_ztemp[,169:172])
res2

```
Surface
              mean_sst_temp max_sst_temp min_sst_temp seas_sst_temp
mean_sst_temp     1.0000000    0.9673832   0.96748430    0.18164518
max_sst_temp      0.9673832    1.0000000   0.87800968    0.41934324
min_sst_temp      0.9674843    0.8780097   1.00000000   -0.06633786
seas_sst_temp     0.1816452    0.4193432  -0.06633786    1.00000000

Mean is well correlated with max, and min both (0.96)
Max and min are medium correlated, but I'll still keep it in (0.87)

I will use max, min, seas

Bottom
              mean_sbt_temp max_sbt_temp min_sbt_temp seas_sbt_temp
mean_sbt_temp     1.0000000    0.9935599    0.9869266     0.8085135
max_sbt_temp      0.9935599    1.0000000    0.9655348     0.8669641
min_sbt_temp      0.9869266    0.9655348    1.0000000     0.7073712
seas_sbt_temp     0.8085135    0.8669641    0.7073712     1.0000000

Mean is well correlated with max (0.99) and min (0.98)
Max and min are well correlated with each other (0.96)
Seasonality is most correlated with max (0.86), and then mean (0.80) and then min (0.70)

I will use min, and seas (least correlated)

___
Colonization base mod surface
```{r base colonization model surface}
col_mod_base_sst <- glmer(col ~ max_sst_temp + min_sst_temp + seas_sst_temp + (1|year) + (1|reg), data = spp_master_ztemp, family = "binomial")
#rank deficient, so dropping a column
summary(col_mod_base_sst)
AICc(col_mod_base_sst)
```

Extinction base mod
```{r}
ext_mod_base_sst <- glmer(now_ext ~ max_sst_temp + min_sst_temp + seas_sst_temp + (1|reg) + (1|year), data = spp_master_ztemp, family = binomial)
summary(ext_mod_base_sst)
AICc(ext_mod_base_sst)
```
The above models perform really poorly, but this doesn't really surprise me. I think change in temperature is a bigger deal.
```{r}
colnames(spp_master_ztemp)
```

#START HERE
Now, let's play around with dredge.
```{r dredge colonization}
#I want to look at possible predictors
dredgesubset <- spp_master_ztemp[,c(1:3,6,45:168)] # this will have to change
#let's set the NA setting correctly
options(na.action = "na.fail")
#fitting global model
fm1_col <- glmer(col ~ . + (1|reg) + (1|year), family = binomial, data = dredgesubset)
#fixed-effect model matrix is rank deficient so dropping 102 columns / coefficients
#I have to trick model here so that it doesn't think it's rank deficient
vars <- colnames(spp_master_ztemp[,45:292])
form<-formula(paste0('col~',paste0(vars,collapse='+'),'+(1|reg)+(1|year)'))
attr(fm1_col@frame,"formula")<-form

#include only models with 1 term (and intercept)
dd <- dredge(fm1_col, m.lim = c(0, 1))

subset(dd, delta < 2)

# Visualize the model selection table:

par(mar = c(3,5,6,4))
plot(dd.col, labAsExpr = TRUE)

# Model average models with delta AICc < 2
model.avg(dd.col, subset = delta < 2)

#or as a 95% confidence set:
model.avg(dd.col, subset = cumsum(weight) <= .95) # get averaged coefficients

#'Best' model
summary(get.models(dd.col, 1)[[1]])
```
Now a dredge for extinction
```{r dredge extinction}
#fitting global model
fm1_ext <- glmm(now_ext ~  + (1|reg) + (1|year), family = binomial, data = dredgesubset)
dd.ext <- dredge(fm1_ext)

subset(dd.ext, delta < 2)

# Visualize the model selection table:

par(mar = c(3,5,6,4))
plot(dd.ext, labAsExpr = TRUE)

# Model average models with delta AICc < 2
model.avg(dd.ext, subset = delta < 2)

#or as a 95% confidence set:
model.avg(dd.ext, subset = cumsum(weight) <= .95) # get averaged coefficients

#'Best' model
summary(get.models(dd.ext, 1)[[1]])
```






*******
```{r surface colonization comparison}
#playing with year as factor
#can't get dredge to run above, so trying here with fewer variables
spp_master_ztemp[,year_factor := as.factor(year)]

#going to make loop to make all models I need to look at with single variables
surface_variables <- colnames(spp_master_ztemp[,45:168])
surface_colonization_model_comparison <- as.data.table(matrix(nrow = length(surface_variables)))
surface_colonization_model_comparison[, variable:=as.factor(V1)][, coef:=as.numeric(V1)][, p_value:=as.numeric(V1)][, AICc:=as.numeric(V1)]
surface_colonization_model_comparison[, V1 := NULL]

for (i in 1:length(surface_variables)){
  mod <- glmer(col ~ get(surface_variables[i]) + (1|reg) + (1|year_factor), family = binomial, data = spp_master_ztemp)
  surface_colonization_model_comparison[i,variable := surface_variables[i]]
  surface_colonization_model_comparison[i,coef := coef(summary(mod))[,"Estimate"][2]]
  surface_colonization_model_comparison[i,p_value := coef(summary(mod))[,"Pr(>|z|)"][2]]
  surface_colonization_model_comparison[i,AICc := AICc(mod)]
  
  print(paste(i, length(surface_variables), sep = "/"))
    
}

#21, 37 failed to converge
head(surface_colonization_model_comparison)

```

Here, looking at surface extinction
```{r surface extinction comparison}
#looping through all possible extinction variables with single variables
surface_variables <- colnames(spp_master_ztemp[,45:168]) #same as above

surface_extinction_model_comparison <- as.data.table(matrix(nrow = length(surface_variables)))
surface_extinction_model_comparison[, variable:=as.factor(V1)][, coef:=as.numeric(V1)][, p_value:=as.numeric(V1)][, AICc:=as.numeric(V1)]
surface_extinction_model_comparison[, V1 := NULL]

for (i in 1:length(surface_variables)){
  mod <- glmer(now_ext ~ get(surface_variables[i]) + (1|reg) + (1|year_factor), family = binomial, data = spp_master_ztemp)
  surface_extinction_model_comparison[i,variable := surface_variables[i]]
  surface_extinction_model_comparison[i,coef := coef(summary(mod))[,"Estimate"][2]]
  surface_extinction_model_comparison[i,p_value := coef(summary(mod))[,"Pr(>|z|)"][2]]
  surface_extinction_model_comparison[i,AICc := AICc(mod)]
  
  print(paste(i, length(surface_variables), sep = "/"))
    
}
```
Here, looking at bottom temp colonization
```{r bottom colonization comparison}
#looping through all possible extinction variables with single variables
bottom_variables <- colnames(spp_master_ztemp[,169:292]) #same as above

bottom_colonization_model_comparison <- as.data.table(matrix(nrow = length(bottom_variables)))
bottom_colonization_model_comparison[, variable:=as.factor(V1)][, coef:=as.numeric(V1)][, p_value:=as.numeric(V1)][, AICc:=as.numeric(V1)]
bottom_colonization_model_comparison[, V1 := NULL]

for (i in 1:length(bottom_variables)){
  mod <- glmer(col ~ get(bottom_variables[i]) + (1|reg) + (1|year_factor), family = binomial, data = spp_master_ztemp)
  bottom_colonization_model_comparison[i,variable := bottom_variables[i]]
  bottom_colonization_model_comparison[i,coef := coef(summary(mod))[,"Estimate"][2]]
  bottom_colonization_model_comparison[i,p_value := coef(summary(mod))[,"Pr(>|z|)"][2]]
  bottom_colonization_model_comparison[i,AICc := AICc(mod)]
  
  print(paste(i, length(bottom_variables), sep = "/"))
    
}
```
Finally, looking at bottom temp extinction
```{r bottom extinction comparison}
#looping through all possible extinction variables with single variables
bottom_variables <- colnames(spp_master_ztemp[,169:292]) #same as above

bottom_extinction_model_comparison <- as.data.table(matrix(nrow = length(bottom_variables)))
bottom_extinction_model_comparison[, variable:=as.factor(V1)][, coef:=as.numeric(V1)][, p_value:=as.numeric(V1)][, AICc:=as.numeric(V1)]
bottom_extinction_model_comparison[, V1 := NULL]

for (i in 1:length(bottom_variables)){
  mod <- glmer(now_ext ~ get(bottom_variables[i]) + (1|reg) + (1|year_factor), family = binomial, data = spp_master_ztemp)
  bottom_extinction_model_comparison[i,variable := bottom_variables[i]]
  bottom_extinction_model_comparison[i,coef := coef(summary(mod))[,"Estimate"][2]]
  bottom_extinction_model_comparison[i,p_value := coef(summary(mod))[,"Pr(>|z|)"][2]]
  bottom_extinction_model_comparison[i,AICc := AICc(mod)]
  
  print(paste(i, length(bottom_variables), sep = "/"))
    
}
```

Testing, as some models between colonization and extinction have the same coefficients


----------
Top models with single variables
```{r best surface colonization models}
head(surface_colonization_model_comparison[order(AICc)])

#mean highly correlated with max, so looking just at max, lag1 and lag8 are best
```
Merge surface and bottom for colonization
```{r merge surface and bottom for colonization}
colonization_model_comparisons <- rbind(bottom_colonization_model_comparison, surface_colonization_model_comparison)
#lowest AICc
head(colonization_model_comparisons[order(AICc)], 30)

```
Merge surface and bottom for extinction
```{r merge surface and bottom for extinction}
extinction_model_comparisons <- rbind(bottom_extinction_model_comparison, surface_extinction_model_comparison)
head(extinction_model_comparisons[order(AICc)], 30)

```
Seems that something is going on with seas_sbt_temp_change_abs_lag9
```{r seas_sbt_temp_change_abs_lag9}
plot(spp_master_ztemp$year, spp_master_ztemp$seas_sbt_temp_change_abs_lag9)
plot(spp_master_ztemp$year, spp_master_ztemp$seas_sst_temp_change_abs_lag9)
```

********
##Below I'm doing a more"organic" and less perscriptive dredge

What model performs better? a model with changes in the past year? or a model with changes in the year before last?

First, let's make a model with changes from the year before last for colonization
```{r}
col_mod_change_abs_lag1 <- glmer(col ~ sst_mean_change_lag1_abs + sst_max_change_lag1_abs + sst_min_change_lag1_abs + sst_seas_change_lag1_abs + (1|reg) + (1|year), data = spp_master_ztemp, family = binomial)
summary(col_mod_change_abs_lag1)
AICc(col_mod_change_abs_lag1)
```
Not much info from summary, but let's compare using AICc with other model
```{r}
AICc(col_mod_change_abs, col_mod_change_abs_lag1)
```

Poor results here! What about extinction?
```{r}
ext_mod_change_abs_lag1 <- glmer(now_ext ~ sst_mean_change_lag1_abs + sst_max_change_lag1_abs + sst_min_change_lag1_abs + sst_seas_change_lag1_abs + (1|reg) + (1|year), data = spp_master_ztemp, family = binomial)
summary(ext_mod_change_abs_lag1)
AICc(ext_mod_change_abs_lag1)
```
For extinction and lag 1 of changes, both absolute value of change in mean experienced last year and absolute value of change in max temp experienced last year appear significant
As absolute value of change in mean experienced last year increases, likelihood of extinction decreases
As absolute value of change in max temp experienced last year increases, likelihood of extinction increases

But what about comparing this to other model?
```{r}
AICc(ext_mod_change_abs, ext_mod_change_abs_lag1)
```
Here, unlike colonization, it appears that changes experienced LAST year are better than changes experienced this year at predicting extinction

Let's go back once more in time for each for kicks

Colonization based on changes in temp experienced between 2-3 years ago
```{r}
col_mod_change_abs_lag2 <- glmer(col ~ sst_mean_change_lag2_abs + sst_max_change_lag2_abs + sst_min_change_lag2_abs + sst_seas_change_lag2_abs + (1|reg) + (1|year), data = spp_master_ztemp, family = binomial)
summary(col_mod_change_abs_lag2)
AICc(col_mod_change_abs_lag2)
```
This is still bad, but what about one more lag
```{r}
col_mod_change_abs_lag3 <- glmer(col ~ sst_mean_change_lag3_abs + sst_max_change_lag3_abs + sst_min_change_lag3_abs + sst_seas_change_lag3_abs + (1|reg) + (1|year), data = spp_master_ztemp, family = binomial)
summary(col_mod_change_abs_lag3)
AICc(col_mod_change_abs_lag3)
```
And compare this using AICc to other models
```{r}
AICc(col_mod_change_abs, col_mod_change_abs_lag1, col_mod_change_abs_lag2, col_mod_change_abs_lag3)
```
Both incorporating lags in changes are equally poor for colonization, so we'll stick with most recent changes in temp variables. 

Next step, I will take away variables and see what's best for colonization using AIC. For dredge, I will check all combinations.
```{r}
summary(col_mod_change_abs)
```

```{r}
#remember, variables with best P values for colonization ARE seas, and then max, and then mean (min not significant)
#three variables, 4 models
col_mod_change_abs_3_var <- glmer(col ~ sst_mean_change_abs + sst_max_change_abs + sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_3_var2 <- glmer(col ~ sst_mean_change_abs + sst_min_change_abs + sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_3_var3 <- glmer(col ~ sst_mean_change_abs + sst_max_change_abs + sst_min_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_3_var4 <- glmer(col ~ sst_seas_change_abs + sst_max_change_abs + sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

AICc(col_mod_change_abs_3_var, col_mod_change_abs_3_var2, col_mod_change_abs_3_var3, col_mod_change_abs_3_var4)

#two variables, 6 models
col_mod_change_abs_2_var <- glmer(col ~ sst_max_change_abs + sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_2_var2 <- glmer(col ~ sst_max_change_abs + sst_min_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_2_var3 <- glmer(col ~ sst_max_change_abs + sst_mean_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_2_var4 <- glmer(col ~ sst_seas_change_abs + sst_min_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_2_var5 <- glmer(col ~ sst_seas_change_abs + sst_mean_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_2_var6 <- glmer(col ~ sst_min_change_abs + sst_mean_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

AICc(col_mod_change_abs_2_var, col_mod_change_abs_2_var2, col_mod_change_abs_2_var3, col_mod_change_abs_2_var4, col_mod_change_abs_2_var5, col_mod_change_abs_2_var6)

#one variable, 4 models
col_mod_change_abs_1_var <- glmer(col ~ sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_1_var2 <- glmer(col ~ sst_mean_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_1_var3 <- glmer(col ~ sst_min_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
col_mod_change_abs_1_var4 <- glmer(col ~ sst_max_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

AICc(col_mod_change_abs_1_var, col_mod_change_abs_1_var2, col_mod_change_abs_1_var3, col_mod_change_abs_1_var4)

AICc(col_mod_change_abs_1_var, col_mod_change_abs_2_var, col_mod_change_abs_3_var, col_mod_change_abs)
summary(col_mod_change_abs_3_var)
```
**The best model uses only seasonality as a predictor**
col_mod_change_abs_1_var <- glmer(col ~ sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

Extinction based on changes in temp experienced between 2-3 years ago
```{r}
ext_mod_change_abs_lag2 <- glmer(now_ext ~ sst_mean_change_lag2_abs + sst_max_change_lag2_abs + sst_min_change_lag2_abs + sst_seas_change_lag2_abs + (1|reg) + (1|year), data = spp_master_ztemp, family = binomial)
summary(ext_mod_change_abs_lag3)
AICc(ext_mod_change_abs, ext_mod_change_abs_lag1, ext_mod_change_abs_lag2, ext_mod_change_abs_lag3)
```
And compare this using AICc to other models
```{r}
AICc(ext_mod_change_abs, ext_mod_change_abs_lag1, ext_mod_change_abs_lag2)
```
Both lags improve model, so I'll go another level back in changes are equally poor, so we'll stick with most recent changes. 
```{r}
ext_mod_change_abs_lag3 <- glmer(now_ext ~ sst_mean_change_lag3_abs + sst_max_change_lag3_abs + sst_min_change_lag3_abs + sst_seas_change_lag3_abs + (1|reg) + (1|year), data = spp_master_ztemp, family = binomial)
summary(ext_mod_change_abs_lag3)
AICc(ext_mod_change_abs, ext_mod_change_abs_lag1, ext_mod_change_abs_lag2, ext_mod_change_abs_lag3)
```
The best is lag 2, so the change from 2-3 years ago is best at determining extinction probability

Next step, I will take away variables and see what's best for extinction using AIC. For dredge, I will check all combinations.

```{r}
summary(ext_mod_change_abs_lag2)
```

```{r}
#remember, variables with best P values for extinction with 2 lag are max change lag 2 abs, mean change lag 2 abs, seas change lag 2 abs, and last min change lag 2 abs

#for three variables (four models)
ext_mod_change_lag2_abs_3_var <- glmer(now_ext ~ sst_mean_change_lag2_abs + sst_max_change_lag2_abs + sst_seas_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_3_var2 <- glmer(now_ext ~ sst_mean_change_lag2_abs + sst_min_change_lag2_abs + sst_seas_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_3_var3 <- glmer(now_ext ~ sst_min_change_lag2_abs + sst_max_change_lag2_abs + sst_seas_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_3_var4 <- glmer(now_ext ~ sst_mean_change_lag2_abs + sst_max_change_lag2_abs + sst_min_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
AICc(ext_mod_change_lag2_abs_3_var2, ext_mod_change_lag2_abs_3_var3, ext_mod_change_lag2_abs_3_var4, ext_mod_change_lag2_abs_3_var)

#for two variables (six models)
ext_mod_change_lag2_abs_2_var <- glmer(now_ext ~ sst_mean_change_lag2_abs + sst_max_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_2_var2 <- glmer(now_ext ~ sst_mean_change_lag2_abs + sst_min_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_2_var3 <- glmer(now_ext ~ sst_mean_change_lag2_abs + sst_seas_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_2_var4 <- glmer(now_ext ~ sst_max_change_lag2_abs + sst_min_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_2_var5 <- glmer(now_ext ~ sst_max_change_lag2_abs + sst_seas_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_2_var6 <- glmer(now_ext ~ sst_min_change_lag2_abs + sst_seas_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

AICc(ext_mod_change_lag2_abs_2_var, ext_mod_change_lag2_abs_2_var2,ext_mod_change_lag2_abs_2_var3,ext_mod_change_lag2_abs_2_var4,ext_mod_change_lag2_abs_2_var5,ext_mod_change_lag2_abs_2_var6)

#for one variable (four models)
ext_mod_change_lag2_abs_1_var <- glmer(now_ext ~ sst_max_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_1_var2 <- glmer(now_ext ~ sst_mean_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_1_var3 <- glmer(now_ext ~ sst_min_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_1_var4 <- glmer(now_ext ~ sst_seas_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

AICc(ext_mod_change_lag2_abs_1_var,ext_mod_change_lag2_abs_1_var2,ext_mod_change_lag2_abs_1_var3,ext_mod_change_lag2_abs_1_var4)

#Now comparing best of each level of variables
AICc(ext_mod_change_lag2_abs_3_var, ext_mod_change_lag2_abs_1_var, ext_mod_change_lag2_abs_2_var, ext_mod_change_abs_lag2)
AICc()
summary(ext_mod_change_lag2_abs_3_var)
```
The best model is less clear here, ext_mod_change_lag2_abs_2_var has the lowest AICc, but, it's not substantially lower, and ext_mod_change_lag2_abs_1_var has fewer predictors, so let's go with that
**ext_mod_change_lag2_abs_1_var <- glmer(now_ext ~ sst_max_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)**: best model for extinction
```{r}
summary(ext_mod_change_lag2_abs_1_var)
```
We're able to explain 12% of variation in extinction from the change in maximum temperature between 2-3 years ago (extinction debt?)

**col_mod_change_abs_1_var <- glmer(col ~ sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)**: best model for colonization
```{r}
summary(col_mod_change_abs_1_var)
```
We're able to explain 13% of the variation in colonization with change in seasonality from last year

_______________

Can we plot what we have so far?

Colonization: col_mod_change_abs_1_var <- glmer(col ~ sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

x axis: absolute change in seasonality this past year
y axis: likelihood of colonization in this year

```{r}
#probability density function for colonization
x <- c(seq(0,1, by=0.001))
k <- sum(spp_master_ztemp$col)
n <- length(complete.cases(spp_master_ztemp$col))
bdist <- dbeta(x,k+1,n-k+1)

                                                
plot(x,bdist, ylab="Density Distribution", xlab="Proportion Of Year/Species Combos Undergoing Colonization", pch='.')

p <- c(0.025,0.975)
confint <- qbeta(p,k+1,n-k+1)
abline(v=confint,col="red",lty=2)

min(spp_master_ztemp$sst_seas_change_abs, na.rm = T)
max(spp_master_ztemp$sst_seas_change_abs, na.rm = T)

w <- seq(0,7,0.01)

plot(x=spp_master_ztemp$sst_seas_change_abs, y=spp_master_ztemp$col)
lines(x=w, y=ilogit(-3.33061+0.12645*w))
```
For each degree increase in temperature change of seasonality, log odds of extinction is predicted to change by 0.12645, or we can say that for a one unit increase in temp change, the odds of having colonization (versus not colonizing) increase by a factor of 1.13.


Extinction: ext_mod_change_lag2_abs_1_var <- glmer(now_ext ~ sst_max_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

x axis: absolute change in maximum temperature between 2 and 3 years ago
y axis: likelihood of extinction in this year

```{r}
#probability density function for colonization
x <- c(seq(0,1, by=0.001))
k <- sum(spp_master_ztemp$now_ext)
n <- length(complete.cases(spp_master_ztemp$now_ext))
bdist <- dbeta(x,k+1,n-k+1)

                                                
plot(x,bdist, ylab="Density Distribution", xlab="Proportion Of Year/Species Combos Undergoing Extinction", pch='.')

p <- c(0.025,0.975)
confint <- qbeta(p,k+1,n-k+1)
abline(v=confint,col="red",lty=2)

min(spp_master_ztemp$sst_max_change_abs, na.rm = T)
max(spp_master_ztemp$sst_max_change_abs, na.rm = T)

w <- seq(0,8,0.01)

plot(x=spp_master_ztemp$sst_max_change_abs, y=spp_master_ztemp$now_ext)
lines(x=w, y=ilogit(-4.25266+0.11950*w))
```
For each degree increase in temperature change of max, log odds of extinction is predicted to change by 0.11950, or we can say that for a one unit increase in temp change, the odds of having extinction (versus not becoming extinct) increase by a factor of 1.12.

Hard to interpret, these graphs are pretty lame, but big take aways are:
1. temperature change is better predictor than plain temperature
1. absolute value of temperature change is better predictor than just change
1. for colonization, temp change over past year is best predictor
1. for extinction, temp change experienced 2-3 years ago is best predictor
1. for colonization, of temp variables, change in seasonality is best predictor
1. for extinction, of temp variables, change in maximum temperature is best predictor
_______________

Now we will see if adding traits can improve the model, we only have traits for fish at this point, so we only care about fish so let's go back to earlier code to extract fish only

```{r}
load("fish_nofishkey.RData")
```

We will now limit spp_master_z to fish

```{r}
spp_master_ztemp_fishnofish <- left_join(spp_master_ztemp, fish_nofishkey, by = c("reg", "spp"))
spp_master_ztemp_fishnofish$fish <- as.factor(spp_master_ztemp_fishnofish$fish)
spp_master_ztemp_fishonly <- spp_master_ztemp_fishnofish %>%
  filter(fish == "Y")
?filter
summary(spp_master_ztemp_fishnofish$fish) # I have NA's but I will ignore for now
```
Let's now link to traits!
```{r}
traits <- read.csv("traits_formanip.csv")
traits$spp <- paste(traits$genus, traits$species, sep = " ")
#restrict to LME 1:9 and 54 according to https://www.st.nmfs.noaa.gov/ecosystems/lme/
levels(as.factor(traits$lme))
traits_NA <- traits %>%
  dplyr::filter(lme < 10 | lme == 54) %>%
  group_by(spp) %>%
  filter(row_number() == 1) #this means we only have one observation per species

spp_master_ztemp_fishonly_traits <- left_join(spp_master_ztemp_fishonly, traits_NA, by = "spp")
```

Traits that actually look reasonable to use

feeding.mode: benthivorous, generalist, herbivorous, piscivorous, planktivorous
  *maybe make this into generalist or specialist? (only 2 NA's) --> $generalist

tl: 2-4.65 (no NA's)

egg.size: 
  *(looks like I need to make this numeric or something) 111 NA's
  
spawning.type ( open_water-substratum, guarder, nonguarder, internal_bearer, external_bearer) 63 NA's

Lmax (2.5-1520) only 1 NA!

tmax (0.55 - 223) 181 NA's

tm (156 NA's) 

Lm (712 NA's) *exclude


Let's simplify feeding.mode to just generalist vs. specialist
```{r}
spp_master_ztemp_fishonly_traits$generalist <- NA
spp_master_ztemp_fishonly_traits$generalist[spp_master_ztemp_fishonly_traits$feeding.mode == "generalists"] <- 1
spp_master_ztemp_fishonly_traits$generalist[spp_master_ztemp_fishonly_traits$feeding.mode != "generalists"] <- 0
```
Let's look at egg size
```{r}
spp_master_ztemp_fishonly_traits$egg.size <- as.numeric(spp_master_ztemp_fishonly_traits$egg.size)
summary(spp_master_ztemp_fishonly_traits$egg.size)
#much better
```
For this analysis, I'm just going to look at trait values we have a lot of
generalist
tl
spawning.type
Lmax

Let's first run best models now that we're looking only at fish
reminders of best models:
col_mod_change_abs_1_var <- glmer(col ~ sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)
ext_mod_change_lag2_abs_1_var <- glmer(now_ext ~ sst_max_change_lag2_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

First colonization
```{r}
col_mod_change_abs_1_var_fishonly <- glmer(col ~ sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp_fishonly_traits)
summary(col_mod_change_abs_1_var_fishonly)
AICc(col_mod_change_abs_1_var_fishonly, col_mod_change_abs_1_var)
#unclear, but the AICc value for the fish only model is much better even though it appears dismal interms of p value, let's try with all variables again
col_mod_change_abs_fishonly <- glmer(col ~ sst_mean_change_abs + sst_max_change_abs + sst_min_change_abs + sst_seas_change_abs + (1|reg) + (1|year), data = spp_master_ztemp_fishonly_traits, family = binomial)
summary(col_mod_change_abs_fishonly) #nothing is significant
#still we'll try to add in traits
col_mod_change_abs_1_var_fishonly_wtraits <- glmer(col ~ sst_seas_change_abs*generalist + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp_fishonly_traits)
summary(col_mod_change_abs_1_var_fishonly_wtraits)
summary(col_mod_change_abs_1_var_fishonly)
AICc(col_mod_change_abs_1_var_fishonly_wtraits, col_mod_change_abs_1_var_fishonly)
```
Let's try the same with extinction
```{r}
ext_mod_change_lag2_abs_1_var_fishonly_wtraits <- glmer(now_ext ~ sst_max_change_lag2_abs*generalist + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp_fishonly_traits)
summary(ext_mod_change_lag2_abs_1_var_fishonly_wtraits)
```


One idea would just be to seperate by mobility as Malin did in Nature 2019
Reminder, best Model Colonization:
col_mod_change_abs_1_var <- glmer(col ~ sst_seas_change_abs + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp)

```{r}
#import Malin's dataset
mobility_traits <- read.csv("dataset_1_traits_nature.csv")
spp_master_ztemp_fishnofish_mobility <- inner_join(spp_master_ztemp_fishnofish, mobility_traits, by = "spp")
col_mod_change_abs_1_var_mobility <- glmer(col ~ sst_seas_change_abs*fish + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp_fishnofish)
summary(col_mod_change_abs_1_var_sessilemobile)
AICc(col_mod_change_abs_1_var_sessilemobile, col_mod_change_abs_1_var)
```
Including sessile/mobile improves our ability to predict likelihood of colonization

What about extinction?
```{r}
ext_mod_change_lag2_abs_1_var_sessilemobile <- glmer(now_ext ~ sst_max_change_lag2_abs*fish + (1 | reg) + (1 | year), family = binomial, data = spp_master_ztemp_fishnofish)
summary(ext_mod_change_lag2_abs_1_var_sessilemobile)
AICc(ext_mod_change_lag2_abs_1_var, ext_mod_change_lag2_abs_1_var_sessilemobile)
```
Here, sessile/mobile also improves fit and also the interaction with change in maximum temperature. The role of mobility changes with the size of the change in temperature between years. 