---
title: "Extracting SODA Temperature from both 2.2.4 and 3.3.1"
output: html_notebook
---
What lat/lon values do we actually need?
```{r make list of all GPS points we want to extract temp data from}
library(data.table)
load("/Users/zoekitchel/Documents/grad school/Rutgers/Coursework/Fall 2018/SDM/hauls_catch_Dec2017 1.RData")
#load("~/col_ext/hauls_catch_Dec2017 1.RData")
hauls <- data.table(hauls)
cols <- c("lat",  "lon",  "region", "depth")
latlongyear <- hauls[,  cols,  with = F]

# set the key to all columns
setkey(latlongyear)

# Get unique lat lon combos
latlongyear.u <- unique(latlongyear[list(lat,  lon),  nomatch = 0])

#only take maximum depth
latlongyear.u_maxdepth <- latlongyear.u[latlongyear.u[, .I[which.max(depth)], by = c("lat","lon")]$V1]

#minimum lat?,  maximum lat? minimum long? maximum long?
summary(latlongyear.u_maxdepth)

minlat <- min(latlongyear.u_maxdepth$lat)
maxlat <- max(latlongyear.u_maxdepth$lat)
minlon <- min(latlongyear.u_maxdepth$lon)
maxlon <- max(latlongyear.u_maxdepth$lon)
maxdepth <- max(latlongyear.u_maxdepth$depth)
mindepth <- min(latlongyear.u_maxdepth$depth)

minlat
maxlat
minlon
maxlon
maxdepth
mindepth
```

We need to follow Jim's lead,  and use SODA 2.2 for pre1980,  and then SODA 3.3 for post1980. Unfortunately,  they're accessible in different ways. We will start with1980+

```{bash import SODA post1980}
#install nco
conda install-c conda-forge nco
#make text file from https://www.atmos.umd.edu/~ocean/index_files/soda3.3.1_mn_ocean_reg.txt
#add "/ocean"after regrided error on their part

wget-i soda3.3.1_mn_ocean_reg_import.txt

```

This takes a while,  as it's100 gb of data and no way to narrow down beforehand as far as I can tell. 
**Downloaded: 36 files, 118G in 46m13s 43.6 MB/s**

```{bash extract just top/bottom layers from1980 to 2015 for each yearly file}
#bottom
for i in `cat soda3.3.1_mn_ocean_reg_import.txt`;do ncks -d depth,-1 "${i}" "${i%.*}_bottom.nc";done

#surface
for i in `cat soda3.3.1_mn_ocean_reg_import.txt`;do ncks -d depth,0 "${i}" "${i%.*}_surface.nc";done #in python, 0 is the first dimension

```

Seeing now that these files in their current state actually contain multiple variables, and I only want one temp
```{bash extracting only temp from 1980 to 2015}
#bottom
for i in `cat soda3.3.1_mn_ocean_reg_import.txt`; do ncks -v temp -O "${i%.*}_bottom.nc" "${i%.*}_sbt.nc";done
#surface
for i in `cat soda3.3.1_mn_ocean_reg_import.txt`; do ncks -v temp -O "${i%.*}_surface.nc" "${i%.*}_sst.nc";done
```

I will now merge these files along time. Ensure that there are no other files that match `soda3.3.1*sst.nc` syntax because then it won't work.
```{bash merge1980-2015 files for sbt and sst}
ncrcat soda3.3.1*sst.nc -O soda3.3.1_1980_2015_sst.nc
ncrcat soda3.3.1*sbt.nc -O soda3.3.1_1980_2015_sbt.nc
```

Let's now crop down to actual points we need:
24.49008 lat
62.01917 lat
-189.397 lon
-43.835 lon

But, currently, this 3.3.1 data is in 0-360, or technically 0.25-359.75 which is equal to 0.25 to -0.25

To match what we will extract from 2.2.4, we need -189.75E and -43.25E. So, we need to keep 170.25-316.75**

Let's crop into two pieces, and then remerge.
Crop to correct lat lon
```{bash crop points to those we actually need from 0-360 from 3.3.1}
#surface
ncks -O -d latitude,24.25,62.25 -d longitude,170.25,316.75 soda3.3.1_1980_2015_sst.nc soda3.3.1_1980_2015_sst_crop.nc


#bottom
ncks -O -d latitude,24.25,62.25 -d longitude,170.25,316.75 soda3.3.1_1980_2015_sbt.nc soda3.3.1_1980_2015_sbt_crop.nc

```


*The first command ncks shifts the data,  the second command ncap2 recalibrates the coordinate to the newly shifted data. One comment on applying this algorithm: Be careful to specify hemispheres that do not overlap,  e.g.,  by inadvertently specifying coordinate ranges in the first command that both include the date line. Some users will find using index-based rather than coordinate-based hyperslabs makes this clearer. Examine a plot of the field to make sure your rotation was correct.*

```{bash convert grid}
#surface
ncap2 -O -s 'longitude=longitude-360' soda3.3.1_1980_2015_sst_crop.nc soda3.3.1_1980_2015_sst_final.nc

#bottom
ncap2 -O -s 'longitude=longitude-360' soda3.3.1_1980_2015_sbt_crop.nc soda3.3.1_1980_2015_sbt_final.nc
```


Change names of dimensions to match earlier time files
```{bash change names of dimensions and variables for1980+}
#surface
ncrename -d latitude,lat -v latitude,lat soda3.3.1_1980_2015_sst_final.nc
ncrename -d longitude,lon -v longitude,lon soda3.3.1_1980_2015_sst_final.nc
#bottom
ncrename -d latitude,lat -v latitude,lat soda3.3.1_1980_2015_sbt_final.nc
ncrename -d longitude,lon -v longitude,lon soda3.3.1_1980_2015_sbt_final.nc
```
*Done with1980-2015*

----------
Malin recommended extracting bottom temperature by extracting from given depth value at each GPS location from trawl rather than trying to create a bottom layer, so I will also merge but keep all depths, so I will also try this way.Surface is fine, but I have to make the bottom layer in a different way.

Maximum depth = 1504 meters, so only up to 1625.7 in SODA data
```{bash limit to depths included in trawl survey 3.3.1}
#limit depth to 1625.7 (closest to max in trawl survey)
for i in `cat soda3.3.1_mn_ocean_reg_import.txt`;do ncks -O -d depth,0.,1625.7 "${i}" "${i%.*}_bottom_trawl.nc";done
#I should keep this depth max even if I don't use Malin's recommendation because bottom temperature below this point (deepest trawl) is illogical

#only look at temperature variable
for i in `cat soda3.3.1_mn_ocean_reg_import.txt`; do ncks -v temp -O "${i%.*}_bottom_trawl.nc" "${i%.*}_sbt_trawl.nc";done

#now merge these files over time
ncrcat soda3.3.1*sbt_trawl.nc -O soda3.3.1_1980_2015_sbt_trawl.nc

#crop to points we need lat lon
ncks -O -d latitude,24.25,62.25 -d longitude,170.25,316.75 soda3.3.1_1980_2015_sbt_trawl.nc soda3.3.1_1980_2015_sbt_trawl_crop.nc

#change grid from 0 360 - -180-180
ncap2 -O -s 'longitude=longitude-360' soda3.3.1_1980_2015_sbt_trawl_crop.nc soda3.3.1_1980_2015_sbt_trawl_final.nc

#change names to lat/lon
ncrename -d latitude,lat -v latitude,lat soda3.3.1_1980_2015_sbt_trawl_final.nc
ncrename -d longitude,lon -v longitude,lon soda3.3.1_1980_2015_sbt_trawl_final.nc
```
---------

**********

Now,  we will import netcdf data for pre1980 from SODA 2.4.

Unfortunately,  It looks like I'm going to have to import files one by one for chunks of depth.

Text file with links: /Scratch/zoek_scratch/soda2.2.4_mn_ocean_reg.txt

In order to give a buffer and correctly download from iridl.ldeo.columbia.edu, use -190E- -43E
SODA 2.2.4 gives -189.75E and -43.25E which is equal to 189.75W and 43.25W OR 170.25E and -43.25E

I made text file in amphiprion to bulk download `wget-i soda2.2.4_mn_ocean_reg.txt--no-check-certificate` DO THIS
file names going in as data.nc.x etc.
5.01-70.02: data.nc
82.92-197.79: data.nc.1
229.48-729.35: data.nc.2
918.37-2375.0: data.nc.3
2625.0-4375.0: data.nc.4
4625.0-5375.0: data.nc.5

```{bash merge SODA files pre1980}
#merge all depth chunks
cdo merge data.n* data.pre1980.nc

#only bottom layer
ncks -d depth,-1 data.pre1980.nc soda2.2.4_1953_1979_sbt.nc #NCO ncks with a negative hyperslab

#only surface layer
ncks -d depth,1 data.pre1980.nc soda2.2.4_1953_1979_sst.nc
```
Get rid of depth dimension for surface temp files
```{bash delete depth dimension and variable for surface temp files}
ncwa -a depth soda2.2.4_1953_1979_sst.nc soda2.2.4_1953_1979_sst.nc
ncwa -a depth soda3.3.1_1980_2015_sst_final.nc soda3.3.1_1980_2015_sst_final.nc

ncks -x -v depth soda2.2.4_1953_1979_sst.nc soda2.2.4_1953_1979_sst.nc
ncks -x -v depth soda3.3.1_1980_2015_sst_final.nc soda3.3.1_1980_2015_sst_final.nc

```
We have to match time units, soda2.2.4 in 360 day and 3.3.1 in gregorian
START HERE, maybe import first, and then play with time
```{bash fixing time units}
cdo -a setcalendar,standard soda3.3.1_1980_2015_sst_final.nc soda3.3.1_1980_2015_sst_final_timefix.nc
cdo -a setcalendar,standard soda3.3.1_1980_2015_sbt_final.nc soda3.3.1_1980_2015_sbt_final_timefix.nc
cdo -a setcalendar,standard soda2.2.4_1953_1979_sst.nc soda2.2.4_1953_1979_sst_timefix.nc
cdo -a setcalendar,standard soda2.2.4_1953_1979_sbt.nc soda2.2.4_1953_1979_sbt_timefix.nc
```
----------
Malin recommended extracting bottom temperature by extracting from given depth value at each GPS location from trawl rather than trying to create a bottom layer, so I will also merge but keep all depths, so I will also try this way.Surface is fine, but I have to make the bottom layer in a different way.

Maximum depth = 1504 meters, so only up to 1625.7 in SODA data
```{bash limit to depths included in trawl survey 2.2.4}
#limit depth to 1625.7 (closest to max in trawl survey)
ncks -O -d depth,0.,1625.7 data.pre1980.nc soda2.2.4_1953_1979_sbt_trawl_final.nc
#I should keep this depth max even if I don't use Malin's recommendation because bottom temperature below this point (deepest trawl) is illogical
```

---------


Can't figure the time thing out, I will upload what I have.
soda3.3.1_1980_2015_sst_final_timefix.nc*
soda3.3.1_1980_2015_sbt_trawl_final_timefix.nc*
soda2.2.4_1953_1979_sst.nc
soda2.2.4_1953_1979_sbt_trawl_final.nc


*see below for timefix specifics
```{bash enumerating 3.3.1 months sequentially}
#I want it to be months since ""months since 1960-01-01""
#1980-01-01 (first value) is 241.5 months since 1960-01-01
ncap2 -s 'time=array(241.5,1,$time)' soda3.3.1_1980_2015_sst_final.nc soda3.3.1_1980_2015_sst_final_timefix.nc
ncap2 -s 'time=array(241.5,1,$time)' soda3.3.1_1980_2015_sbt_trawl_final.nc soda3.3.1_1980_2015_sbt_trawl_final_timefix.nc

```

************

I didn't use this link,  but it appears to be helpful. [ERDAPP]https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdSoda331oceanmday.html.

************

Now I have 2 files for surface temperature (3 dimensions) and surface-deepest trawls (bottom temperature, 4 dimensions). Surface temperature should be straight forward enough. However, I think the next step for bottom temperature is to split them both into monthly files, and these will be of three dimensions (depth, lat, lon). Then, I can extract depth, lat, lon, data and loop through all month files.

Let's start with extracting GPS points from surface temperature, this should be easy enough.
```{r extract temp values at GPS points for surface temperature}
#First, pre 1980

# =========================================
# = Function to Read in SODA, Grab Surface = by Ryan
# =========================================
get.soda <- function(file){

	soda.info <- nc_open(file)
	name.soda.sizes <- sapply(soda.info$var$temp$dim, function(x)x$name)
	soda.sizes <- soda.info$var$temp$size
	dim.units <- sapply(soda.info$var$temp$dim, function(x)x$units)
	print(dim.units)
	stopifnot(grepl("months since ", dim.units[3])) # make sure time is in correct units and in right place
	names(soda.sizes) <- name.soda.sizes
	ntime <- soda.sizes["time"]
	ndepth <- soda.sizes["depth"]

	soda.time0 <- soda.info$var$temp$dim[[3]]$vals
	ref.date <- as.Date(gsub("months since ", "", dim.units[3]))
	start.before.ref <- grepl("-", soda.time0[1]) # is the first date before ref.date?
	n.month.before <- ceiling(abs(soda.time0[1])) + as.integer(start.before.ref)
	start.increment <- ifelse(start.before.ref, "-1 month", "1 month")
	time.start <- rev(seq.Date(ref.date, by=start.increment, length.out=n.month.before))[1]
	soda.time <- seq.Date(time.start, by="1 month", length.out=ntime)
	
	soda <- brick(file)
	names(soda) <- soda.time

		
	return(soda)
	
}

soda_surface_2.2.4 <- get.soda("soda2.2.4_1953_1979_sst.nc")

#Now, post 1980

# =========================================
# = Function to Read in SODA, Grab Surface = by Ryan (NOTICE EDITS FROM ABOVE BECAUSE OF TIME)
# =========================================
get.soda.3.3.1 <- function(file){

	soda.info <- nc_open(file)
	name.soda.sizes <- sapply(soda.info$var$temp$dim, function(x)x$name)
	soda.sizes <- soda.info$var$temp$size
	dim.units <- sapply(soda.info$var$temp$dim, function(x)x$units)
	print(dim.units)
	names(soda.sizes) <- name.soda.sizes
	ntime <- soda.sizes["time"]

	time.start <- as.Date("1980-01-01") #set this to date you want to start at
	soda.time <- seq.Date(time.start, by="1 month", length.out=ntime)
	
	soda <- brick(file)
	names(soda) <- soda.time

		
	return(soda)
	
}

soda_surface_3.3.1 <- get.soda.3.3.1("soda3.3.1_1980_2015_sst_final_timefix.nc")

```

Now, in R, merge these two raster bricks
```{r merge stacks for surface temperature across time}
soda_surface <- stack(soda_surface_2.2.4, soda_surface_3.3.1)
#save
writeRaster(soda_surface, "soda_surface_sep2019.grd", overwrite = T)
```
Done with surface temperature.

Now, we have to do the same for bottom temperature, will use Ryan and Becca's code.

First, 2.2.4 `soda2.2.4_1953_1979_sbt_trawl_final.nc`
```{r bring in bottom 2.2.4}
# =========================================
# = Function to Read in SODA, Grab Bottom =
# =========================================
get.soda.bot.2.2.4 <- function(file){

	soda.info <- nc_open(file)
	name.soda.sizes <- sapply(soda.info$var$temp$dim, function(x)x$name)
	soda.sizes <- soda.info$var$temp$size
	dim.units <- sapply(soda.info$var$temp$dim, function(x)x$units)
	print(dim.units)
	stopifnot(grepl("months since ", dim.units[4])) # make sure time is in correct units and in right place
	names(soda.sizes) <- name.soda.sizes
	ntime <- soda.sizes["time"]
	ndepth <- soda.sizes["depth"]

	soda.time0 <- soda.info$var$temp$dim[[4]]$vals
	ref.date <- as.Date(gsub("months since ", "", dim.units[4]))
	start.before.ref <- grepl("-", soda.time0[1]) # is the first date before ref.date?
	n.month.before <- ceiling(abs(soda.time0[1])) + as.integer(start.before.ref)
	start.increment <- ifelse(start.before.ref, "-1 month", "1 month")
	time.start <- rev(seq.Date(ref.date, by=start.increment, length.out=n.month.before))[1]
	soda.time <- seq.Date(time.start, by="1 month", length.out=ntime)


	pb <- txtProgressBar(min=1, max=ntime, style=3)
	for(i in 1:ntime){
		t.soda <- brick(file, lvar=4, level=i)
		# need to switch missing value to actual NA # TODO is this automatically done by brick()? I think so.
		
		soda.depths <- as.numeric(gsub("X", "", names(t.soda))) # does weird rounding that I don't understand
	
		# get the deepest available temperature in each gridd cell
		t.soda.bot <- do.call(cover, unstack(subset(t.soda, length(soda.depths):1)))
		names(t.soda.bot) <- soda.time[i]
		# the subsetting piece flips the order of the layers (so that deepest is first layer)
		# by unstacking I reformat from raster with layers, to list of rasters
		# that list can be used as the list of arguments to a function taking ...
		# cover keeps the values in the first object of the ..., and replaces the NA's with values from the second ...
		# that process repeats until through all object listed in the ...
		# in other words, the final value will be the value in the first layer that has a non-NA value in that cell
		
		# accumulate over time periods (monthly in original data sets I'm using)
		# this will be for the first time period (1958-1978)
		if(i==1){
			soda.bot <- t.soda.bot
		}else{
			soda.bot <- addLayer(soda.bot, t.soda.bot)
		}
		setTxtProgressBar(pb, i)
	}
	close(pb)
	
	return(soda.bot)
	
}

soda_bottom_2.2.4 <- get.soda.bot.2.2.4("soda2.2.4_1953_1979_sbt_trawl_final.nc")

```

Then, 3.3.1 `soda3.3.1_1980_2015_sbt_trawl_final_timefix.nc`
```{r bring in bottom 3.3.1}
# =========================================
# = Function to Read in SODA, Grab Bottom =
# =========================================
get.soda.bot.3.3.1 <- function(file){

	soda.info <- nc_open(file)
	name.soda.sizes <- sapply(soda.info$var$temp$dim, function(x)x$name)
	soda.sizes <- soda.info$var$temp$size
	dim.units <- sapply(soda.info$var$temp$dim, function(x)x$units)
	print(dim.units)
	names(soda.sizes) <- name.soda.sizes
	ntime <- soda.sizes["time"]
	ndepth <- soda.sizes["depth"]

  time.start <- as.Date("1980-01-01") #set this to date you want to start at
	soda.time <- seq.Date(time.start, by="1 month", length.out=ntime)


	pb <- txtProgressBar(min=1, max=ntime, style=3)
	for(i in 1:ntime){
		t.soda <- brick(file, lvar=4, level=i)
		# need to switch missing value to actual NA # TODO is this automatically done by brick()? I think so.
		
		soda.depths <- as.numeric(gsub("X", "", names(t.soda))) # does weird rounding that I don't understand
	
		# get the deepest available temperature in each gridd cell
		t.soda.bot <- do.call(cover, unstack(subset(t.soda, length(soda.depths):1)))
		names(t.soda.bot) <- soda.time[i]
		# the subsetting piece flips the order of the layers (so that deepest is first layer)
		# by unstacking I reformat from raster with layers, to list of rasters
		# that list can be used as the list of arguments to a function taking ...
		# cover keeps the values in the first object of the ..., and replaces the NA's with values from the second ...
		# that process repeats until through all object listed in the ...
		# in other words, the final value will be the value in the first layer that has a non-NA value in that cell
		
		# accumulate over time periods (monthly in original data sets I'm using)
		# this will be for the first time period (1958-1978)
		if(i==1){
			soda.bot <- t.soda.bot
		}else{
			soda.bot <- addLayer(soda.bot, t.soda.bot)
		}
		setTxtProgressBar(pb, i)
	}
	close(pb)
	
	return(soda.bot)
	
}

soda_bottom_3.3.1 <- get.soda.bot.3.3.1("soda3.3.1_1980_2015_sbt_trawl_final_timefix.nc")

```

Now, in R, merge these two raster bricks
```{r merge stacks for bottom temperature across time}
soda_bottom <- stack(soda_bottom_2.2.4, soda_bottom_3.3.1)
#save
writeRaster(soda_bottom, "soda_bottom_sep2019.grd", overwrite = T)
```