---
title: "Extracting SODA Temperature Update"
output: html_notebook
---
We need to follow Jim's lead, and use SODA 2.4 for pre 1980, and then SODA 3.3 for post 1980. Unfortunately, they're accessible in different ways. We will start with 1980+

```{terminal import SODA post 1980}
#install nco
conda install -c conda-forge nco
#make text file from https://www.atmos.umd.edu/~ocean/index_files/soda3.3.1_mn_ocean_reg.txt
#add "/ocean"after regrided (error on their part)

wget -i soda3.3.1_mn_ocean_reg_import.txt

```

This takes a while, as it's 100 gb of data and no way to narrow down beforehand as far as I can tell. 
**Downloaded: 36 files, 118G in 46m 13s (43.6 MB/s)**

```{terminal merge these files post 1980 into one}
ncrcat -h soda3.3.1_mn_ocean_reg_*.nc soda3.3.1_mn_ocean_reg_1980_2015.nc
```
This merge doesn't seem possible, so seems better to loop rather than merge

```{terminal extract just top/bottom layers from 1980 to 2015}
#bottom?
for i in `cat soda3.3.1_mn_ocean_reg_import.txt`;do ncks -d depth,-1 "${i}" "${i%.*}_bottom.nc";done

#top?
for i in `cat soda3.3.1_mn_ocean_reg_import.txt`;do ncks -d depth,1 "${i}" "${i%.*}_surface.nc";done
```

For top, and to actually extract bottom, I think I can use Ryan's code
```{r 1980+ extract bottom and top temp}
#Ryan's function to extract from nc_open
get.soda.sst <- function(file){
  
  soda.info <- nc_open(file)
  name.soda.sizes <- sapply(soda.info$var$temp$dim, function(x)x$name)
  soda.sizes <- soda.info$var$temp$size
  dim.units <- sapply(soda.info$var$temp$dim, function(x)x$units)
  print(dim.units)
  stopifnot(grepl("months since ", dim.units[4])) # make sure time is in correct units and in right place
  names(soda.sizes) <- name.soda.sizes
  ntime <- soda.sizes["time"]
  ndepth <- soda.sizes["depth"]
  
  soda.time0 <- soda.info$var$temp$dim[[4]]$vals
  ref.date <- as.Date(gsub("months since ", "", dim.units[4]))
  start.before.ref <- grepl("-", soda.time0[1]) # is the first date before ref.date?
  n.month.before <- ceiling(abs(soda.time0[1])) + as.integer(start.before.ref)
  start.increment <- ifelse(start.before.ref, "-1 month", "1 month")
  time.start <- rev(seq.Date(ref.date, by=start.increment, length.out=n.month.before))[1]
  soda.time <- seq.Date(time.start, by="1 month", length.out=ntime)
  
  soda.sst <- brick(file)
  names(soda.sst) <- soda.time
  
  
  return(soda.sst)
  
}

## Using Ryan's code
soda_sst <- get.soda.sst("col_ext/CARTON-GIESE_SODA_v2p1p6_sstemp.nc")


```



Now, let's average temperature values for every month over all GPS points in each region
```{r make list of all GPS points we want to extract temp data from, and then average}
library(data.table)
load("/Users/zoekitchel/Documents/grad school/Rutgers/Coursework/Fall 2018/SDM/hauls_catch_Dec2017 1.RData")
#load("col_ext/hauls_catch_Dec2017 1.RData")
hauls <- data.table(hauls)
cols <- c("lat", "lon", "region")
latlongyear <- hauls[, cols, with = F]

# set the key to all columns
setkey(latlongyear)

# Get unique lat lon combos
latlongyear.u <- unique(latlongyear[list(lat, lon), nomatch = 0])

```
Now, we will import OPENDAP files (netcdf off web) for pre 1980 from SODA 2.4. But I will have to run this in terminal.
```{r import SODA pre 1980}
conda activate /local/home/zoek/enter/envs/myRenv3_5
R
library(ncdf4)
library(raster)
```
Unfortunately, It looks like I'm going to have to import files one by one.
I made text file in amphiprion to bulk download `wget -i soda2.2.4_mn_ocean_reg.txt --no-check-certificate`
file names going in as data.nc.x etc.
5.01-70.02: data.nc
82.92-197.79: data.nc.1
229.48-729.35: data.nc.2
918.37-2375.0: data.nc.3
2625.0-4375.0: data.nc.4
4625.0-5375.0: data.nc.5

May need to merge in terminal
```{terminal merge SODA files pre 1980}
conda install -c conda-forge cdo #installing climate data operators
conda install -c conda-forge nco
cdo merge data.n* data.pre1980.nc

#only bottom layer
ncks -d depth,-1 in.nc out.nc #NCO ncks with a negative hyperslab
```

```{trying to import}
get.soda <- function(file){

	soda.info <- nc_open(file)
	name.soda.sizes <- sapply(soda.info$var$temp$dim, function(x)x$name)
	soda.sizes <- soda.info$var$temp$size
	dim.units <- sapply(soda.info$var$temp$dim, function(x)x$units)
	print(dim.units)
	stopifnot(grepl("months since ", dim.units[4])) # make sure time is in correct units and in right place
	names(soda.sizes) <- name.soda.sizes
	ntime <- soda.sizes["time"]
	ndepth <- soda.sizes["depth"]

	soda.time0 <- soda.info$var$temp$dim[[4]]$vals
	ref.date <- as.Date(gsub("months since ", "", dim.units[4]))
	start.before.ref <- grepl("-", soda.time0[1]) # is the first date before ref.date?
	n.month.before <- ceiling(abs(soda.time0[1])) + as.integer(start.before.ref)
	start.increment <- ifelse(start.before.ref, "-1 month", "1 month")
	time.start <- rev(seq.Date(ref.date, by=start.increment, length.out=n.month.before))[1]
	soda.time <- seq.Date(time.start, by="1 month", length.out=ntime)
	
	soda <- brick(file)
	names(soda) <- soda.time

		
	return(soda)
}

soda_pre1980 <- get.soda(data.nc)

	
```

```{merge SODA files pre 1980}
data.nc <- nc_open('data.nc')
data.nc.1 <- nc_open('data.nc.1')
data.nc.2 <- nc_open('data.nc.2')
data.nc.3 <- nc_open('data.nc.3')
data.nc.4 <- nc_open('data.nc.4')
data.nc.5 <- nc_open('data.nc.5')

data.nc.merge <- cbind(
  ncvar_get(data.nc, 'temp'),
  ncvar_get(data.nc.1, 'temp'),
  ncvar_get(data.nc.2, 'temp'),
  ncvar_get(data.nc.3, 'temp'),
  ncvar_get(data.nc.4, 'temp'),
  ncvar_get(data.nc.5, 'temp'))
dim(data.nc.merge)
var <- data.nc$var['temp']$temp

#new file
data.nc.pre1980 <- nc_create(
  filename = 'SODA_monthly_2.2.4_1953_1979.nc',
  #define variables
  vars = ncvar_def(
    name = 'temp',
    units = var$units,
    dim = dim(data.nc.merge)))

#write to it
ncvar_put(
  nc = data.nc.pre1980,
  varid = 'temp',
  vals = data.nc.merge)

#close
nc_close(data.nc.pre1980)

```
Now I only want surface and bottom temperature for this.

```{r import SODA pre 1980 sst}
# =========================================
# = Function to Read in SODA, Grab Surface =
# =========================================
get.soda.sst <- function(file){

	soda.info <- nc_open(file)
	name.soda.sizes <- sapply(soda.info$var$temp$dim, function(x)x$name)
	soda.sizes <- soda.info$var$temp$size
	dim.units <- sapply(soda.info$var$temp$dim, function(x)x$units)
	print(dim.units)
	stopifnot(grepl("months since ", dim.units[4])) # make sure time is in correct units and in right place
	names(soda.sizes) <- name.soda.sizes
	ntime <- soda.sizes["time"]
	ndepth <- soda.sizes["depth"]

	soda.time0 <- soda.info$var$temp$dim[[4]]$vals
	ref.date <- as.Date(gsub("months since ", "", dim.units[4]))
	start.before.ref <- grepl("-", soda.time0[1]) # is the first date before ref.date?
	n.month.before <- ceiling(abs(soda.time0[1])) + as.integer(start.before.ref)
	start.increment <- ifelse(start.before.ref, "-1 month", "1 month")
	time.start <- rev(seq.Date(ref.date, by=start.increment, length.out=n.month.before))[1]
	soda.time <- seq.Date(time.start, by="1 month", length.out=ntime)
	
	soda.sst <- brick(file)
	names(soda.sst) <- soda.time

		
	return(soda.sst)
	
}

#still may need to extract first column somewhere

```

And now, the challenge, bottom temperature

```{r bottom temp SODA}
bottom <- apply(!is.na(haul_temp), 1, sum)
soda.sbt <- haul_temp[cbind(1:nrow(haul_temp), bottom)]
```


