---
title: "Colonization Extinction Summary"
output: html_notebook
---
```{r setup, include = F}

library(data.table)


load(file = "colonization_model_comparisons_dataframe.Rdata")
```

```{r manipulating colonization models, include = F}
#only use top 10 models
colonization_model_comparisons.r <- colonization_model_comparisons.nomean[order(AICc)][1:10,]
cols <- names(colonization_model_comparisons.r)[2:5]
colonization_model_comparisons.r[,(cols) := round(.SD,3), .SDcols=cols]
colonization_model_comparisons.r
```


##Colonization Models

variable | coefficient | Pr(>|z|) | AICc | $\Delta$ AICc |
---------|-------------|----|-----|---|---|
| max_sbt_temp_change_abs |	0.239 |	0.000 |	6878.288 |	0.000 |
| max_sst_temp_change_abs_lag1 |	-0.240 |	0.001 |	6887.400 |	9.112 |
| min_sbt_temp_change_abs_lag5 |	-0.244 |	0.002 |	6888.859 |	10.572 |
| max_sbt_temp_change_abs_lag3 |	0.171 |	0.001 |	6888.938 |	10.650 |
| seas_sbt_temp_change_abs |	0.142 |	0.002 |	6889.498 |	11.211 |
| max_sst_temp_change_abs_lag8 |	0.191 |	0.002 |	6889.901 |	11.614 |
| seas_sst_temp_change_abs_lag8 |	0.127 |	0.004 |	6890.941 |	12.654 |
| seas_sbt_temp_change_abs_lag4 |	0.136 |	0.005 |	6891.397 |	13.109 |
| max_sbt_temp_change_lag8 |	-0.115 |	0.006 |	6891.566 |	13.279 |
| seas_sbt_temp_change_abs_lag1 |	0.126 |	0.007 |	6891.960 |	13.672 |

Relative Variable Importance 
Calculated Akaike Weight for each model, divided by sum of all weights to get a value for relative model importance. Then, summed values for all models including the following variables. This only works when each variable is equally represented.

* Surface (n = 93) importance = 0.02 **versus** Bottom (n = 93) importance = 0.97
* Non lagged (n = 18) importance = 0.96 **versus** Lag (n = 168) importance = 0.03 (despite difference in n)
* Absolute change (n = 60) importance = 0.99 **versus** Raw temp (n = 66) importance = 0.003 **versus** change (n = 60) = 0.005


```{r manipulating extinction models, include = F}
#only use top 10 models
extinction_model_comparisons.r <- extinction_model_comparisons.nomean[order(AICc)][1:10,]
cols <- names(extinction_model_comparisons.r)[2:5]
extinction_model_comparisons.r[,(cols) := round(.SD,3), .SDcols=cols]
extinction_model_comparisons.r
```

##Extinction Models

variable | coefficient | Pr(>|z|) | AICc | $\Delta$ AICc |
---------|-------------|----|-----|---|---|
| seas_sbt_temp_change_abs_lag9 |	0.168 |	0.002 |	6207.213 |	0.000 |
| min_sst_temp_change_lag2 |	-0.095 |	0.011 |	6209.860 |	2.647 |
| min_sbt_temp_change_lag7 |	-0.116 |	0.011 |	6210.022 |	2.808 |
| seas_sbt_temp_change_abs |	0.116 |	0.013 |	6210.184 |		2.971 |
| max_sbt_temp_change_lag7 |	-0.100 |	0.015 |	6210.285 |	3.072 |
| seas_sbt_temp_lag9 |	0.087 |	0.017 |	6210.766 |	3.553 |
| seas_sst_temp_change_lag2 |	0.069 |	0.020 |	6210.908 |	3.695 |
| max_sbt_temp_change_abs_lag9 |	0.177 |	0.028 |	6211.488 |	4.275 |
| max_sst_temp_change_lag1 |	-0.098 |	0.038 |	6212.029 |	4.816 |
| max_sbt_temp_change_abs_lag3 |	0.117 |	0.043 |	6212.264 |	5.050 |

Relative Variable Importance

* Surface (n = 93) importance = 0.37 **versus** Bottom (n = 93) importance = 0.63
* Non lagged (n = 18) importance = 0.10 **versus** Lag (n = 168) importance = 0.90 (keep in mind change in n)
* Absolute change (n = 60) importance = 0.47 **versus** Raw temp (n = 66) importance = 0.20 **versus** change (n = 60) = 0.34

Comparisons in lags between colonizations and extinctions
```{r RVI lags, echo = F}
load("RVI_lags.Rdata")
RVI_col_plot
RVI_ext_plot
RVI_lags + scale_x_continuous(breaks = c(0:9))
```


##Traits

###Maximum Age

Best Colonization Models

| | Variable | Coefficient | Pr(>|z|) | AICc | Rank for Full Models | Coefficient for Full Models
|--|---------|-------------|------|-------|----------------|-----------------|
| 1 | seas_sst_temp_change_abs |	0.16747228 |	0.001059902 |	4239.108 |	12 | 0.104 |
| 2 | max_sbt_temp_change_abs_lag3 |	0.19569357 |	0.001885719 |	4239.951 | 4	| 0.171 |
| 3 | max_sst_temp_change_abs_lag1 |	-0.26280961 |	0.003919178 |	4240.645 | 2 | -0.24 |
| *4 | max_sbt_temp_change_abs |	0.16256132 |	0.009148480 |	4242.650 | 1 | 0.239 |
| 5 | max_sbt_temp_change_abs_lag1 |	0.16616085 |	0.010718533 |	4242.893 | 11 | 0.14 | 
| 6 | seas_sbt_temp_change_abs |	0.13226661 |	0.016386407 |	4243.561 | 5 | 0.14 
| 7 | seas_sbt_temp_change_abs_lag4 |	0.14075757 |	0.019188279 |	4243.750 | 8 | 0.14
| 8 | max_sst_temp_change_abs_lag6 |	-0.20512821 |	0.024157883 |	4243.823 |	not in top 15 | |
| 9 | seas_sbt_temp_change_abs_lag1 |	0.12951678 |	0.026914136 |	4244.384 | 9 | 0.13 |	
| 10 | max_sst_temp_change_abs_lag8 |	0.17240059 |	0.028828987 |	4244.709 | 6 |	0.19
| 11 | seas_sst_temp_change_abs_lag6 |	-0.12608343 |	0.043620010 |	4244.899 | not in top 15 | |	
| 12 | max_sst_temp_change_abs |	0.19930194 |	0.042940256 |	4245.196 | 14 | 0.17 |	
| 13 | seas_sst_temp_change_lag5 |	-0.07459168 |	0.049345998 |	4245.232 | not in top 15 | |	
| 14 | seas_sbt_temp_lag10 |	0.08104661 |	0.048357908 |	4245.336 | not in top 15 | |	
| 15 | min_sbt_temp_change_abs_lag5 |	-0.18602894 |	0.060853920 |	4245.388 | 3 | -0.24 |	
**Used for models below. Should I instead use top ranking model here?*
In top 15, 11/15 shared. 

Looked at:
* col ~ temp
* col ~ temp + max.age
* col ~ temp * max.age
* col ~ max.age

Best model includes interaction, (Δ = 1.34 vs w/o interaction)
  variable                        coefficient Pr(>|z|) AICc
  max_sbt_temp_change_abs          0.055470    0.5295   4203.399
  age.max                         -0.031236  1.15e-07 ***
  max_sbt_temp_change_abs:age.max  0.008543    0.0922 .  

As max age increases, impact of temperature on log odds of colonization increases. As max age increases, log odds of colonization decreases. 

Best Extinction Models

| | Variable | Coefficient | Pr(>|z|) | AICc | Rank for Full Models | Coefficient for Full Models
---------|-------------|-----------|----------|---------|-----------|
| *1 | min_sst_temp_change_lag2 |	-0.14436022 |	0.003038827 |	3930.580 | 2 | -0.09 |	
| 2 |  seas_sbt_temp_change_lag9 |	0.12993554 |	0.005210505 |	3931.628 | not in top 15 | |	
| 3 |  seas_sbt_temp_lag9 |	0.10972665 |	0.013229764 |	3933.519 | 6 | 0.09 |	
| 4 |  max_sbt_temp_change_abs |	0.15427483 |	0.017742274 |	3933.821 | not in top 15 | |	
| 5 |  min_sst_temp_change_abs_lag6 |	0.16578612 |	0.020123707 |	3934.268 | not in top 15 | |	
| 6 |  min_sbt_temp_change_abs_lag5 |	-0.23173199 |	0.036076360 |	3934.596 | not in top 15| |	
| 7 | seas_sbt_temp_change_abs |	0.12245319 |	0.032188793 |	3934.761 | 4 | 0.12 |	
| 9 |  max_sbt_temp_change_lag9 |	0.11439746 |	0.035588475 |	3934.905 | not in top 15| |	
| 10 | seas_sst_temp |	-0.09460166 |	0.050126076 |	3935.655 | not in top 15 | |	
| 11 | min_sst_temp_change_abs_lag5 |	-0.15868171 |	0.063866384 |	3935.761 | not in top 15 | |	
| 12 | seas_sst_temp_lag2 |	0.07970154 |	0.044236722 |	3935.820 | 12 |  0.06 |	
| 13 |	seas_sst_temp_change_abs_lag5 |	-0.11465158 |	0.071676787 |	3935.918 | not in top 15 | |	
| 14 | seas_sst_temp_change_lag2 |	0.06818593 |	0.069610050 |	3936.028 | 7 | 0.07 |	
| 15 | seas_sst_temp_change |	-0.07118644 |	0.077429810 |	3936.153 | not in top 15 | |	

Of top 15 models, they only share 6... this is a problem? Seems like *min_sst_temp_change_lag2* is best to continue with?
*Continued with this model

Looked at:
* now_ext ~ temp
* now_ext ~ temp + age.max
* now_ext ~ temp * age.max
* now_ext ~ age.max

Best model includes both temp and max age but does not include interaction, but the AICcs are very close (Δ = 0.717 vs w interaction, Δ = 6.656 from other models)

                          Estimate Std. Error z value Pr(>|z|)       AICc
(Intercept)              -3.330222   0.228016 -14.605  < 2e-16 *** 3889.425
min_sst_temp_change_lag2 -0.144548   0.048814  -2.961  0.00306 ** 
age.max                  -0.029182   0.004982  -5.857  4.7e-09 ***

This tells us that as maximum age increases, log odds of going extinct decreases. As change in min temperature experienced between 2-3 years ago increases (positive change, towards warmer), the log odds of extinction decreases.


###Age at Maturity

Hypothesis: as age at maturity increases, odds of going extinct decreases because they likely have a longer generation time, and therefore the population will take longer to respond. 

Best Colonization Models

| | Variable | Coefficient | Pr(>|z|) | AICc | Rank for Full Models | Coefficient for Full Models
---------|-------------|-----------|------------|------------|--------|
| 1 | seas_sst_temp_change_abs | 	0.16732168 |	0.001055896 |	4237.134 | 12 | 0.10	| 
| 2 | max_sst_temp_change_abs_lag1	 | -0.26275332 |	0.004143103 |	4238.738 | 2 | -0.24 |
| 3 | max_sbt_temp_change_abs_lag3 | 	0.18628409 |	0.003083958 |	4238.804 | 4 | 0.17 |
| 4 | max_sst_temp_change_abs_lag6	 | -0.22361557 |	0.014782142 |	4240.889 | not in top 15 | |
| *5 | max_sbt_temp_change_abs | 	0.15801135 |	0.011635865 |	4241.039 | 1 | 0.24 |
| 6 | max_sst_temp_change_abs_lag8 | 	0.19598072 |	0.012668985 |	4241.366 | 6 | 0.19 |
| 7 | max_sbt_temp_change_abs_lag1 | 	0.15825795 |	0.014922452 |	4241.440 | 11 | 0.14 |
| 9 | seas_sbt_temp_change_abs | 	0.13079589 |	0.017830284 |	4241.690 | 5 | 0.14 | 
| 10 | seas_sbt_temp_change_abs_lag4 | 	0.13424702 |	0.024563505 |	4242.176 | 8 | 0.14 |
| 11 | seas_sst_temp_change_abs_lag6	 | -0.13386926 |	0.033532762 |	4242.414 | not in top 15 | |
| 12 | seas_sbt_temp_change_abs_lag1 | 	0.12498712 |	0.031649631 |	4242.646 | 10 | 0.13 |
| 13 | 	min_sbt_temp_change_abs_lag5	 | -0.19180573 |	0.054000799 |	4243.150 | 3 | -0.24 |
| 14 | seas_sst_temp_change_lag5	 | -0.07497315 |	0.049187743 |	4243.216 |	not in top 15 | |
| 15 | seas_sbt_temp_lag10 | 	0.07724898 |	0.059318089 |	4243.639 | not in top 15 ||
**Used for models below*
In top 15, 11/15 shared. 

Looked at:
* col ~ temp
* col ~ temp + age.maturity
* col ~ temp * age.maturity
* col ~ age.maturity

Best model includes temperature and age at maturity, but not an interaction ($/Delta$ = 1.94 vs w/ interaction):
  max_sbt_temp_change_abs  0.15503   0.0132 4236.753
  age.maturity            -0.03818   0.0171 
  
  
Extinction Model

| | Variable | Coefficient | Pr(>|z|) | AICc | Rank for Full Models | Coefficient for Full Models
|---|---------|-------------|-----------|------------|-----------|----------|
| 1 | min_sst_temp_change_lag2	| -0.14805497 |	0.002710857 |	3920.109 | 2 | -0.09 |	
| 2 |  seas_sbt_temp_change_lag9	| 0.12861632 |	0.006329875 |	3921.697 | not in top 15 | |	
| 3 |  seas_sbt_temp_lag9	| 0.11207541 |	0.011862688 |	3923.042 | 6 | 0.09 |	
| 4 |  max_sbt_temp_change_abs	| 0.15619946 |	0.016642456 |	3923.456 | not in top 15  | |	
| 5 |  min_sst_temp_change_abs_lag6	| 0.16612378 |	0.020897770 |	3924.076 | not in top 15 | |	
| 6 |  min_sbt_temp_change_abs_lag5	| -0.23921218 |	0.032703088 |	3924.162 | not in top 15 | |	
| 7 | seas_sbt_temp_change_abs	| 0.12215822 |	0.032750203 |	3924.536 | 4  | 0.12 |	
| 9 |  seas_sst_temp_lag2	| 0.08894160 |	0.021276070 |	3924.730 | 12 | 0.06 |	
| 10 | seas_sst_temp_change	| -0.07914067 |	0.053147439 |	3925.275 | not in top 15  | |	
| 11 | max_sbt_temp_change_lag9	| 0.10522327 |	0.055746506 |	3925.392 | not in top 15 | |	
| 12 | min_sst_temp_change_abs_lag5	| -0.16304438 |	0.059969373 |	3925.406 | not in top 15 | |	
| 13 |	 seas_sst_temp_change_lag2	| 0.07202288 |	0.059464663 |	3925.501 | 7 | 0.07 |	
| 14 | seas_sst_temp	| -0.09394699 |	0.055661858 |	3925.636 | not in top 15 | |	
| 15 | min_sst_temp_change	| 0.09385684 |	0.072201812 |	3925.738 | not in top 15  | |	

Of top 15, only 5 are shared

Best model includes both age at maturity and temp but not the interaction ($\Delta$  = 1.926 vs w/ interaction).

                         Estimate Std. Error z value Pr(>|z|)    AICc
(Intercept)              -3.54614    0.23907 -14.833  < 2e-16 *** 3906.268
min_sst_temp_change_lag2 -0.14810    0.04939  -2.998 0.002714 ** 
age.maturity             -0.06855    0.01846  -3.714 0.000204 ***

This tells us that as age at maturity increases,log odds of extinction decrease. Also, as change in minimum temperature experienced between 2-3 years ago increases (towards positive), log odds of extinction decreases.

###Trophic Level

Hypothesis:

Best Colonization Models

| | Variable | Coefficient | Pr(>|z|) | AICc | Rank for Full Models | Coefficient for Full Models
|---------|-------------|----------------|------------------|--------------|------------|
| 1 | seas_sst_temp_change_abs |	0.17121179 |	0.0005599279 |	4525.486 |	12 | 0.10 |
| 2 | max_sbt_temp_change_abs_lag3 |	0.19678089 |	0.0013246068 |	4526.919 | 4 | 0.17 |
| *3 | max_sbt_temp_change_abs |	0.18509626 |	0.0021952362 |	4527.742 | 1 | 0.24
| 4 | max_sst_temp_change_abs_lag1 |	-0.25709293 |	0.0033255361 |	4527.955 | 2 | -0.24
| 5 | seas_sbt_temp_change_abs |	0.15728890 |	0.0034775899 |	4528.408 |	5 | 0.14
| 6 | max_sbt_temp_change_abs_lag1 |	0.16544380 |	0.0096094080 |	4530.271 |	11 | 0.14
| 7 | seas_sbt_temp_change_abs_lag1 |	0.13431316 |	0.0183946976 |	4531.309 |	10 | 0.13
| 8 | seas_sbt_temp_change_abs_lag4 |	0.13162107 |	0.0232351577 |	4531.668 |	8 | 0.14
| 9 | min_sbt_temp_change_abs |	0.17182671 |	0.0196874934 |	4531.795 | not in top 15 ||
| 10 | max_sst_temp_change_abs_lag8 |	0.16963749 |	0.0255186622 |	4532.091 |	6 | 0.19 |
| 11 | min_sbt_temp_change_abs_lag5 |	-0.20316841 |	0.0397350249 |	4532.176 |	3 | -0.24 |
| 12 | max_sst_temp_change_abs_lag6 |	-0.18049426 |	0.0381593586 |	4532.251 | not in top 15 ||
| 13 | seas_sst_temp_change_lag5 |	-0.07645147 |	0.0360758538 |	4532.280 |	not in top 15 ||
| 14 | max_sst_temp_change_abs |	0.20292370 |	0.0327597085 |	4532.339 |	14 | 0.17|
| 15 | min_sst_temp_change_abs |	0.13663573 |	0.0333662358 |	4532.452 |	not in top 15||
**Used for models below*
Of top 15 models, 11/15 shared. 

Looked at:
* col ~ temp
* col ~ temp + trophic level
* col ~ temp * trophic level
* col ~ trophic level

Best model includes temperature and trophic level, including interaction ((Δ = 6.027 vs w/o interaction):
  variable                    coefficient Pr(>|z|) AICc  
  max_sbt_temp_change_abs     0.93859   0.000332 4509.615
  tl                         -0.14904   0.226275    
  max_sbt_temp_change_abs:tl -0.20651   0.003700 
  
Best Extinction Models

| | Variable | Coefficient | Pr(>|z|) | AICc | Rank for Full Models | Coefficient for Full Models
|---------|-------------|----------------|------------------|--------------|------------|
|*1 | min_sst_temp_change_lag2 |	-0.12605189 |	0.008882239 |	4205.455 |2	|-0.09|
|2 | seas_sbt_temp_change_abs	| 0.13647286 | 	0.014522389	| 4206.321 |4 |0.12 | 	
|3 | seas_sbt_temp_change_lag9 |	0.11068561 |	0.016010438 |	4206.530 |not in top 15 | |	
|4 | seas_sbt_temp_lag9 |	0.10832214 |	0.014587647 |	4206.647 | 6 | 0.09 |	
|5 | min_sst_temp_change_abs_lag6 |	0.17395683 |	0.014698403 |	4206.661 | not in top 15	||
|6 | min_sbt_temp_change_abs_lag5 |	-0.25022141 |	0.023335227 |	4206.734 | not in top 15 | |	
|7 | max_sbt_temp_change_abs |	0.15091446 |	0.018910307 |	4206.845 | not in top 15 | |	
|8 |min_sst_temp_change_abs_lag7 |	0.16683645 |	0.017394540 |	4206.863 |15	|0.11|
|9 | min_sst_temp_change_abs_lag5 |	-0.18898669 |	0.027881516 |	4207.251 |not in top 15	||
|10| min_sbt_temp_change_lag7 |	-0.10660642 |	0.056611513 |	4208.757 | 3| -0.12|	
|11|  seas_sst_temp_lag2 |	0.07887305 |	0.050759977 |	4208.927 |12	|0.07|
|12 | seas_sst_temp_change_abs_lag5 |	-0.11074422 |	0.077570136 |	4208.999 |not in top 15	||
|13 |  min_sst_temp_change |	0.08871614 |	0.078886342 |	4209.073 |not in top 15	||
|14 |  seas_sst_temp |	-0.08790929 |	0.067876817 |	4209.092 |not in top 15	||
|15 | max_sbt_temp_change_lag9 |	0.09267424 |	0.081161794 |	4209.182 |not in top 15 | |	
**Used for models below*

Of top 15, only 6 are shared. 

Looked at:
* now_Ext ~ temp
* now_Ext ~ temp + trophic level
* now_Ext ~ temp * trophic level
* now_Ext ~ trophic level

Best model includes both trophic level and temperature, but not an interaction ($/Delta$ = 1.914 vs. w/ interaction). 

                         Estimate Std. Error z value Pr(>|z|)      AICc
(Intercept)              -2.36179    0.44387  -5.321 1.03e-07 *** 4192.295
min_sst_temp_change_lag2 -0.12635    0.04823  -2.620   0.0088 ** 
tl                       -0.39633    0.10089  -3.928 8.55e-05 ***

This tells us that as trophic level increases, log odds of extinction decreases. As change in minimum temperature experienced between 2-3 years ago towards positive increases, the log odds of extinction decreases. 
