---
title: "Colonization and Extinction Final Models with Traits Included"
output: html_notebook
---

This is a cleaned version of dredge_trait_models
Changes

* species and region as random effects (not year)
* scaling temperature values
* setting nACQ to 0, see why [here](https://stats.stackexchange.com/questions/77313/why-cant-i-match-glmer-family-binomial-output-with-manual-implementation-of-g). 

```{r setup}
library(MuMIn)
library(lme4)
library(data.table)
library(ggplot2)

load("spp_master_ztemp.traits.RData")
```

Running, and ranking colonization models without traits

```{r scaling temperature values}
#extract just temperature variables to scale
spp_master_ztemp_varsonly <- spp_master_ztemp.traits[,c(1,45:292, 301, 308, 310, 312, 313)]

#names of temp variables, so I can add "_scaled" suffix
temptraitvars <- colnames(spp_master_ztemp.traits[,c(45:292, 301, 308, 310, 312, 313)])

#scale all variables, make new columns, do this by region
spp_master_ztemp_varsonly[, paste0(temptraitvars, "_scaled") := lapply(.SD, function(x) as.vector(scale(x))), .SDcols = temptraitvars]

#names of scaled columns
scaled_cols <- grep("*_scaled", names(spp_master_ztemp_varsonly), value = T)

#merge new columns back with spp_master_ztemp
spp_master_ztemp.traits_scaled <- cbind(spp_master_ztemp.traits, spp_master_ztemp_varsonly[,scaled_cols, with = FALSE])

save(spp_master_ztemp.traits_scaled, file = "spp_master_ztemp.traits_scaled.RData")

```

I will be looking at:

*trophic level
*maximum age
*age at maturity
*maximum body size

```{r missing values}
summary(spp_master_ztemp.traits_scaled$age.max) #702 missing
summary(spp_master_ztemp.traits_scaled$age.maturity) #551 missing
summary(spp_master_ztemp.traits_scaled$tl) #none missing
summary(spp_master_ztemp.traits_scaled$length.max) #none missing
summary(spp_master_ztemp.traits_scaled$length.infinity) #443 missing, this is highly correlated (0.82) with length max, so we'll just use length.max

cor(spp_master_ztemp.traits_scaled[,.(age.max_scaled, age.maturity_scaled, tl_scaled, length.max_scaled, length.infinity_scaled)], use = "complete.obs")

pairs(spp_master_ztemp.traits_scaled[,.(age.max_scaled, age.maturity_scaled, tl_scaled, length.max_scaled, length.infinity_scaled)])

```

Prep for dredge
```{r prep for model runs}

#temp scaled vars only
tempvars <- names(spp_master_ztemp.traits_scaled[,314:561])
othervars_col <- c("reg", "year", "col", "spp", "age.max_scaled", "age.maturity_scaled", "tl_scaled", "length.max_scaled")
colstosaveforcol <- c(othervars_col, tempvars)
#Colonization table
colonization.reduced <- spp_master_ztemp.traits_scaled[,colstosaveforcol, with = F] #reduces to columns we need

#Extinction table

othervars_ext <- c("reg", "year", "now_ext", "spp", "age.max_scaled", "age.maturity_scaled", "tl_scaled", "length.max_scaled")
colstosaveforext <- c(othervars_ext, tempvars)
#Colonization table
extinction.reduced <- spp_master_ztemp.traits_scaled[,colstosaveforext, with = F] #reduces to columns we need



```


Making all models for colonization
```{r all models colonization}

#playing with year as factor
#can't get dredge to run above, so trying here with fewer variables
colonization.reduced[,year := as.factor(year)]

#for now, going with two traits and one temp variable
#going to make loop to make all models I need to look at with single variables
all.temp.variables <- colnames(colonization.reduced[,9:256])
all.temp.variables.0 <- c(0, all.temp.variables) #add 0 as option so models can just include traits
all.temp.variables.0 <- all.temp.variables.0[!grepl("*mean*", all.temp.variables.0)]#get rid of MEAN temp variables
all.trait.variables <- c(0, "tl_scaled", "age.max_scaled","age.maturity_scaled", "length.max_scaled",
                         "tl_scaled + age.max_scaled", "age.max_scaled + tl_scaled",
                         "tl_scaled + age.maturity_scaled", "age.maturity_scaled + tl_scaled",
                         "tl_scaled + length.max_scaled", "length.max_scaled + tl_scaled",
                         "age.max_scaled + age.maturity_scaled", "age.maturity_scaled + age.max_scaled",
                         "age.max_scaled + length.max_scaled", "length.max_scaled + age.max_scaled",
                         "age.maturity_scaled + length.max_scaled", "length.max_scaled + age.maturity_scaled")

sign <- c("+", "*")

full_var_set <- as.data.table(expand.grid(all.temp.variables.0, sign, all.trait.variables))

colonization_model_comparison_traits <- as.data.table(matrix(nrow = nrow(full_var_set))) 
colonization_model_comparison_traits[, temp_variable:=as.factor(V1)][, sign:=as.factor(V1)][, trait_variable:=as.factor(V1)][, AICc:=as.numeric(V1)][, converge:= as.logical(V1)]
colonization_model_comparison_traits[, V1 := NULL]

for (i in 1:nrow(full_var_set)){
 
  temp <- full_var_set$Var1[i]
  sign <- full_var_set$Var2[i]
  trait <- full_var_set$Var3[i]
  
  formula <- as.formula(paste("col ~", temp, sign, trait, "+ (1|reg) + (1|spp)"))
  
  mod <- glmer(formula, family = binomial, data = colonization.reduced, nAGQ = 0)
  
  colonization_model_comparison_traits[i,temp_variable := full_var_set$Var1[i]]
  colonization_model_comparison_traits[i,sign := full_var_set$Var2[i]]
  colonization_model_comparison_traits[i,trait_variable := full_var_set$Var3[i]]
  colonization_model_comparison_traits[i,AICc := AICc(mod)]
  
  warning <- mod@optinfo$conv$lme4$messages[2]
  converges <- ifelse(is.null(warning) == TRUE, T, F)
  
  colonization_model_comparison_traits[i, converge := converges]
  
  print(paste(i, nrow(colonization_model_comparison_traits), sep = "/"))
    
}


save(colonization_model_comparison_traits, file = "colonization_model_comparison_traits.RData")


```

Making all models for extinction
```{r all models extinction}

#for now, going with two traits and one temp variable
#going to make loop to make all models I need to look at with single variables
#don't want mean

extinction_model_comparison_traits <- as.data.table(matrix(nrow = nrow(full_var_set))) 
extinction_model_comparison_traits[, temp_variable:=as.factor(V1)][, sign:=as.factor(V1)][, trait_variable:=as.factor(V1)][, AICc:=as.numeric(V1)][, converge:= as.logical(V1)]
extinction_model_comparison_traits[, V1 := NULL]

for (i in 1:nrow(full_var_set)){
 
  temp <- full_var_set$Var1[i]
  sign <- full_var_set$Var2[i]
  trait <- full_var_set$Var3[i]
  
  formula <- as.formula(paste("now_ext ~", temp, sign, trait, "+ (1|reg) + (1|spp)"))
  
  mod <- glmer(formula, family = binomial, data = extinction.reduced, nAGQ = 0)
  
  extinction_model_comparison_traits[i,temp_variable := full_var_set$Var1[i]]
  extinction_model_comparison_traits[i,sign := full_var_set$Var2[i]]
  extinction_model_comparison_traits[i,trait_variable := full_var_set$Var3[i]]
  extinction_model_comparison_traits[i,AICc := AICc(mod)]
  
  warning <- mod@optinfo$conv$lme4$messages[2]
  converges <- ifelse(is.null(warning) == TRUE, T, F)
  
  extinction_model_comparison_traits[i, converge := converges]
  
  print(paste(i, nrow(extinction_model_comparison_traits), sep = "/"))
    
}

save(extinction_model_comparison_traits, file = "extinction_model_comparison_traits.RData")

```

Next step is to look at RVI (relative variable importance)

COLONIZATION

```{r RVI colonization traits}
#only raw temp has lag of 10 years, which means variable importance comparisons are unbalanced so I will get rid of these rows
colonization_model_comparisons_traits.lag9max <- colonization_model_comparison_traits[!grepl("lag10", temp_variable), ]

#add ∆AIC to tables
min_col_AICc_trait <- min(colonization_model_comparisons_traits.lag9max[, AICc])
colonization_model_comparisons_traits.lag9max[,"deltaAICc" := (AICc - min_col_AICc_trait)]

#add relative likelihood exp( -0.5 * ∆AIC score for that model)
colonization_model_comparisons_traits.lag9max[,"rel_likelihood" := exp((-0.5 * deltaAICc))]
#sum relative likelihoods across all models
colonization_model_comparisons_traits.lag9max.likelihoodsum <- sum(colonization_model_comparisons_traits.lag9max[, rel_likelihood])
#Akaike weight for a model is this rel_likelihood devided by the sum of these values across all models
colonization_model_comparisons_traits.lag9max[,"akaike_weight" := rel_likelihood/colonization_model_comparisons_traits.lag9max.likelihoodsum]

#I want to look at relative importance FOR 
#bottom/surface (bottom = includes sbt, grepl("sbt", variable))
col_bottom_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("sbt", temp_variable), ]$akaike_weight)
col_surface_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("sst", temp_variable), ]$akaike_weight)
  col_bottom_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("sbt", temp_variable), ]) #3060
  col_surface_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("sst", temp_variable), ]) #3060

#lag/not lagged (grepl("lag", variable))
col_lag_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("lag", temp_variable), ]$akaike_weight)
col_nolag_importance <- sum(colonization_model_comparisons_traits.lag9max[!grep("lag", temp_variable), ]$akaike_weight)
  col_lag_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("lag", temp_variable), ]) #5508 models include a lag
  col_nolag_importance.n <- count(colonization_model_comparisons_traits.lag9max[!grepl("lag", temp_variable), ]) #646 don't include a lag (no lag OR trait only)


  #34 models have NO temp variable
  
  
#absolute (grepl("abs", variable))
col_abs_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("abs", temp_variable), ]$akaike_weight)
  col_abs_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("abs", temp_variable), ])#2040
  
#raw (!grepl("change"))
col_raw_importance <- sum(colonization_model_comparisons_traits.lag9max[!grep("change", temp_variable), ][temp_variable != 0,]$akaike_weight)
  col_raw_importance.n <- count(colonization_model_comparisons_traits.lag9max[!grepl("change", temp_variable), ][temp_variable != 0]) #2040

#change (grepl("change", temp_variable), (!grepl("abs")))
col_change_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("change", temp_variable), ][!grep("abs", temp_variable),]$akaike_weight)
  col_change_importance.n <- count(colonization_model_comparisons_traits.lag9max[grep("change", temp_variable), ][!grep("abs", temp_variable),])#2040


#type of temp temp_variable (max, min, seas)
col_max_temp_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("max", temp_variable), ]$akaike_weight)
col_min_temp_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("min", temp_variable), ]$akaike_weight)
col_seas_temp_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("seas", temp_variable), ]$akaike_weight)
  col_max_temp_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("max", temp_variable), ]) #2040
  col_min_temp_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("min", temp_variable), ]) #2040
  col_seas_temp_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("seas", temp_variable), ]) #2040
  

#traits included
col_age.max_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("age.max", trait_variable), ]$akaike_weight)
col_age.maturity_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("age.maturity", trait_variable), ]$akaike_weight)
col_tl_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("tl", trait_variable), ]$akaike_weight)
col_length.max_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("length", trait_variable),]$akaike_weight)
  col_age.max_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("age.max", trait_variable), ]) #2534
  col_age.maturity_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("age.maturity", trait_variable), ]) #2534
  col_tl_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("tl", trait_variable), ]) #2534
  col_length.max_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("length", trait_variable),]) #2534


#interaction between trait and temperature
col_interaction_temp_trait_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("\\*", sign), ]$akaike_weight)
  col_interaction_temp_trait_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("\\*", sign), ]) #3077


#save this table
save(colonization_model_comparisons_traits.lag9max, file = "colonization_models_traits_with_likelihoods.Rdata")
```
Best model stats?
```{r best model stats for models with traits}
mod_col_trait <- glmer(col ~ max_sst_temp_change_abs_lag1_scaled * age.maturity_scaled + age.max_scaled + (1|spp) + (1|reg), data = colonization.reduced, nAGQ = 0, family = binomial)

summary(mod_col_trait)

#R^2 (coefficient of determination based on likelihood ratio test)
#adj.r.squared: Nagelkerke's modified R^2 statistic
library(MuMIn)
r.squaredLR(mod_col_trait)

```

```{r put RVIs for colonization into data table}
col_ext <- c("col","col","col","col","col", "col","col", "col", "col","col", "col","col", "col", "col","col")
type <- c("depth", "depth", "lag", "lag", "Change?", "Change?", "Change?", "Temp", "Temp", "Temp", "Trait", "Trait", "Trait","Trait", "Interaction")
variable <- c("bottom", "surface", "no_lag", "lagged", "absolute_value_change", "change", "raw", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "length.max", "tempxtrait")
value <- c(col_bottom_importance, col_surface_importance, col_nolag_importance, col_lag_importance, col_abs_importance, col_change_importance, col_raw_importance, col_min_temp_importance, col_max_temp_importance, col_seas_temp_importance, col_age.max_importance, col_age.maturity_importance, col_tl_importance, col_length.max_importance, col_interaction_temp_trait_importance)

RVI_col_traits <- data.table(col_ext, type, variable, value)
```


EXTINCTION

```{r RVI extinction traits}
#only raw temp has lag of 10 years, which means variable importance comparisons are unbalanced so I will get rid of these rows
extinction_model_comparisons_traits.lag9max <- extinction_model_comparison_traits[!grepl("lag10", temp_variable), ]


#add ∆AIC to tables
min_ext_AICc_trait <- min(extinction_model_comparisons_traits.lag9max[, AICc])
extinction_model_comparisons_traits.lag9max[,"deltaAICc" := (AICc - min_ext_AICc_trait)]

#add relative likelihood exp( -0.5 * ∆AIC score for that model)
extinction_model_comparisons_traits.lag9max[,"rel_likelihood" := exp((-0.5 * deltaAICc))]
#sum relative likelihoods across all models
extinction_model_comparisons_traits.lag9max.likelihoodsum <- sum(extinction_model_comparisons_traits.lag9max[, rel_likelihood])
#Akaike weight for a model is this rel_likelihood devided by the sum of these values across all models
extinction_model_comparisons_traits.lag9max[,"akaike_weight" := rel_likelihood/extinction_model_comparisons_traits.lag9max.likelihoodsum]

#I want to look at relative importance FOR 
#bottom/surface (bottom = includes sbt, grepl("sbt", variable))
ext_bottom_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("sbt", temp_variable), ]$akaike_weight)
  ext_bottom_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("sbt", temp_variable), ]) #3060
ext_surface_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("sst", temp_variable), ]$akaike_weight)
  ext_surface_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("sst", temp_variable), ]) #3060
  
#lag/not lagged (grepl("lag", variable))
ext_lag_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("lag", temp_variable), ]$akaike_weight)
  ext_lag_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("lag", temp_variable),]) #5508
ext_nolag_importance <- sum(extinction_model_comparisons_traits.lag9max[!grep("lag", temp_variable), ]$akaike_weight)
  ext_nolag_importance.n <- count(extinction_model_comparisons_traits.lag9max[!grepl("lag", temp_variable), ]) #646

  #absolute (grepl("abs", variable))
ext_abs_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("abs", temp_variable), ]$akaike_weight)
  ext_abs_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("abs", temp_variable), ]) #2040

  #raw (!grepl("change"))
ext_raw_importance <- sum(extinction_model_comparisons_traits.lag9max[!grep("change", temp_variable), ][temp_variable != 0,]$akaike_weight)
  ext_raw_importance.n <- count(extinction_model_comparisons_traits.lag9max[!grepl("change", temp_variable), ][temp_variable != 0,]) #2040

#change (grepl("change", temp_variable), (!grepl("abs")))
ext_change_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("change", temp_variable), ][!grep("abs", temp_variable),]$akaike_weight)
  ext_change_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("change", temp_variable), ][!grepl("abs", temp_variable),])#2040

#type of temp temp_variable (max, min, seas)
ext_max_temp_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("max", temp_variable), ]$akaike_weight)
  ext_max_temp_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("max", temp_variable), ]) #2040
ext_min_temp_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("min", temp_variable), ]$akaike_weight)
  ext_min_temp_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("min", temp_variable), ]) #2040
ext_seas_temp_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("seas", temp_variable), ]$akaike_weight)
  ext_seas_temp_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("seas", temp_variable), ]) #2040

#traits included
ext_age.max_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("age.max", trait_variable), ]$akaike_weight)
ext_age.maturity_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("age.maturity", trait_variable), ]$akaike_weight)
ext_tl_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("tl", trait_variable), ]$akaike_weight)
ext_length.max_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("length", trait_variable),]$akaike_weight)

  ext_age.max_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("age.max", trait_variable), ]) #2534
  ext_age.maturity_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("age.maturity", trait_variable), ]) #2534
  ext_tl_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("tl", trait_variable), ]) #2534
  ext_length.max_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("length", trait_variable),]) #2534

#interaction between trait and temperature
ext_interaction_temp_trait_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("\\*", sign), ]$akaike_weight)
  ext_interaction_temp_trait_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("\\*", sign), ]) #3077


#save this table
save(extinction_model_comparisons_traits.lag9max, file = "extinction_models_traits_with_likelihoods.Rdata")
head(extinction_model_comparisons_traits.lag9max)
```

```{r put RVIs for extinction into data table}
col_ext <- c("ext","ext","ext","ext","ext", "ext","ext", "ext", "ext","ext", "ext","ext", "ext", "ext","ext")
type <- c("depth", "depth", "lag", "lag", "Change?", "Change?", "Change?", "Temp", "Temp", "Temp", "Trait", "Trait", "Trait","Trait", "Interaction")
variable <- c("bottom", "surface", "no_lag", "lagged", "absolute_value_change", "change", "raw", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "length.max", "tempxtrait")
value <- c(ext_bottom_importance, ext_surface_importance, ext_nolag_importance, ext_lag_importance, ext_abs_importance, ext_change_importance, ext_raw_importance, ext_min_temp_importance, ext_max_temp_importance, ext_seas_temp_importance, ext_age.max_importance, ext_age.maturity_importance, ext_tl_importance, ext_length.max_importance, ext_interaction_temp_trait_importance)

RVI_ext_traits <- data.table(col_ext, type, variable, value)
```

Best model stats?
```{r best model stats for extinction models with traits}
mod_ext_trait <- glmer(now_ext ~max_sst_temp_scaled * age.maturity_scaled + age.max_scaled + (1|spp) + (1|reg), data = extinction.reduced, nAGQ = 0, family = binomial)

summary(mod_ext_trait)

#R^2 (coefficient of determination based on likelihood ratio test)
#adj.r.squared: Nagelkerke's modified R^2 statistic
r.squaredLR(mod_ext_trait)


```

Merge RVI tables for colonization and extinction with traits, and then graph
```{r merge RVI trait tables and plot}
RVI_trait_table <- as.data.table(rbind(RVI_col_traits, RVI_ext_traits)) #bind RVI for colonizations and extinctions

RVI_trait_table.r <- RVI_trait_table[!grep("surface", variable)][!grep("lagged", variable)]

#change order of factor levels
RVI_trait_table.r[, variable := factor(variable, levels = c("bottom", "no_lag", "raw", "change", "absolute_value_change", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "length.max", "tempxtrait"))]

save(RVI_trait_table, file = "RVI_trait_table.Rdata")

ggplot(data = RVI_trait_table.r, aes(x=variable, y = value, fill = col_ext)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_discrete(breaks = c("bottom", "no_lag", "raw", "change", "absolute_value_change", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "length.max", "tempxtrait"), labels = c("Bottom\nTemp", "No\nLag", "Raw\nTemp", "Change in \nTemp", "| Change in\n Temp |", "Min\nTemp", "Max\nTemp", "Seas", "Age\nMax", "Age\nMaturity", "Trophic\nLevel", "Max\nLength" , "Temp*\nTrait")) +
  labs(x = "Temperature and Trait Variables", y = "Relative Variable Importance")  +
  theme_bw()

```

Another way to visualize
```{r visualizing RVI with traits}
#put levels into correct order for viewing
RVI_trait_table[
  , variable := factor(variable, levels = c("bottom", "surface", "absolute_value_change", "change", "raw", "no_lag", "lagged", "Min", "Max", "Seas", "age.max", "age.maturity", "tl","length.max", "tempxtrait"))][
    ,type := factor(type, levels = c("Change?", "depth", "lag", "Temp", "Trait", "Interaction"))]
#make types factors as well

#colonization
RVI_col_trait_plot <- ggplot(data = RVI_trait_table[col_ext == "col" & (type == "Trait" | type == "Interaction")], aes(x=variable, y = value)) +
  geom_bar(stat= "identity") +
  scale_x_discrete(breaks = c("age.max", "age.maturity", "tl", "length.max", "tempxtrait"), labels = c("Maximum\nAge", "Age of\nMaturity", "Trophic\nLevel","Max\nLength", "Temp Trait\nInteraction")) +
  labs(x = "Trait Variables", y = "Relative Variable Importance") +
  theme_classic() +
  theme(text = element_text(size = 20))
#extinction
RVI_ext_trait_plot <- ggplot(data = RVI_trait_table[col_ext == "ext" & (type == "Trait" | type == "Interaction")], aes(x=variable, y = value)) +
  geom_bar(stat= "identity") +
  scale_x_discrete(breaks = c("age.max", "age.maturity", "tl","length.max",  "tempxtrait"), labels = c("Maximum\nAge", "Age of\nMaturity", "Trophic\nLevel","Max\nLength", "Temp Trait\nInteraction")) +
  labs(x = "Trait Variables", y = "Relative Variable Importance") +
  theme_classic() +
  theme(text = element_text(size = 20))

RVI_col_trait_plot
RVI_ext_trait_plot

save(RVI_trait_table, file = "RVI_trait_table.Rdata")
```
Let's see how the RVI importance varies by looking at specific lag values for these trait models
```{r make a plot here for variability across lag values for trait models}
#first, table for lags
lags_v_RVI_trait <- data.table(col_lag = c(0:9), col_RVI = 0, ext_RVI = 0)
#colonization
lags_v_RVI_trait[1,2] <- 1-sum(colonization_model_comparisons_traits.lag9max[grep("lag", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[2,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag1", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[3,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag2", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[4,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag3", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[5,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag4", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[6,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag5", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[7,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag6", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[8,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag7", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[9,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag8", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[10,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag9", temp_variable), ]$akaike_weight)
#extinction
lags_v_RVI_trait[1,3] <- 1-sum(extinction_model_comparisons_traits.lag9max[grep("lag", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[2,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag1", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[3,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag2", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[4,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag3", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[5,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag4", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[6,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag5", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[7,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag6", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[8,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag7", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[9,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag8", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[10,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag9", temp_variable), ]$akaike_weight)

RVI_lags_traits <- ggplot(data = lags_v_RVI_trait, aes(x = col_lag)) +
  geom_line(aes(y = col_RVI), col = "blue") +
  geom_line(aes(y = ext_RVI), col = "red", linetype = "dashed") +
  labs(x = "Lag (years)", y = "Relative Variable Importance") +
  theme_classic() +
  theme(text = element_text(size = 22, color = "black")) +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = 0:1)

save(RVI_col_trait_plot, RVI_ext_trait_plot, RVI_lags_traits, file = "RVI_plots_traits.Rdata")

setEPS()
postscript("RVI_col_trait_plot.eps")
RVI_col_trait_plot
dev.off()

setEPS()
postscript("RVI_ext_trait_plot.eps")
RVI_ext_trait_plot
dev.off()

setEPS()
postscript("RVI_lags_trait_plot.eps")
RVI_lags_traits
dev.off()
```

What do interactions look like?

##Colonization with Traits
Best model: max_sst_temp_change_abs_lag1_scaled * age.maturity_scaled + age.max_scaled
```{r col with traits}
library(ggeffects)

#model
mod_col_trait <- glmer(col ~ max_sst_temp_change_abs_lag1_scaled * age.maturity_scaled + age.max_scaled + (1|reg) + (1|spp), family = "binomial", data = colonization.reduced, nAGQ = 0)
min(colonization.reduced$max_sst_temp_change_abs_lag1_scaled)
max(colonization.reduced$max_sst_temp_change_abs_lag1_scaled)
#plot marginal effects of interactions
values <- seq(-1.2, 5.5, by = 0.01)
dat <- ggpredict(mod_col_trait, c("max_sst_temp_change_abs_lag1_scaled[all]", "age.maturity_scaled [-0.78, 1.53, 6.67, 11.80, 24.63]"))

summary(colonization.reduced$max_sst_temp_change_abs_lag1_scaled)

ggplot(dat, aes(x=x, y = predicted, colour = group)) +
  geom_point(size = 1) +
  labs(colour="Age at Maturity Values", x = "Absolute change in maximum surface temperature\nexperienced a year previously (˚C centered and scaled)", y = "Probability of Colonization",title = "Predicted Probabilities of Colonization") +
  scale_y_continuous(breaks=c(0, 0.5, 1)) +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  theme(text=element_text(size = 14))

ggsave(path = "plots/marginal_effects", file = "predicted_probs_col_maxage.jpg")

mod <- lm(spp_master_ztemp.traits_scaled$age.maturity_scaled~spp_master_ztemp.traits_scaled$age.maturity)
summary(mod)

df <- data.frame(unscaled = c(1,10,30,50,100))
df$scaled <- mod$coefficients[1]+mod$coefficients[2]*df$unscaled

```
##Extinction with Traits
Best model: max_sst_temp_scaled * age.maturity_scaled + age.max_scaled
```{r ext with traits}
library(ggeffects)

#model
mod_ext_trait <- glmer(now_ext ~ max_sst_temp_scaled * age.maturity_scaled + age.max_scaled + (1|reg) + (1|spp), family = "binomial", data = extinction.reduced, nAGQ = 0)

#plot marginal effects of interactions
dat <- ggpredict(mod_ext_trait, c("max_sst_temp_scaled [all]", "age.maturity_scaled [-0.78, 1.53, 6.67, 11.80, 24.63]"))

ggplot(dat, aes(x=x, y = predicted, colour = group)) +
  geom_point(size = 1) +
  labs(colour="Age at Maturity Values", x = "Maximum surface temperature in the\npast year (˚C centered and scaled)", y = "Probability of Extinction",title = "Predicted Probabilities of Extinction") +
  scale_y_continuous(breaks=c(0, 0.5, 1)) +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  theme(text=element_text(size = 14))

```

How well do these models actually predict the # of colonizations and the # of extinctions?

```{r predicted vs. observed colonizations scaled}
library(tidyr)
library(plyr)
#best colonization trait models (all delta less than 2)
  #max_sst_temp_change_abs_lag1_scaled * age.maturity_scaled + age.max_scaled


bestcolmod_scaled_traits <- glmer(col ~ max_sst_temp_change_abs_lag1_scaled * age.maturity_scaled + age.max_scaled + (1|reg) + (1|spp), family = binomial, data = spp_master_ztemp.traits_scaled, nAGQ = 0)

interceptcolmod_scaled_traits <- glmer(col ~ (1|reg) + (1|spp), family = binomial, data = spp_master_ztemp.traits_scaled, nAGQ = 0)

summary(bestcolmod_scaled_traits)

#I wanted predicted # of colonizations in each region per region/year
#year region observed colonizations predicted colonizations

#plotting predicted versus modeled
#unique combinations of spp, region, and max_sbt_temp_change_abs_scaled found in dataset
model_imputs_col_traits <- spp_master_ztemp.traits_scaled[,.(spp, reg, year, col, max_sst_temp_change_abs_lag1_scaled, age.maturity_scaled, age.max_scaled)]

model_imputs_col_traits <- na.omit(model_imputs_col_traits) #get rid of any rows without data for all variables

t2_col <- predict(bestcolmod_scaled_traits, newdata = model_imputs_col_traits, type = "response", re.form = NULL) #setting re.form to NULL includes all random effects
t2_col_null <- predict(interceptcolmod_scaled_traits, newdata = model_imputs_col_traits, type = "response", re.form = NULL)

model_imputs_col_traits$pred <- t2_col
model_imputs_col_traits$pred_null <- t2_col_null

#group by region and year, sum over "col" = observations, and "pred" = predict colonizations by summing probabilities
col_model_comparisons_traits <- model_imputs_col_traits[, .(col_predicted = sum(pred), col_predicted_null = sum(pred_null), col_observed = sum(col)), by=.(year,reg)]


ggplot(aes(x=year), data = col_model_comparisons_traits) +
  geom_point(aes(y = col_observed), data = col_model_comparisons_traits, col = "red") +
  geom_point(data = col_model_comparisons_traits, aes(y = col_predicted), col = "black") +
  facet_wrap(~reg) +
  theme_classic()

```

And extinctions?

```{r predicted vs. observed extinctions scaled}
#best extinction trait models (all delta less than 2)
 #max_sst_temp_scaled    * age.maturity_scaled + age.max_scaled 0.000000


bestextmod_scaled_traits <- glmer(now_ext ~ max_sst_temp_scaled * age.maturity_scaled + age.max_scaled + (1|reg) + (1|spp), family = binomial, data = spp_master_ztemp.traits_scaled, nAGQ = 0)

interceptextmod_scaled_traits <- glmer(now_ext ~ (1|reg) + (1|spp), family = binomial, data = spp_master_ztemp.traits_scaled, nAGQ = 0)

summary(bestextmod_scaled_traits)

#I wanted predicted # of extinctions in each region per region/year
#year region observed extinctions predicted extinctions

#plotting predicted versus modeled
#unique combinations of spp, region, and max_sst_temp_scaled found in dataset
model_imputs_ext_traits <- spp_master_ztemp.traits_scaled[,.(spp, reg, year, now_ext, max_sst_temp_scaled, age.maturity_scaled, age.max_scaled)]

model_imputs_ext_traits <- na.omit(model_imputs_ext_traits)

t2_ext <- predict(bestextmod_scaled_traits, newdata = model_imputs_ext_traits, type = "response", re.form = NULL) #setting re.form to NULL includes all random effects
t2_ext_null <- predict(interceptextmod_scaled_traits, newdata = model_imputs_ext_traits, type = "response", re.form = NULL)

model_imputs_ext_traits$pred <- t2_ext
model_imputs_ext_traits$pred_null <- t2_ext_null

#group by region and year, sum over "now_ext" = observations, and "pred" = predict extinctions by summing probabilities
ext_model_comparisons_traits <- model_imputs_ext_traits[, .(ext_predicted = sum(pred), ext_predicted_null = sum(pred_null),  ext_observed = sum(now_ext)), by=.(year,reg)]


ggplot(aes(x=year), data = ext_model_comparisons_traits) +
  geom_point(aes(y = ext_observed), data = ext_model_comparisons_traits, col = "red") +
  geom_point(data = ext_model_comparisons_traits, aes(y = ext_predicted), col = "black") +
  facet_wrap(~reg) +
  theme_classic()

```

What about richness accumulation??

Start with colonization models
```{r richness accumulation for trait models}
col_model_comparisons_traits[,col_obs_sum := cumsum(col_observed), by = reg]
col_model_comparisons_traits[,col_pred_sum := cumsum(col_predicted), by = reg]
col_model_comparisons_traits[,col_pred_sum_null := cumsum(col_predicted_null), by = reg]

ggplot(aes(x=year), data = col_model_comparisons_traits) +
  geom_point(aes(y = col_obs_sum), data = col_model_comparisons_traits, col = "red") +
  geom_point(data = col_model_comparisons_traits, aes(y = col_pred_sum), col = "black", size = 1) +
  labs(x="Year", y = "Cumulative sum of colonizations") +
  facet_wrap(~reg) +
  theme_classic()


```


Now, accumulation of extinctions
```{r plot extinction with cumulative sum of extinctions}

ext_model_comparisons_traits[,ext_obs_sum := cumsum(ext_observed), by = reg]
ext_model_comparisons_traits[,ext_pred_sum := cumsum(ext_predicted), by = reg]
ext_model_comparisons_traits[,ext_pred_sum_null := cumsum(ext_predicted_null), by = reg]

ggplot(aes(x=year), data = ext_model_comparisons_traits) +
  geom_point(aes(y = ext_obs_sum), data = ext_model_comparisons_traits, col = "red") +
  geom_point(data = ext_model_comparisons_traits, aes(y = ext_pred_sum), col = "black", size = 1) +
  labs(x="Year", y = "Cumulative sum of extinctions") +
  facet_wrap(~reg) +
  theme_classic()


```



Cumulative species richness change
```{r cumulative species richness change}
#from colonizations
cumulative_richness_traits <- col_model_comparisons_traits[,.(reg, year, col_obs_sum, col_pred_sum, col_pred_sum_null)]
ext_obs_sum <- ext_model_comparisons_traits[,.(ext_obs_sum)]
ext_pred_sum <- ext_model_comparisons_traits[,.(ext_pred_sum)]
ext_pred_sum_null <- ext_model_comparisons_traits[,.(ext_pred_sum_null)]
cumulative_richness_traits[,ext_obs_sum := ext_obs_sum]
cumulative_richness_traits[,ext_pred_sum := ext_pred_sum]
cumulative_richness_traits[,ext_pred_sum_null := ext_pred_sum_null]

cumulative_richness_traits[, Observed := col_obs_sum-ext_obs_sum][, Predicted := col_pred_sum-ext_pred_sum][, Predicted_null := col_pred_sum_null-ext_pred_sum_null]

cumulative_richness_traits.r <- cumulative_richness_traits[,.(reg, year, Observed, Predicted, Predicted_null)]

#wide to long
cumulative_richness_traits.r <- melt(cumulative_richness_traits.r, id.vars = c("reg", "year"), measure.vars = c("Observed", "Predicted", "Predicted_null"), variable.name = "Richness", value.name = "value")



#rename region levels for plotting
levels(cumulative_richness_traits.r$reg) <- c("Aleutian Islands", "E. Bering Sea", "Gulf of Mexico", "Gulf of Alaska", "Northeast US", "Newfoundland", "Southeast US", "Scotian Shelf", "West Coast US")

#plot richness over time predicted versus observed
cum_richness_plot_traits <- ggplot(aes(x=year, y = value, color = Richness, shape = Richness), data = cumulative_richness_traits.r) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0) +
  labs(x="Year", y = "Cumulative Richness") +
  facet_wrap(~reg, scales = "free_y") +
  theme_classic() +
  scale_color_manual(values = c("Observed" = 'coral3' , "Predicted" = 'deepskyblue4', "Predicted_null" = 'black')) +
  scale_shape_manual(values = c("Observed" = 19 , "Predicted" = 19, "Predicted_null" = 20)) +
  theme(plot.title = element_text(size = 17),
        legend.title = element_text(size = 17),
        legend.text=element_text(size = 15),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))

cum_richness_plot_traits

ggsave(cum_richness_plot_traits, file = "cum_richness_plot_traits.jpg")
ggsave(cum_richness_plot_traits, file = "cum_richness_plot_traits.eps")

```

In response to plots above, Malin: "Perhaps worth one more set of plots to understand what’s going on: each point would be a year, x-axis would be observed change in richness, y-axis would be predicted change in richness. One set of plots for each of the three models (null, temp, temp+trait), each sub-plot would be a region, as you have here."

```{r observed versus predicted change in richness}
setkey(cumulative_richness_traits.r, Richness, reg, year)
cumulative_richness_traits.r[, delta_value := value-shift(value), by = .(reg, Richness)]

#have to cast to wide to make x = observed change and y = predicted change
cumulative_richness_traits.r.w <- dcast(cumulative_richness_traits.r, reg + year ~ Richness, value.var = "delta_value")

```

```{r plotting observed versus predicted change in richness for trait models}

#plot richness over time predicted null versus observed
delta_richness_plot_null_traits <- ggplot(data = cumulative_richness_traits.r.w, aes(x = Observed, y = Predicted_null)) +
  geom_point() +
  labs(x="Observed Change\nin Richness", y = "Predicted Change in\nRichness (Null Model)") +
  facet_wrap(~reg) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, color = "grey", linetype = "dashed") +
  theme(plot.title = element_text(size = 17),
        legend.title = element_text(size = 17),
        legend.text=element_text(size = 15),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))

delta_richness_plot_null_traits

#plot richness over time predicted versus observed
delta_richness_plot_traits <- ggplot(data = cumulative_richness_traits.r.w, aes(x = Observed, y = Predicted)) +
  geom_point() +
  labs(x="Observed Change\nin Richness", y = "Predicted Change in\nRichness (Temperature & Trait Model)") +
  facet_wrap(~reg) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, color = "grey", linetype = "dashed") +
  theme(plot.title = element_text(size = 17),
        legend.title = element_text(size = 17),
        legend.text=element_text(size = 15),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))

delta_richness_plot_traits



ggsave(delta_richness_plot_null_traits, file = "delta_richness_plot_null_traits.jpg")
ggsave(delta_richness_plot_traits, file = "delta_richness_plot_traits.jpg")
```