---
title: "R Notebook"
output: html_notebook
---
```{r setup}

library(data.table)


load(file = "colonization_model_comparisons_dataframe.Rdata")
```

```{r manipulating colonization models}
#only use top 13 models
colonization_model_comparisons.r <- colonization_model_comparisons.nomean[order(AICc)][1:13,]
cols <- names(colonization_model_comparisons.r)[2:5]
colonization_model_comparisons.r[,(cols) := round(.SD,3), .SDcols=cols]
colonization_model_comparisons.r
```


Colonization Models

variable | coefficient | p | AICc | $\Delta$ AICc |
---------|-------------|----|-----|---|---|
| max_sbt_temp_change_abs |	0.239 |	0.000 |	6878.288 |	0.000 |
| max_sst_temp_change_abs_lag1 |	-0.240 |	0.001 |	6887.400 |	9.112 |
| min_sbt_temp_change_abs_lag5 |	-0.244 |	0.002 |	6888.859 |	10.572 |
| max_sbt_temp_change_abs_lag3 |	0.171 |	0.001 |	6888.938 |	10.650 |
| seas_sbt_temp_change_abs |	0.142 |	0.002 |	6889.498 |	11.211 |
| max_sst_temp_change_abs_lag8 |	0.191 |	0.002 |	6889.901 |	11.614 |
| seas_sst_temp_change_abs_lag8 |	0.127 |	0.004 |	6890.941 |	12.654 |
| seas_sbt_temp_change_abs_lag4 |	0.136 |	0.005 |	6891.397 |	13.109 |
| max_sbt_temp_change_lag8 |	-0.115 |	0.006 |	6891.566 |	13.279 |
| seas_sbt_temp_change_abs_lag1 |	0.126 |	0.007 |	6891.960 |	13.672 |

Relative Variable Importance: Calculated Akaike Weight for each model, divided by sum of all weights to get a value for relative model importance. Then, summed values for all models including the following variables. This only works when each variable is equally represented.

* Surface (n = 93) importance = 0.02 **versus** Bottom (n = 93) importance = 0.97
* Non lagged (n = 18) importance = 0.96 **versus** Lag (n = 168) importance = 0.03 (despite difference in n)
* Absolute change (n = 60) importance = 0.99 **versus** Raw temp (n = 66) importance = 0.003 **versus** change (n = 60) = 0.005


Extinction Models

variable | coefficient | p | AICc | $\Delta$ AICc |
---------|-------------|----|-----|---|---|
| seas_sbt_temp_change_abs_lag9 |	0.168 |	0.002 |	6207.213 |	0.000 |
| min_sst_temp_change_lag2 |	-0.095 |	0.011 |	6209.860 |	2.647 |
| min_sbt_temp_change_lag7 |	-0.116 |	0.011 |	6210.022 |	2.808 |
| seas_sbt_temp_change_abs |	0.116 |	0.013 |	6210.184 |		2.971 |
| max_sbt_temp_change_lag7 |	-0.100 |	0.015 |	6210.285 |	3.072 |
| seas_sbt_temp_lag9 |	0.087 |	0.017 |	6210.766 |	3.553 |
| seas_sst_temp_change_lag2 |	0.069 |	0.020 |	6210.908 |	3.695 |
| max_sbt_temp_change_abs_lag9 |	0.177 |	0.028 |	6211.488 |	4.275 |
| max_sst_temp_change_lag1 |	-0.098 |	0.038 |	6212.029 |	4.816 |
| max_sbt_temp_change_abs_lag3 |	0.117 |	0.043 |	6212.264 |	5.050 |

* Surface (n = 93) importance = 0.37 **versus** Bottom (n = 93) importance = 0.63
* Non lagged (n = 18) importance = 0.10 **versus** Lag (n = 168) importance = 0.90 (keep in mind change in n)
* Absolute change (n = 60) importance = 0.47 **versus** Raw temp (n = 66) importance = 0.20 **versus** change (n = 60) = 0.34

Comparison in lags between colonizations and extinctions
```{r RVI lags}
load("RVI_lags.Rdata")
RVI_lags
```

