---
title: "Dredge for Trait Models"
output: html_notebook
---

Here, I'm going to try again to use dredge. 

```{r setup}
library(MuMIn)
library(data.table)
library(lme4)
library(ggplot2)
library(dplyr)
load("spp_master_ztemp.traits.Rdata")
```

First, get rid of all rows that don't have all three trait variables
```{r get rid of rows without three trait variables}
spp_master_ztemp.traits.nonas <- na.omit(spp_master_ztemp.traits, cols=c("age.max", "age.maturity", "tl")) #gets rid of any rows missing any of these trait data values
spp_master_ztemp.traits.nonas.nomean <- spp_master_ztemp.traits.nonas[, grep("mean*", colnames(spp_master_ztemp.traits.nonas)):=NULL] #get rid of columns with mean temp variables
length(unique(spp_master_ztemp.traits.nonas$spp))
```

To see which variables to include... 
```{r}
cor(spp_master_ztemp.traits.nonas.nomean[,c("age.max", "age.maturity", "tl")])

mod1 <- lm(age.max ~ age.maturity, data = spp_master_ztemp.traits.nonas.nomean)
mod2 <- lm(age.max ~ tl, data = spp_master_ztemp.traits.nonas.nomean)
mod3 <- lm(age.maturity ~ tl, data = spp_master_ztemp.traits.nonas.nomean)

mod1
mod2
mod3
                                  
``` 

None super correlated, so I'll keep all in models

Prep for dredge
```{r prep for dredge}

#Colonization table
colonization.reduced <- spp_master_ztemp.traits.nonas.nomean[,c(1, 3, 6, 45:230, 239, 246, 251)]#reduces to columns we need

#Extinction table

extinction.reduced <- spp_master_ztemp.traits.nonas.nomean[,c(1, 3, 9, 45:230, 239, 246, 251)]


```

Making all models for colonization
```{r all models colonization}

#playing with year as factor
#can't get dredge to run above, so trying here with fewer variables
colonization.reduced[,year := as.factor(year)]

#going to make loop to make all models I need to look at with single variables
all.temp.variables <- colnames(colonization.reduced[,4:189])
all.temp.variables.0 <- c(0, all.temp.variables) #add 0 as option
all.trait.variables <- c(0, "tl", "age.max","age.maturity", "tl + age.max", "tl * age.max", "tl + age.maturity", "tl * age.maturity", "age.max + age.maturity", "age.max * age.maturity", "tl + age.max + age.maturity", "tl + age.max * age.maturity", "tl * age.max + age.maturity", "tl * age.max * age.maturity")

sign <- c("+", "*")

full_var_set <- as.data.table(expand.grid(all.temp.variables.0, sign, all.trait.variables))

full_var_set <- full_var_set[!(Var1 == "0" & Var3 == "0")] #eliminate models with both 0s

colonization_model_comparison_traits <- as.data.table(matrix(nrow = nrow(full_var_set))) 
colonization_model_comparison_traits[, temp_variable:=as.factor(V1)][, sign:=as.factor(V1)][, trait_variable:=as.factor(V1)][, AICc:=as.numeric(V1)][, converge:= as.logical(V1)]
colonization_model_comparison_traits[, V1 := NULL]

for (i in 1:nrow(full_var_set)){
 
  temp <- full_var_set$Var1[i]
  sign <- full_var_set$Var2[i]
  trait <- full_var_set$Var3[i]
  
  formula <- as.formula(paste("col ~", temp, sign, trait, "+ (1|reg) + (1|year)"))
  
  mod <- glmer(formula, family = binomial, data = colonization.reduced)
  
  colonization_model_comparison_traits[i,temp_variable := full_var_set$Var1[i]]
  colonization_model_comparison_traits[i,sign := full_var_set$Var2[i]]
  colonization_model_comparison_traits[i,trait_variable := full_var_set$Var3[i]]
  colonization_model_comparison_traits[i,AICc := AICc(mod)]
  
  warning <- mod@optinfo$conv$lme4$messages[2]
  converges <- ifelse(is.null(warning) == TRUE, T, F)
  
  colonization_model_comparison_traits[i, converge := converges]
  
  print(paste(i, nrow(colonization_model_comparison_traits), sep = "/"))
    
}

```
Above couldn't complete with all, system crashes with three interactions for both colonizatino and below in extinction. 

Making all models for extinction (RUN THIS TONIGHT)
```{r all models extinction}

#to figure out whether or not model converged
mod@optinfo$conv$lme4$messages[2]
mod2@optinfo$conv$lme4$messages[2]

#playing with year as factor
extinction.reduced[,year := as.factor(year)]

#full var set made above

extinction_model_comparison_traits <- as.data.table(matrix(nrow = nrow(full_var_set))) 
extinction_model_comparison_traits[, temp_variable:=as.factor(V1)][, sign:=as.factor(V1)][, trait_variable:=as.factor(V1)][, AICc:=as.numeric(V1)][, converge:= as.logical(V1)]
extinction_model_comparison_traits[, V1 := NULL]

for (i in 1:nrow(extinction_model_comparison_traits)){
 
  temp <- full_var_set$Var1[i]
  sign <- full_var_set$Var2[i]
  trait <- full_var_set$Var3[i]
  
  formula <- as.formula(paste("now_ext ~", temp, sign, trait, "+ (1|reg) + (1|year)"))
  
  mod <- glmer(formula, family = binomial, data = extinction.reduced)
  
  extinction_model_comparison_traits[i,temp_variable := full_var_set$Var1[i]]
  extinction_model_comparison_traits[i,sign := full_var_set$Var2[i]]
  extinction_model_comparison_traits[i,trait_variable := full_var_set$Var3[i]]
  extinction_model_comparison_traits[i,AICc := AICc(mod)]
  
  warning <- mod@optinfo$conv$lme4$messages[2]
  converges <- ifelse(is.null(warning) == TRUE, T, F)
  
  extinction_model_comparison_traits[i, converge := converges]
  
  print(paste(i, nrow(extinction_model_comparison_traits), sep = "/"))
    
}

save(extinction_model_comparison_traits, file = "col_ext_mod_ext_wtraits_extonly.Rdata")
load("col_ext_mod_ext_wtraits_extonly.Rdata")
save(colonization_model_comparison_traits, file = "col_ext_mod_ext_wtraits_colonly.Rdata")
load("col_ext_mod_ext_wtraits_colonly.Rdata")
```

Next step is to look at RVI

COLONIZATION

```{r RVI colonization traits}
#only raw temp has lag of 10 years, which means variable importance comparisons are unbalanced so I will get rid of these rows
colonization_model_comparisons_traits.lag9max <- colonization_model_comparison_traits[!grepl("lag10", temp_variable), ][converge == TRUE,]#don't know what to do about the fact so many models (1/) don't converge, but for here will just get rid of all models that don't converge


#add ∆AIC to tables
min_col_AICc_trait <- min(colonization_model_comparisons_traits.lag9max[, AICc])
colonization_model_comparisons_traits.lag9max[,"deltaAICc" := (AICc - min_col_AICc_trait)]

#add relative likelihood exp( -0.5 * ∆AIC score for that model)
colonization_model_comparisons_traits.lag9max[,"rel_likelihood" := exp((-0.5 * deltaAICc))]
#sum relative likelihoods across all models
colonization_model_comparisons_traits.lag9max.likelihoodsum <- sum(colonization_model_comparisons_traits.lag9max[, rel_likelihood])
#Akaike weight for a model is this rel_likelihood devided by the sum of these values across all models
colonization_model_comparisons_traits.lag9max[,"akaike_weight" := rel_likelihood/colonization_model_comparisons_traits.lag9max.likelihoodsum]

#I want to look at relative importance FOR 
#bottom/surface (bottom = includes sbt, grepl("sbt", variable))
col_bottom_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("sbt", temp_variable), ]$akaike_weight)
col_surface_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("sst", temp_variable), ]$akaike_weight)
  col_bottom_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("sbt", temp_variable), ]) #1103
  col_surface_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("sst", temp_variable), ]) #1092

#lag/not lagged (grepl("lag", variable))
col_lag_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("lag", temp_variable), ]$akaike_weight)
col_nolag_importance <- sum(colonization_model_comparisons_traits.lag9max[!grep("lag", temp_variable), ]$akaike_weight)
  col_lag_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("lag", temp_variable), ]) #1976
  col_nolag_importance.n <- count(colonization_model_comparisons_traits.lag9max[!grepl("lag", temp_variable), ]) #239


#absolute (grepl("abs", variable))
col_abs_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("abs", temp_variable), ]$akaike_weight)
  col_abs_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("abs", temp_variable), ])#788
  
#raw (!grepl("change"))
col_raw_importance <- sum(colonization_model_comparisons_traits.lag9max[!grep("change", temp_variable), ]$akaike_weight)
  col_raw_importance.n <- count(colonization_model_comparisons_traits.lag9max[!grepl("change", temp_variable), ]) #647

#change (grepl("change", temp_variable), (!grepl("abs")))
col_change_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("change", temp_variable), ][!grep("abs", temp_variable),]$akaike_weight)
  col_change_importance.n <- count(colonization_model_comparisons_traits.lag9max[grep("change", temp_variable), ][!grep("abs", temp_variable),])#780


#type of temp temp_variable (max, min, seas)
col_max_temp_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("max", temp_variable), ]$akaike_weight)
col_min_temp_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("min", temp_variable), ]$akaike_weight)
col_seas_temp_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("seas", temp_variable), ]$akaike_weight)
  col_max_temp_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("max", temp_variable), ]) #729
  col_min_temp_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("min", temp_variable), ]) #723
  col_seas_temp_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("seas", temp_variable), ]) #743
  

#traits included
col_age.max_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("age.max", trait_variable), ]$akaike_weight)
col_age.maturity_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("age.maturity", trait_variable), ]$akaike_weight)
col_tl_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("tl", trait_variable), ]$akaike_weight)
  col_age.max_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("age.max", trait_variable), ]) #963
  col_age.maturity_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("age.maturity", trait_variable), ]) #1076
  col_tl_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("tl", trait_variable), ]) #1004


#interaction between trait and temperature
col_interaction_temp_trait_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("\\*", sign), ]$akaike_weight)
  col_interaction_temp_trait_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("\\*", sign), ]) #793

#interaction between traits
col_interaction_trait_trait_importance <- sum(colonization_model_comparisons_traits.lag9max[grep("\\*", trait_variable), ]$akaike_weight)
  col_interaction_trait_trait_importance.n <- count(colonization_model_comparisons_traits.lag9max[grepl("\\*", trait_variable), ]) #44

#save this table
save(colonization_model_comparisons_traits.lag9max, file = "colonization_models_traits_with_likelihoods.Rdata")
```

```{r put RVIs for colonization into data table}
col_ext <- c("col","col","col","col","col", "col","col", "col", "col","col", "col","col", "col", "col","col")
type <- c("depth", "depth", "lag", "lag", "Change?", "Change?", "Change?", "Temp", "Temp", "Temp", "Trait", "Trait", "Trait", "Interaction", "Interaction")
variable <- c("bottom", "surface", "no_lag", "lagged", "absolute_value_change", "change", "raw", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "tempxtrait", "traitxtrait")
value <- c(col_bottom_importance, col_surface_importance, col_nolag_importance, col_lag_importance, col_abs_importance, col_change_importance, col_raw_importance, col_min_temp_importance, col_max_temp_importance, col_seas_temp_importance, col_age.max_importance, col_age.maturity_importance, col_tl_importance, col_interaction_temp_trait_importance, col_interaction_trait_trait_importance)

RVI_col_traits <- data.table(col_ext, type, variable, value)
```

EXTINCTION

```{r RVI extinction traits}
#only raw temp has lag of 10 years, which means variable importance comparisons are unbalanced so I will get rid of these rows
extinction_model_comparisons_traits.lag9max <- extinction_model_comparison_traits[!grepl("lag10", temp_variable), ][converge == TRUE,]#don't know what to do about the fact so many models (1/) don't converge, but for here will just get rid of all models that don't converge


#add ∆AIC to tables
min_ext_AICc_trait <- min(extinction_model_comparisons_traits.lag9max[, AICc])
extinction_model_comparisons_traits.lag9max[,"deltaAICc" := (AICc - min_ext_AICc_trait)]

#add relative likelihood exp( -0.5 * ∆AIC score for that model)
extinction_model_comparisons_traits.lag9max[,"rel_likelihood" := exp((-0.5 * deltaAICc))]
#sum relative likelihoods across all models
extinction_model_comparisons_traits.lag9max.likelihoodsum <- sum(extinction_model_comparisons_traits.lag9max[, rel_likelihood])
#Akaike weight for a model is this rel_likelihood devided by the sum of these values across all models
extinction_model_comparisons_traits.lag9max[,"akaike_weight" := rel_likelihood/extinction_model_comparisons_traits.lag9max.likelihoodsum]

#I want to look at relative importance FOR 
#bottom/surface (bottom = includes sbt, grepl("sbt", variable))
ext_bottom_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("sbt", temp_variable), ]$akaike_weight)
  ext_bottom_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("sbt", temp_variable), ]) #1142
ext_surface_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("sst", temp_variable), ]$akaike_weight)
  ext_surface_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("sst", temp_variable), ]) #1114
  
#lag/not lagged (grepl("lag", variable))
ext_lag_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("lag", temp_variable), ]$akaike_weight)
  ext_lag_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("lag", temp_variable),]) #2030
ext_nolag_importance <- sum(extinction_model_comparisons_traits.lag9max[!grep("lag", temp_variable), ]$akaike_weight)
  ext_nolag_importance.n <- count(extinction_model_comparisons_traits.lag9max[!grepl("lag", temp_variable), ]) #245

  #absolute (grepl("abs", variable))
ext_abs_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("abs", temp_variable), ]$akaike_weight)
  ext_abs_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("abs", temp_variable), ]) #810

  #raw (!grepl("change"))
ext_raw_importance <- sum(extinction_model_comparisons_traits.lag9max[!grep("change", temp_variable), ]$akaike_weight)
  ext_raw_importance.n <- count(extinction_model_comparisons_traits.lag9max[!grepl("change", temp_variable), ]) #669

#change (grepl("change", temp_variable), (!grepl("abs")))
ext_change_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("change", temp_variable), ][!grep("abs", temp_variable),]$akaike_weight)
  ext_change_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("change", temp_variable), ][!grepl("abs", temp_variable),])#796

#type of temp temp_variable (max, min, seas)
ext_max_temp_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("max", temp_variable), ]$akaike_weight)
  ext_max_temp_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("max", temp_variable), ]) #742
ext_min_temp_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("min", temp_variable), ]$akaike_weight)
  ext_min_temp_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("min", temp_variable), ]) #762
ext_seas_temp_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("seas", temp_variable), ]$akaike_weight)
  ext_seas_temp_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("seas", temp_variable), ]) #752

#traits included
ext_age.max_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("age.max", trait_variable), ]$akaike_weight)
ext_age.maturity_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("age.maturity", trait_variable), ]$akaike_weight)
ext_tl_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("tl", trait_variable), ]$akaike_weight)

  ext_age.max_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("age.max", trait_variable), ]) #1000
  ext_age.maturity_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("age.maturity", trait_variable), ]) #1112
  ext_tl_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("tl", trait_variable), ]) #1051

#interaction between trait and temperature
ext_interaction_temp_trait_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("\\*", sign), ]$akaike_weight)
  ext_interaction_temp_trait_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("\\*", sign), ]) #820
#interaction between traits
ext_interaction_trait_trait_importance <- sum(extinction_model_comparisons_traits.lag9max[grep("\\*", trait_variable), ]$akaike_weight)
  ext_interaction_trait_trait_importance.n <- count(extinction_model_comparisons_traits.lag9max[grepl("\\*", trait_variable), ]) #54

#save this table
save(extinction_model_comparisons_traits.lag9max, file = "extinction_models_traits_with_likelihoods.Rdata")
head(extinction_model_comparisons_traits.lag9max)
```

```{r put RVIs for extinction into data table}
col_ext <- c("ext","ext","ext","ext","ext", "ext","ext", "ext", "ext","ext", "ext","ext", "ext", "ext","ext")
type <- c("depth", "depth", "lag", "lag", "Change?", "Change?", "Change?", "Temp", "Temp", "Temp", "Trait", "Trait", "Trait", "Interaction", "Interaction")
variable <- c("bottom", "surface", "no_lag", "lagged", "absolute_value_change", "change", "raw", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "tempxtrait", "traitxtrait")
value <- c(ext_bottom_importance, ext_surface_importance, ext_nolag_importance, ext_lag_importance, ext_abs_importance, ext_change_importance, ext_raw_importance, ext_min_temp_importance, ext_max_temp_importance, ext_seas_temp_importance, ext_age.max_importance, ext_age.maturity_importance, ext_tl_importance, ext_interaction_temp_trait_importance, ext_interaction_trait_trait_importance)

RVI_ext_traits <- data.table(col_ext, type, variable, value)
```

Merge RVI tables for colonization and extinction with traits, and then graph
```{r merge RVI trait tables and plot}
RVI_trait_table <- as.data.table(rbind(RVI_col_traits, RVI_ext_traits)) #bind RVI for colonizations and extinctions

RVI_trait_table.r <- RVI_trait_table[!grep("surface", variable)][!grep("lagged", variable)]

#change order of factor levels
RVI_trait_table.r[, variable := factor(variable, levels = c("bottom", "no_lag", "raw", "change", "absolute_value_change", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "tempxtrait", "traitxtrait"))]

save(RVI_trait_table, file = "RVI_trait_table.Rdata")

ggplot(data = RVI_trait_table.r, aes(x=variable, y = value, fill = col_ext)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_discrete(breaks = c("bottom", "no_lag", "raw", "change", "absolute_value_change", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "tempxtrait", "traitxtrait"), labels = c("Bottom\nTemp", "No\nLag", "Raw\nTemp", "Change in \nTemp", "| Change in\n Temp |", "Min\nTemp", "Max\nTemp", "Seas", "Age\nMax", "Age\nMaturity", "Trophic\nLevel", "Temp*\nTrait", "Trait*\nTrait")) +
  labs(x = "Temperature and Trait Variables", y = "Relative Variable Importance")  +
  theme_bw()

```

Another way to visualize
```{r visualizing RVI with traits}
#put levels into correct order for viewing
RVI_trait_table[, variable := factor(variable, levels = c("bottom", "surface", "absolute_value_change", "change", "raw", "no_lag", "lagged", "Min", "Max", "Seas", "age.max", "age.maturity", "tl", "tempxtrait", "traitxtrait"))][,type := factor(type, levels = c("Change?", "depth", "lag", "Temp", "Trait", "Interaction"))]
#make types factors as well

#colonization
RVI_col_trait_plot <- ggplot(data = RVI_trait_table[col_ext == "col" & (type == "Trait" | type == "Interaction")], aes(x=variable, y = value)) +
  geom_bar(stat= "identity") +
  scale_x_discrete(breaks = c("age.max", "age.maturity", "tl", "tempxtrait", "traitxtrait"), labels = c("Maximum\nAge", "Age of\nMaturity", "Trophic\nLevel", "Temp Trait\nInteraction", "Trait Trait\nInteraction")) +
  labs(x = "Trait Variables", y = "Relative Variable Importance") +
  theme_classic() +
  theme(text = element_text(size = 20))
#extinction
RVI_ext_trait_plot <- ggplot(data = RVI_trait_table[col_ext == "ext" & (type == "Trait" | type == "Interaction")], aes(x=variable, y = value)) +
  geom_bar(stat= "identity") +
  scale_x_discrete(breaks = c("age.max", "age.maturity", "tl", "tempxtrait", "traitxtrait"), labels = c("Maximum\nAge", "Age of\nMaturity", "Trophic\nLevel", "Temp Trait\nInteraction", "Trait Trait\nInteraction")) +
  labs(x = "Trait Variables", y = "Relative Variable Importance") +
  theme_classic() +
  theme(text = element_text(size = 20))

save(RVI_trait_table, file = "RVI_trait_table.Rdata")
```
Let's see how the RVI importance varies by looking at specific lag values for these trait models
```{r make a plot here for variability across lag values for trait models}
#first, table for lags
lags_v_RVI_trait <- data.table(col_lag = c(0:9), col_RVI = 0, ext_RVI = 0)
#colonization
lags_v_RVI_trait[1,2] <- 1-sum(colonization_model_comparisons_traits.lag9max[grep("lag", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[2,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag1", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[3,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag2", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[4,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag3", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[5,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag4", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[6,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag5", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[7,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag6", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[8,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag7", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[9,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag8", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[10,2] <- sum(colonization_model_comparisons_traits.lag9max[grep("lag9", temp_variable), ]$akaike_weight)
#extinction
lags_v_RVI_trait[1,3] <- 1-sum(extinction_model_comparisons_traits.lag9max[grep("lag", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[2,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag1", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[3,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag2", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[4,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag3", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[5,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag4", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[6,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag5", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[7,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag6", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[8,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag7", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[9,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag8", temp_variable), ]$akaike_weight)
lags_v_RVI_trait[10,3] <- sum(extinction_model_comparisons_traits.lag9max[grep("lag9", temp_variable), ]$akaike_weight)

RVI_lags_traits <- ggplot(data = lags_v_RVI_trait, aes(x = col_lag)) +
  geom_line(aes(y = col_RVI), col = "blue") +
  geom_line(aes(y = ext_RVI), col = "red", linetype = "dashed") +
  labs(x = "Lag (years)", y = "Relative Variable Importance") +
  theme_classic() +
  theme(text = element_text(size = 22, color = "black")) +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = 0:1)

save(RVI_col_trait_plot, RVI_ext_trait_plot, RVI_lags_traits, file = "RVI_plots_traits.Rdata")

setEPS()
postscript("RVI_col_trait_plot.eps")
RVI_col_trait_plot
dev.off()

setEPS()
postscript("RVI_ext_trait_plot.eps")
RVI_ext_trait_plot
dev.off()

setEPS()
postscript("RVI_lags_trait_plot.eps")
RVI_lags_traits
dev.off()
```
