---
title: "Regional Richness Temp Change Figure"
output: html_notebook
---

```{r setup}
library(dplyr)
library(data.table)
library(ggplot2)

load("spp_master_ztemp_seus_buoys.Rdata")
load('comm_master.RData') #ryan's table
```


##Malin's way: average trends over all years
```{r functions}

trend <- function(y, x){ #function to extract coefficient
	return(coef(lm(y ~ x))[2])
}

scale <- function(x){ # function to scale by (ni-nmean)/sdn
	return((x-mean(x))/sd(x))
}

```

##Ryan's code and table
```{r ryan code table}
comm_master[reg=='neus', plot(year, reg_rich, type='o')]


# add trends and other vars
comm_master[,bt_trend:=trend(bt_ann,year), by=reg] #bottom temp change per year
comm_master[,bt_trend_abs:=abs(bt_trend)] #absolute value of bottom temp change per year
comm_master[,bt_trend_abs_sc:=scale(bt_trend_abs)] #scaled bottom temp change per year
comm_master[,year_sc:=scale(year)] #scaled year

# create summary data modeled richness and naive
trends <- comm_master[,.(rich=trend(reg_rich, year), rich_naive = trend(naive_rich, year), temp=trend(bt_ann,year)), by=reg] #calculate trend in richness per year and trend in annual temp per year

#modeled richness
setkey(comm_master, reg, year)
fd <- comm_master[,.(fyear=year[1:(.N-1)], richdiff=c(diff(reg_rich)), richdiff_naive = c(diff(naive_rich)), btdiff=c(diff(bt_ann))),by=reg]

#naive richness

# statistics
mod <- trends[,lm(rich ~ abs(temp))]
summary(mod)

mod2 <- fd[,lmer(richdiff ~ abs(btdiff) + (1|reg))]
summary(mod2)

mod3 <- comm_master[,lmer(reg_rich ~ year_sc + bt_trend_abs_sc + year_sc:bt_trend_abs_sc + (year_sc|reg))]
summary(mod3)
mod3b <- comm_master[,lmer(reg_rich ~ year_sc + bt_trend_abs_sc + (year_sc|reg))]
anova(mod3, mod3b)



# plot
quartz(width=5, height=4)
# pdf(width=5, height=4, file='figures/richnesstrend_vs_temperaturetrend.pdf')
par(las=1, mai=c(1,1,0.2, 0.2))
trends[,plot(abs(temp), rich, pch=16, cex=2, col='#1f78b4', ylab='Change in richness (spp/yr)', xlab='abs(Temperature trend) (°C/yr)')]
trends[,abline(mod, col='grey', lty=2)]

#dev.off()

# plot naive richness
#quartz(width=5, height=4)
# pdf(width=5, height=4, file='figures/richnesstrend_vs_temperaturetrend.pdf')
par(las=1, mai=c(1,1,0.2, 0.2))
trends[,plot(abs(temp), rich_naive, pch=16, cex=2, col='#1f78b4', ylab='Change in raw richness (spp/yr)', xlab='abs(Temperature trend) (°C/yr)')]
trends[,abline(mod, col='grey', lty=2)]

dev.off()
```

##My code and table
Add trends and new variables
```{r trends and new var}
#new column for richness
spp_master_ztemp_seus_buoy[,richness:=sum(present), by = .(year, reg)]

#raw
spp_master_ztemp_seus_buoy[,bt_trend_max:=trend(max_sbt_temp,year), by=reg] #average yearly bottom temp
spp_master_ztemp_seus_buoy[,bt_trend_min:=trend(min_sbt_temp,year), by=reg] #average yearly bottom temp
spp_master_ztemp_seus_buoy[,bt_trend_seas:=trend(seas_sbt_temp,year), by=reg] #average yearly bottom temp
spp_master_ztemp_seus_buoy[,bt_trend_mean:=trend(mean_sbt_temp,year), by=reg] #average yearly bottom temp
spp_master_ztemp_seus_buoy[,st_trend_mean:=trend(mean_sst_temp,year), by=reg] #average yearly bottom temp

```

Summary data
```{r summary data}
# create summary data
spp_master_richness <- unique(spp_master_ztemp_seus_buoy[,.(reg, year, richness,bt_trend_max, bt_trend_min, bt_trend_mean, bt_trend_seas, st_trend_mean)])

trends_zoe <- spp_master_richness[,.(rich=trend(richness, year), #calculate trend in richness per year and trend in annual temp per year
                                 bt_trend_max_raw = mean(bt_trend_max) ,
                                 bt_trend_min_raw = mean(bt_trend_min) ,
                                 bt_trend_mean_raw = mean(bt_trend_mean),
                                 bt_trend_seas_raw = mean(bt_trend_seas),
                                 st_trend_mean_raw = mean(st_trend_mean)),  by=reg]


trends_zoe[,bt_trend_max_raw_scaled := scale(bt_trend_max_raw)][,
                                 bt_trend_min_raw_scaled := scale(bt_trend_min_raw)][,
                                 bt_trend_mean_raw_scaled := scale(bt_trend_mean_raw)][,
                                 bt_trend_seas_raw_scaled := scale(bt_trend_seas_raw)][,
                                 st_trend_mean_raw_scaled := scale(st_trend_mean_raw)]

trends_zoe.r <- trends_zoe[,.SD[1], by = reg]


```

Plot Max Change and Abs
```{r plot max}
#Raw

#raw mods for abline
raw_max_mod <- lm(rich~bt_trend_max_raw, data = trends_zoe.r)
summary(raw_max_mod)$r.squared
anova(raw_max_mod)$'Pr(>F)'[1]
# R^2 =  0.1420429
#p value = 0.3173871


ggplot(data = trends_zoe) +
  geom_point(aes(x = bt_trend_max_raw, y = rich, col = reg), size = 4) +
  geom_abline(intercept = raw_max_mod$coefficients[[1]], slope = raw_max_mod$coefficients[[2]], col = "slategrey", linetype = "dashed") +
  #geom_point(aes(x = bt_trend_min_raw, y = rich), col = "blue") +
  #geom_abline(intercept = raw_min_mod$coefficients[[1]], slope = raw_min_mod$coefficients[[2]], col = "blue") +
  labs(x = "Average\nMaximum Temperature (°C)", y = "Average change\nin richness (spp/yr)") +
  theme_classic() +
  theme(text = element_text(size = 14))

ggsave(width = 4, height = 3.5, units = "in", path = "plots/richness_change", filename = "richness_trend_max_raw.jpeg")

```

Plot Min for Change and Abs
```{r min plots}
#Raw

#raw mods for abline
raw_min_mod <- lm(rich~bt_trend_min_raw, data = trends_zoe)
summary(raw_min_mod)$r.squared
anova(raw_min_mod)$'Pr(>F)'[1]
# R^2 = 0.05175603
#p value = 0.5560681



ggplot(data = trends_zoe) +
  geom_point(aes(x = bt_trend_min_raw, y = rich, col = reg), size = 4) +
  geom_abline(intercept = raw_min_mod$coefficients[[1]], slope = raw_min_mod$coefficients[[2]], col = "slategrey", linetype = "dashed") +
  #geom_point(aes(x = bt_trend_min_raw, y = rich), col = "blue") +
  #geom_abline(intercept = raw_min_mod$coefficients[[1]], slope = raw_min_mod$coefficients[[2]], col = "blue") +
  labs(x = "Average\nMinimum Temperature (°C)", y = "Average change\nin richness (spp/yr)") +
  theme_classic() +
  theme(text = element_text(size = 14))

ggsave(width = 4, height = 3.5, units = "in", path = "plots/richness_change", filename = "richness_trend_min_raw.jpeg")
  

```


Plot Mean for Change and Abs bottom
```{r mean plots}
#Raw

#raw mods for abline
raw_mean_mod <- lm(rich~bt_trend_mean_raw, data = trends_zoe)
raw_mean_mod_abs <- lm(rich~abs(bt_trend_mean_raw), data = trends_zoe)


#change in temp over time (unhash to compare with Ryan)
ggplot() +
  geom_point(data = trends_zoe, aes(x = bt_trend_mean_raw, y = rich, col = reg), size = 4) +
  geom_point(data = trends, aes(x = temp, y = rich_naive, col = reg), size = 4, shape = 1) + #compare with ryan's data
  geom_abline(intercept = raw_mean_mod$coefficients[[1]], slope = raw_mean_mod$coefficients[[2]], col = "slategrey", linetype = "dashed") +
  geom_abline(intercept = mod$coefficients[[1]], slope = mod$coefficients[[2]], col = "black", linetype = "dashed") +
  labs(x = "Trend in Mean Temperature (°C/yr)", y = "Average change\nin richness (spp/yr)") +
  theme_classic() +
  theme(text = element_text(size = 14))

#rename region levels for plotting
levels(trends_zoe$reg) <- c("Aleutian Islands", "E. Bering Sea", "Gulf of Mexico", "Gulf of Alaska", "Northeast US", "Newfoundland", "Southeast US", "Scotian Shelf", "West Coast US")

#absolute value of change in temp over time (unhash to compare with Ryan)
ggplot() +
  geom_point(data = trends_zoe, aes(x = abs(bt_trend_mean_raw), y = rich, col = reg), size = 4) +
  #geom_point(data = trends, aes(x = abs(temp), y = rich_naive, col = reg), size = 4, shape = 1) + #compare with ryan's data
  geom_abline(intercept = raw_mean_mod_abs$coefficients[[1]], slope = raw_mean_mod_abs$coefficients[[2]], col = "slategrey", linetype = "dashed") +
  #geom_abline(intercept = mod$coefficients[[1]], slope = mod$coefficients[[2]], col = "black", linetype = "dashed") + #compare with Ryan
  labs(x = "abs(Trend in Mean Temperature) (°C/yr)", y = "Average change\nin richness (spp/yr)", col = "Region") +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  scale_color_brewer(palette = "Set1")

ggsave(path = "plots/richness_change", filename = "richness_trend_mean_raw_abs.jpeg")
ggsave(path = "plots/richness_change", filename = "richness_trend_mean_raw_abs.eps")
  

```
Plot change and abs for surface (just for comparison)
```{r mean plots surface}
#Raw

#raw mods for abline
raw_mean_mod_st <- lm(rich~st_trend_mean_raw, data = trends_zoe)

#change in temp over time
ggplot() +
  geom_point(data = trends_zoe, aes(x = st_trend_mean_raw, y = rich, col = reg), size = 4) +
  geom_abline(intercept = raw_mean_mod$coefficients[[1]], slope = raw_mean_mod$coefficients[[2]], col = "slategrey", linetype = "dashed") +
  labs(x = "Trend in Mean STemperature (°C/yr)", y = "Average change\nin richness (spp/yr)") +
  theme_classic() +
  theme(text = element_text(size = 14))

#absolute value of change in temp over time
ggplot() +
  geom_point(data = trends_zoe, aes(x = abs(st_trend_mean_raw), y = rich, col = reg), size = 4) +
  geom_abline(intercept = raw_mean_mod_st$coefficients[[1]], slope = raw_mean_mod_st$coefficients[[2]], col = "slategrey", linetype = "dashed") +
  labs(x = "abs(Trend in Mean STemperature) (°C/yr)", y = "Average change\nin richness (spp/yr)") +
  theme_classic() +
  theme(text = element_text(size = 14))

ggsave(width = 4, height = 3.5, units = "in", path = "plots/richness_change", filename = "richness_trend_mean_raw.jpeg")
  

```

Plot Seasonality for Change and Abs
```{r Seasonality plots}
#Raw

#raw mods for abline
raw_seas_mod <- lm(rich~bt_trend_seas_raw, data = trends_zoe)
summary(raw_seas_mod)$r.squared
anova(raw_seas_mod)$'Pr(>F)'[1]
# R squared value: 0.3593094
#p value: 0.08800772


ggplot(data = trends_zoe) +
  geom_point(aes(x = bt_trend_seas_raw, y = rich, col = reg), size = 4) +
  geom_abline(intercept = raw_seas_mod$coefficients[[1]], slope = raw_seas_mod$coefficients[[2]], col = "slategrey", linetype = "dashed") +
  #geom_point(aes(x = bt_trend_seas_raw, y = rich), col = "blue") +
  #geom_abline(intercept = raw_seas_mod$coefficients[[1]], slope = raw_seas_mod$coefficients[[2]], col = "blue") +
  labs(x = "Average\nTemperature\nSeasonality (°C)", y = "Average change\nin richness (spp/yr)") +
  theme_classic() +
  theme(text = element_text(size = 14))

ggsave(width = 4, height = 3.5, units = "in", path = "plots/richness_change", filename = "richness_trend_seas_raw.jpeg")
  
 
```


Compare Zoe temp patterns to Ryan's temp patterns
```{r temp comparisons}
load("comm_master.RData")
ryan_reduced <- comm_master[,.(reg, year, naive_rich, bt_ann)][,name:="RYAN"]
zoe_reduced <- unique(spp_master_ztemp_seus_buoy[, bt_ann_z := mean(mean_sbt_temp), by = .(reg, year)][,name:="ZOE"][,.(reg, year, bt_ann_z)])

setkey(ryan_reduced, reg, year)
setkey(zoe_reduced, reg, year)
merge <- ryan_reduced[zoe_reduced]

ggplot(data = merge, aes(x = year)) +
  geom_point(aes(y = bt_ann), color = "blue") +
  geom_point(aes(y = bt_ann_z), color = "red") +
  facet_wrap(~reg) +
  theme_bw()
```


Zoe's temp
```{r zoe temp plots}
ggplot(data = spp_master_ztemp_seus_buoy, aes(x = year, y = mean_sbt_temp)) +
  geom_point() +
  facet_wrap(~reg) +
  theme_bw()

```

Ryan's temp
```{r Ryans temp plots}
ggplot(data = comm_master, aes(x = year, y = bt_ann)) +
  geom_point() +
  facet_wrap(~reg) +
  theme_bw()
```


