---
title: "SEUS Temp Data from Buoys"
output: html_notebook
---

This code models SEUS temp through time from NOAA temp buoys. Output: seus_temp_final_frombuoys.RData

```{r setup}
library(data.table)
library(ggplot2)
library(usmap)
library(lme4)
library(tidyverse)
library(lubridate)
library(here)

#Excel files made by copy pasting buoy data from https://www.ndbc.noaa.gov/ July 22, 2020

savannah <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","41008_savannah.csv"))
fryingpanNC <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","fpsn7_fryingpanNC.csv"))
canaveral <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","41009_canaveral.csv"))
hatteras <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","41001_east_hatteras.csv"))
staugustine <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","sauf1_staugustineFL.csv"))
SC_old <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","41005_SC.csv"))
charleston <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","41004_charleston.csv"))
savannahlight <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","svls1_savannahlight.csv"))
wrightsvilleNC <- fread(here::here("Data","Temperature","SEUS_buoy","NOAA_buoy_data","41037_wrightsvilleNC.csv"))

```

```{r subset each}
hatteras.r <- hatteras[, .(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]
fryingpanNC.r <- fryingpanNC[, .(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]
wrightsvilleNC.r <- wrightsvilleNC[, .(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]
charleston.r <- charleston[, .(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]
SC_old.r <- SC_old[, .(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]
savannah.r <- savannah[,.(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]
savannahlight.r <- savannahlight[,.(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]
staugustine.r <- staugustine[, .(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]
canaveral.r <- canaveral[, .(lat, lon, Location, YYYY, MM, DD, hh, ATMP, WTMP)]

#summary(hatteras.r) 
#summary(fryingpanNC.r) 
#summary(wrightsvilleNC.r) 
#summary(charleston.r) 
#summary(SC_old.r) 
#summary(savannah.r) 
#summary(savannahlight.r)
#summary(staugustine.r) 
#summary(canaveral.r) 

```

```{r merge}
se_hourly_temp <- rbind(savannah.r, fryingpanNC.r, canaveral.r, hatteras.r, staugustine.r, wrightsvilleNC.r, charleston.r, savannahlight.r, SC_old.r)

se_hourly_temp <- data.table(se_hourly_temp)

se_hourly_temp[, Location := as.factor(Location)]

summary(se_hourly_temp)
str(se_hourly_temp)

```

Turn any temp values over 50 to NA (probably incorrect temp reading)
```{r get rid of 999 temp values}
se_hourly_temp[WTMP > 50, WTMP := NA]

#delete all rows without data
se_hourly_temp.narm <- se_hourly_temp[complete.cases(WTMP, Location)]

```

Plot year versus whether we have data
```{r year versus data}
year_site_combos <- expand.grid(Location = levels(se_hourly_temp$Location), YYYY = c(1980:2014))

true_year_site_combos <- unique(se_hourly_temp, by = c("Location", "YYYY"))[,.(Location, YYYY)][, exist := 1]

data_over_time <- rbind(year_site_combos, true_year_site_combos, fill = T)

data_over_time.u <- unique(data_over_time, by = c("Location", "YYYY"), fromLast = T)

data_over_time.u[is.na(data_over_time.u)] <- 0

data_over_time.u$Location <- factor(data_over_time.u$Location, levels = c( "41009_canaveral", "sauf1_staugustineFL", "41008_savannah","41005_SC", "svls1_savannahlight", "41004_charleston", "fpsn7_fryingpanNC","41037_wrightsville",  "41001_east_hatteras"), labels = c("41009\ncanaveral", "sauf1\nstaugustineFL", "41008\nsavannah","41005\nSC", "svls1\nsavannahlight", "41004\ncharleston", "fpsn7\nfryingpanNC","41037\nwrightsville",  "41001\neast.hatteras"))


dates_over_time <- ggplot(data = data_over_time.u, aes(x = YYYY, y = Location, fill = as.factor(exist))) +
  geom_tile() +
  scale_fill_manual(breaks = c("0", "1"), values = c("white", "black"), labels = c("No", "Yes")) +
  theme_classic() +
  labs(x="Year", y = "Buoy ID and Location", fill = "Data\nAvailability")
  

```


```{r monthly averages}
se_hourly_temp.narm[,month_avg := mean(WTMP, na.rm = T), by = c("Location", "YYYY", "MM")]
se_hourly_temp.narm[,date := paste(YYYY, MM, sep = "-")]
se_hourly_temp.narm[,year := as.numeric(YYYY)]
se_hourly_temp.narm[,month := as.numeric(MM)]

se_hourly_temp.narm[,date.f := ymd(date, truncated = 1)]

se_monthly_temp <- unique(se_hourly_temp.narm[,.(Location, date.f, month_avg)])



```

Now, I will model 

mod <- lmer(T_t,i ~ YM_t + (1|B_i)) 

for each month-year t and each buoy i

```{r model monthly temp}


seus_temp_mod <- lmer(month_avg ~ as.factor(date.f) + (1|Location), data = se_monthly_temp) #month and year together

summary(seus_temp_mod)


```


```{r generate all possible Location, MM, YYYY combos and then apply predict}
se_hourly_temp.narm$Location <- factor(se_hourly_temp.narm$Location)

Location <- levels(as.factor(se_hourly_temp.narm$Location))
date.f <- levels(as.factor(se_hourly_temp.narm$date.f))
all_combinations <- expand.grid(Location, date.f)
names(all_combinations) <- c("Location", "date.f")

```

```{r predict}
preds <- predict(seus_temp_mod, newdata = all_combinations)

predicted_sst_seus_monthly <- cbind(all_combinations, preds)

```

Average across months 
```{r average across months}
predicted_sst_seus_monthly <- data.table(predicted_sst_seus_monthly)
predicted_sst_seus_monthly[, year := year(date.f)][, month := month(date.f)]
predicted_sst_seus_monthly[, sst_month_sst := mean(preds), by = c("month", "year")]

saveRDS(predicted_sst_seus_monthly, here::here("Data","Temperature","predicted_sst_seus_monthly.rds"))


```
Creating full temp variables, will have to go 12 months before first trawl which occurs in April 

 * mean_sbt_temp_change_lag
 * mean_sbt_temp_change_abs_lag
 
```{r when did earliest trawl occur?}
load(here::here("Data","Spp_observations","hauls_catch_Dec2017 1.RData")) #before = hauls_fulltempdata.Rdata
hauls <- data.table(hauls)
hauls.seus <- hauls[region == "SCDNR_SEUS"] #it occured in April typically except for 1989

hauls.seus[,min(as.numeric(month)), by = "year"]

#add month julian
#how many months?
year_month_key <- predicted_sst_seus_monthly[, .N, by = c('year', 'month')][,1:2]
max_julian_month <- nrow(year_month_key)

#sequence of 1:420
julian_months <- seq(1, max_julian_month, by=1)

#earliest = April, 12 months B4 April  

year_month_key <- cbind(year_month_key, "month_julian" = julian_months)

```

I need to use this key to add julian month to data tables for sst and sbt. 
```{r add julian month}
#surface
predicted_sst_seus_monthly <- predicted_sst_seus_monthly[year_month_key, on = c(year = 'year', month = 'month')]

```
 
```{r lags and change}
  #surface
    #raw
temp.vars <- c("mean", "max", "min", "seas")
      predicted_sst_seus_monthly[,mean_sst_temp :=numeric(.N)]
      predicted_sst_seus_monthly[,max_sst_temp :=numeric(.N)]
      predicted_sst_seus_monthly[,min_sst_temp :=numeric(.N)]
      predicted_sst_seus_monthly[,seas_sst_temp :=numeric(.N)]

#lag raw
      #new columns
raw.lag.sst.col.names <- CJ(temp.vars, 1:10)[, paste(temp.vars, 1:10, sep ="_sst_temp_lag")] #generate new raw column names

      #loop to make new column names
predicted_sst_seus_monthly[ , raw.lag.sst.col.names[1:40] := numeric(.N) ]
      #ended here
     
      
    #change
predicted_sst_seus_monthly[, mean_sst_temp_change := numeric(.N)]
predicted_sst_seus_monthly[, max_sst_temp_change := numeric(.N)]
predicted_sst_seus_monthly[, min_sst_temp_change := numeric(.N)]
predicted_sst_seus_monthly[, seas_sst_temp_change := numeric(.N)]

    
    #lag change
      #new columns
      change.lag.sst.col.names <- CJ(temp.vars, 1:9)[, paste(temp.vars, 1:9, sep ="_sst_temp_change_lag")]
      
      #loop to make new column names
      predicted_sst_seus_monthly[ ,change.lag.sst.col.names[1:36]:=numeric(.N) ]
      
    #abs
    predicted_sst_seus_monthly[, mean_sst_temp_change_abs := numeric(.N)]
    predicted_sst_seus_monthly[, max_sst_temp_change_abs := numeric(.N)]
    predicted_sst_seus_monthly[, min_sst_temp_change_abs := numeric(.N)]
    predicted_sst_seus_monthly[, seas_sst_temp_change_abs := numeric(.N)]
      
    #lag abs
      abs.change.lag.sst.col.names <- CJ(temp.vars, 1:9)[, paste(temp.vars, 1:9, sep ="_sst_temp_change_abs_lag")]
      
      #loop to make new column names
      predicted_sst_seus_monthly[ ,abs.change.lag.sst.col.names[1:36]:=numeric(.N) ]
  

  
#now, loop to fill these new columns, START AT 121
setorder(predicted_sst_seus_monthly, year, month)
first_year <- which(predicted_sst_seus_monthly==1990, arr.ind = T)[1,1][[1]]
for (i in first_year:nrow(predicted_sst_seus_monthly)) {
    this_month_julian <- predicted_sst_seus_monthly$month_julian[i] #identify month/year
    monthstosubtract <- seq(from = 11, by = 12, length.out = 11) 
    
    this_month_julian_reverse <- NA #month/year combo back in time
    
    for (j in 1:length(monthstosubtract)){
      this_month_julian_reverse[j] <- this_month_julian-monthstosubtract[j]
    }
    
  #surface raw
    subset_sst <- predicted_sst_seus_monthly[month_julian >= this_month_julian_reverse[1] & predicted_sst_seus_monthly$month_julian <= this_month_julian]
    predicted_sst_seus_monthly[(i),mean_sst_temp :=mean(subset_sst$sst_month_sst)] #mean temp over the past 12 months
    predicted_sst_seus_monthly[(i),max_sst_temp :=max(subset_sst$sst_month_sst)] #max temp over the past 12 months
    predicted_sst_seus_monthly[(i),min_sst_temp :=min(subset_sst$sst_month_sst)] #min temp over the past 12 months
    predicted_sst_seus_monthly[(i),seas_sst_temp := max_sst_temp - min_sst_temp] #seasonality over the past 12 months
    
  #surface raw lags 
      for (w in 2:length(this_month_julian_reverse)) {
        
        subset_sst <- predicted_sst_seus_monthly[month_julian >= this_month_julian_reverse[w] & month_julian < this_month_julian_reverse[w-1]] #subsets to correct year
        
        indx <- which(colnames(predicted_sst_seus_monthly) == paste0("mean_sst_temp_lag", w-1)) #finds column number with right name
        predicted_sst_seus_monthly[(i),(indx) := mean(subset_sst$sst_month_sst)] #assigns value to that cell
        indx <- which(colnames(predicted_sst_seus_monthly) == paste0("max_sst_temp_lag", w-1))
        predicted_sst_seus_monthly[(i),(indx) := max(subset_sst$sst_month_sst)]
        indx <- which(colnames(predicted_sst_seus_monthly) == paste0("min_sst_temp_lag", w-1))
        predicted_sst_seus_monthly[(i),(indx) := min(subset_sst$sst_month_sst)]
        indx <- which(colnames(predicted_sst_seus_monthly) == paste0("seas_sst_temp_lag", w-1))
        predicted_sst_seus_monthly[(i),(indx) := (get(paste0("max_sst_temp_lag", w-1))-get(paste0("min_sst_temp_lag", w-1)))]
      }
    
    #surface change
    predicted_sst_seus_monthly[(i),mean_sst_temp_change :=  mean_sst_temp-mean_sst_temp_lag1]
    predicted_sst_seus_monthly[(i),max_sst_temp_change :=  max_sst_temp-max_sst_temp_lag1]
    predicted_sst_seus_monthly[(i),min_sst_temp_change :=  min_sst_temp-min_sst_temp_lag1]
    predicted_sst_seus_monthly[(i),seas_sst_temp_change :=  seas_sst_temp-seas_sst_temp_lag1]
      
    #surface lag change
    for (h in 1:9) {
  #--mean
      indx_raw <- which(colnames(predicted_sst_seus_monthly) == paste0("mean_sst_temp_lag", h))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- predicted_sst_seus_monthly[(i), ..indx_raw]-predicted_sst_seus_monthly[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(predicted_sst_seus_monthly) == paste0("mean_sst_temp_change_lag", h)) #finds column number with right name to fill now
      predicted_sst_seus_monthly[(i),(indx_change) := temp_change] #assigns value to that cell
  #--max
      indx_raw <- which(colnames(predicted_sst_seus_monthly) == paste0("max_sst_temp_lag", h))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- predicted_sst_seus_monthly[(i), ..indx_raw]-predicted_sst_seus_monthly[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(predicted_sst_seus_monthly) == paste0("max_sst_temp_change_lag", h)) #finds column number with right name to fill now
      predicted_sst_seus_monthly[(i),(indx_change) := temp_change] #assigns value to that cell
  #--min
      indx_raw <- which(colnames(predicted_sst_seus_monthly) == paste0("min_sst_temp_lag", h))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- predicted_sst_seus_monthly[(i), ..indx_raw]-predicted_sst_seus_monthly[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(predicted_sst_seus_monthly) == paste0("min_sst_temp_change_lag", h)) #finds column number with right name to fill now
      predicted_sst_seus_monthly[(i),(indx_change) := temp_change] #assigns value to that cell
  #--seas
      indx_raw <- which(colnames(predicted_sst_seus_monthly) == paste0("seas_sst_temp_lag", h))
      indx_raw.p1 <- indx_raw + 1
      
      temp_change <- predicted_sst_seus_monthly[(i), ..indx_raw]-predicted_sst_seus_monthly[(i), ..indx_raw.p1] #change from lag2-lag1 in raw temp
      
      indx_change <- which(colnames(predicted_sst_seus_monthly) == paste0("seas_sst_temp_change_lag", h)) #finds column number with right name to fill now
      predicted_sst_seus_monthly[(i),(indx_change) := temp_change] #assigns value to that cell
    }
    
    #absolute value of change in sst
    predicted_sst_seus_monthly[(i), mean_sst_temp_change_abs := abs(mean_sst_temp_change)]
    predicted_sst_seus_monthly[(i), max_sst_temp_change_abs := abs(max_sst_temp_change)]
    predicted_sst_seus_monthly[(i), min_sst_temp_change_abs := abs(min_sst_temp_change)]
    predicted_sst_seus_monthly[(i), seas_sst_temp_change_abs := abs(seas_sst_temp_change)]
  
    #absolute value of change in sst LAGS
    
    for (j in 1:9) {
  #--mean
      indx_raw <- which(colnames(predicted_sst_seus_monthly) == paste0("mean_sst_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(predicted_sst_seus_monthly[(i),..indx_raw])
      indx_abs <- which(colnames(predicted_sst_seus_monthly) == paste0("mean_sst_temp_change_abs_lag", j)) #find the right column to input new value
      predicted_sst_seus_monthly[(i), (indx_abs) := temp_change_abs]
  #--max
      indx_raw <- which(colnames(predicted_sst_seus_monthly) == paste0("max_sst_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(predicted_sst_seus_monthly[(i),..indx_raw])
      indx_abs <- which(colnames(predicted_sst_seus_monthly) == paste0("max_sst_temp_change_abs_lag", j)) #find the right column to input new value
      predicted_sst_seus_monthly[(i), (indx_abs) := temp_change_abs]
  #--min
      indx_raw <- which(colnames(predicted_sst_seus_monthly) == paste0("min_sst_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(predicted_sst_seus_monthly[(i),..indx_raw])
      indx_abs <- which(colnames(predicted_sst_seus_monthly) == paste0("min_sst_temp_change_abs_lag", j)) #find the right column to input new value
      predicted_sst_seus_monthly[(i), (indx_abs) := temp_change_abs]
  #--seas
      indx_raw <- which(colnames(predicted_sst_seus_monthly) == paste0("seas_sst_temp_change_lag", j)) #find column with raw change value
      temp_change_abs <- abs(predicted_sst_seus_monthly[(i),..indx_raw])
      indx_abs <- which(colnames(predicted_sst_seus_monthly) == paste0("seas_sst_temp_change_abs_lag", j)) #find the right column to input new value
      predicted_sst_seus_monthly[(i), (indx_abs) := temp_change_abs]
    }
}  
```

Extract only April (and avg temperatures calculated in relation to April), because that's what's actually in trawl dataset
```{r only April}
seus_temp_surface <- predicted_sst_seus_monthly[year >= 1990 & month == 4, c(4, 8:131)][,reg := "sa"]

surface_colnames <- colnames(seus_temp_surface)

bottom_colnames <- str_replace(surface_colnames, "sst", "sbt")

seus_temp_bottom <- seus_temp_surface[,2:125]

colnames(seus_temp_bottom) <- bottom_colnames[2:125]

seus_temp_final <- cbind(seus_temp_surface[,.(reg)], seus_temp_surface[,1:125], seus_temp_bottom)

seus_temp_final_unique <- unique(seus_temp_final)

saveRDS(seus_temp_final_unique, here::here("Data","Temperature","SEUS_buoy", "seus_temp_final_frombuoys.rds"))
seus_temp_final <- readRDS(here::here("Data","Temperature","SEUS_buoy", "seus_temp_final_frombuoys.rds"))

#plot avg temp over time
seus_temp_final.r <- unique(seus_temp_final[,.(reg,year,max_sst_temp, seas_sst_temp, min_sst_temp)])
seus_temp_final.r.long <- melt(seus_temp_final.r, id.vars = 1:2, measure.vars = 3:5)

temp_avgs_over_time <- ggplot(data = seus_temp_final.r.long, aes(x = year, y = value, color = variable)) +
  geom_line() +
  scale_color_manual(name = "Temperature\nMetric", values = c("red","darkgrey","blue"), labels = c("Maximum\nTemperature\n","Temperature\nSeasonality\n","Minimum\nTemperature\n")) +
  labs(x = "Year", y = paste("Temperature (\u00b0C)")) +
  #lims(y = c(15,30)) +
  theme_classic()

ggsave(here::here("figures","Supplemental","seus_buoy_temp.eps"), height = 3, width = 8, unit = "in")

```
 



******
Graphing monthly temperatures for each region
```{r yearly min and max and seas from model}


predicted_sst_seus_monthly <- data.table(predicted_sst_seus_monthly)
predicted_sst_seus_monthly[, year := year(date.f)]
predicted_sst_seus_monthly[,sst_min_annual := min(preds), by = c("Location", "year")]
predicted_sst_seus_monthly[,sst_max_annual := max(preds), by = c("Location", "year")]
predicted_sst_seus_monthly[,sst_seas_annual := sst_max_annual - sst_min_annual]
```


```{r yearly min and max and seas from observations}
se_hourly_temp.narm <- data.table(se_hourly_temp.narm)
se_hourly_temp.narm[,sst_min_annual := min(month_avg), by = c("Location", "year")]
se_hourly_temp.narm[,sst_max_annual := max(month_avg), by = c("Location", "year")]
se_hourly_temp.narm[,sst_seas_annual := sst_max_annual - sst_min_annual]
```


Unique values
```{r unique model temp values}
seus_temp <- unique(predicted_sst_seus_monthly, by = c("Location", "year", "month"))

#reorder factors
seus_temp$Location <- factor(seus_temp$Location, levels = c("41001_east_hatteras","41037_wrightsville" , "fpsn7_fryingpanNC", "41004_charleston","svls1_savannahlight", "41005_SC", "41008_savannah","sauf1_staugustineFL", "41009_canaveral"))

ggplot(data = seus_temp, aes(x = year)) +
  geom_point(aes(y = sst_min_annual), color= "blue") +
  geom_point(aes(y = sst_max_annual), color = "red") +
  facet_wrap(~Location, ncol = 1) +
  ylim(10,35) +
  theme_bw()

ggsave(width = 3, height = 9, here::here("Data", "Temperature", "SEUS_buoy", "seus_buoy_data_modeled.png"))
```


```{r unique observations}
seus_temp_obs <- unique(se_hourly_temp.narm, by = c("Location", "year", "month"))

seus_temp_obs$Location <- factor(seus_temp_obs$Location, levels = c("41001_east_hatteras","41037_wrightsville" , "fpsn7_fryingpanNC", "41004_charleston","svls1_savannahlight", "41005_SC", "41008_savannah","sauf1_staugustineFL", "41009_canaveral"))

ggplot(data = seus_temp_obs, aes(x = year)) +
  geom_point(aes(y = sst_min_annual), color= "blue") +
  geom_point(aes(y = sst_max_annual), color = "red") +
  facet_wrap(~Location, ncol = 1) +
  ylim(10,35) +
  theme_bw()

ggsave(width = 3, height = 9, here::here("Data", "Temperature", "SEUS_buoy", "seus_buoy_data_observed.png"))


```
Plot modeled versus observed (modeled on y axis and observed on x axis). We need Location, Year, sst_min_annual, sst_max_annual, sst_seas_annual.
```{r observed versus modeled for each region}
#observed
observed_reduced_seus <- seus_temp_obs[,.(Location, year, month, sst_min_annual, sst_max_annual, sst_seas_annual, month_avg)][,type := "obs"]
predicted_reduced_seus <- seus_temp[,.(Location, year, month, sst_min_annual, sst_max_annual, sst_seas_annual, sst_month_sst)][,type := "pred"]

pred_obs_seus <-  predicted_reduced_seus[observed_reduced_seus, on = c("Location", "year", "month")]


#monthly
summary(pred_obs_seus)
ggplot(data = pred_obs_seus) +
  geom_point(aes(x = month_avg , y = sst_month_sst)) +
  geom_abline(slope = 1) +
  facet_wrap(~Location) +
  labs(x = "Observed Monthly Avg Temp", y = "Predicted Monthly Avg Temp") +
  theme_classic()

#ggsave(file = "predicted_obs_temp_seus_monthly.jpg")

#yearly
ggplot(data = pred_obs_seus) +
  geom_point(aes(x = i.sst_max_annual, y = sst_max_annual)) +
  geom_abline(slope = 1) +
  facet_wrap(~Location) +
  labs(x = "Observed Maximum Annual Temperature", y = "Predicted Maximum Annual Temperature") +
  theme_classic()

#ggsave(file = "predicted_obs_temp_seus_maxannual.jpg")

ggplot(data = pred_obs_seus) +
  geom_point(aes(x = i.sst_min_annual, y = sst_min_annual)) +
  geom_abline(slope = 1) +
  facet_wrap(~Location) +
  labs(x = "Observed Minimum Annual Temperature", y = "Predicted Minimum Annual Temperature") +
  theme_classic()

#ggsave(file = "predicted_obs_temp_seus_minannual.jpg")

ggplot(data = pred_obs_seus) +
  geom_point(aes(x = i.sst_seas_annual, y = sst_seas_annual)) +
  geom_abline(slope = 1) +
  facet_wrap(~Location) +
  labs(x = "Observed Annual Temperature Seasonality", y = "Predicted Annual Temperature Seasonality") +
  theme_classic()
  
#ggsave(file = "predicted_obs_temp_seus_seusannual.jpg")
```


```{r plot map of where buoys are}
library(dplyr)
library(ggrepel)
library(maps)
latlon_seus <- se_hourly_temp.narm[,.(lat,lon, Location)]

latlon_seus <- latlon_seus %>% 
  select(lat,lon, Location) %>% 
  unique


states <- map_data("state")
southeast <- subset(states, region %in% c("north carolina", "south carolina", "virginia", "georgia", "florida"))

(buoypoints <- ggplot(data = southeast) +
    geom_polygon(aes(x = long, y = lat, group = group), fill = NA, color = "black") +
    coord_fixed(1.3) +
	geom_point(data = latlon_seus,  # Add points
		aes(x = lon, y = lat), color = "red") +
    #geom_text_repel(data = latlon_seus, aes(x = lon, y = lat, label = Location), size = 2.5, nudge_x = 6) +
	theme_classic() +  # Remove ugly grey background
  labs( x = paste("Longitude (\u00b0E)"), y = paste("Latitude (\u00b0N)")))

ggsave(here::here("figures","Supplemental", "FigureS1a.eps"))





```
Merge figures for supplementary Figure S1
```{r supplementary figure S1}
library(cowplot)
(sup1_1 <- plot_grid(buoypoints, dates_over_time
                    +
                   theme(legend.position = "none")
                    ,ncol = 2, rel_widths = c(1,1), rel_heights = c(1,1), axis = "tblr", align = "h", labels = c("a.","b.")))

legend <- get_legend(dates_over_time)

(FigureS2_upper <- plot_grid(sup1_1, legend, ncol = 2, rel_widths = c(9,1)))

(FigureS2_final <- plot_grid(FigureS2_upper,temp_avgs_over_time, nrow = 2,ncol = 1, labels = c("","c."), rel_heights = c(5,3.5),vjust = 0))
  

ggsave(here::here("figures","Supplemental", "FigureS1_final.eps"), height = 6, width = 8)


```

