---
title: "Temp Through Time Gains Losses"
output: html_notebook
author: "Zoë Kitchel"
---

Temperature trends through time, checking for significance

We fit linear models using generalized least squares (errors are allowed to be correlated and/or have unequal variances), which is typically an effective way to model time series regression. 

These results are included in Figure S4 as *

```{r}
library(data.table)
library(nlme)

spp_master_ztemp_seus_buoy <- readRDS(here::here("Data","Spp_master","spp_master_ztemp_seus_buoy.rds"))
```

Unique year temp values
```{r}
spp_master_ztemp_seus_buoy.u <- unique(spp_master_ztemp_seus_buoy[,.(year, reg, min_sbt_temp, max_sbt_temp, seas_sbt_temp, min_sst_temp, max_sst_temp, seas_sst_temp)])

regions <- unique(spp_master_ztemp_seus_buoy.u[,reg])
```

Make GLS for each region and temp var to see which are significant, include in supplement and main text

corAR1: autoregressive 1 error structure (simple, means that closer points are more correlated, not independent observations)
```{r}
temp_change_time.full <- data.table(region = factor(NA), temp_var = as.character(NA), coef = as.numeric(NA), p = as.numeric(NA))

for (i in 1:length(regions)) {
  #empty dt to add to
  temp_change_time <- data.table(matrix(ncol = 4, nrow = 6))
  colnames(temp_change_time) <- c("region", "temp_var", "coef", "p")
  temp_change_time[,region := as.factor(region)][,temp_var := as.character(temp_var)][,coef := as.numeric(coef)][,p := as.numeric(p)]
  
  
  #subset full data table to single region
  subset <- spp_master_ztemp_seus_buoy.u[reg == regions[i],]
  
  #assign region column 
  temp_change_time[,region := regions[i]]
  
  #bottom
  max_sbt_mod <- gls(max_sbt_temp ~ year, correlation=corAR1(), method="ML", data = subset)
    temp_change_time[1,"temp_var"] <- "max_sbt_temp"
    temp_change_time[1,"coef"] <- summary(max_sbt_mod)$tTable[2]
    temp_change_time[1,"p"] <- summary(max_sbt_mod)$tTable[8]
  min_sbt_mod <- gls(min_sbt_temp ~ year,  correlation=corAR1(), data = subset)
    temp_change_time[2,"temp_var"] <- "min_sbt_temp"
    temp_change_time[2,"coef"] <- summary(min_sbt_mod)$tTable[2]
    temp_change_time[2,"p"] <- summary(min_sbt_mod)$tTable[8]
  seas_sbt_mod <- gls(seas_sbt_temp ~ year,  correlation=corAR1(), data = subset)
    temp_change_time[3,"temp_var"] <- "seas_sbt_temp"
    temp_change_time[3,"coef"] <- summary(seas_sbt_mod)$tTable[2]
    temp_change_time[3,"p"] <- summary(seas_sbt_mod)$tTable[8]
    
  #surface
  max_sst_mod <- gls(max_sst_temp ~ year,  correlation=corAR1(), data = subset)
     temp_change_time[4,"temp_var"] <- "max_sst_temp"
         temp_change_time[4,"coef"] <- summary(max_sst_mod)$tTable[2]
    temp_change_time[4,"p"] <- summary(max_sst_mod)$tTable[8]
  min_sst_mod <- gls(min_sst_temp ~ year,  correlation=corAR1(), data = subset)
    temp_change_time[5,"temp_var"] <- "min_sst_temp"
        temp_change_time[5,"coef"] <- summary(min_sst_mod)$tTable[2]
    temp_change_time[5,"p"] <- summary(min_sst_mod)$tTable[8]
  seas_sst_mod <- gls(seas_sst_temp ~ year,  correlation=corAR1(), data = subset)
    temp_change_time[6,"temp_var"] <- "seas_sst_temp"
    temp_change_time[6,"coef"] <- summary(seas_sst_mod)$tTable[2]
    temp_change_time[6,"p"] <- summary(seas_sst_mod)$tTable[8]
  
  temp_change_time.full <- rbind(temp_change_time.full, temp_change_time)
  
}


temp_change_time.full[,p_round := round(p,3)]

View(temp_change_time.full)
```

