---
title: "Figure 4: New with Revision, temp versus likelihood of gains vs. losses"
output: html_notebook
---


```{r setup}
library(ggplot2)
library(data.table)
library(lme4)
library(here)
library(cowplot)
library(DHARMa) #check model assumptions
library(glmmTMB)
```

Pull in spp master data

```{r pull in data}
spp_master_ztemp_seus_buoy_scaled <- readRDS(here::here("Data","Spp_master", "spp_master_ztemp_seus_buoy_scaled.rds")) #spp master, gain loss
spp_key <- fread(here::here("Data","Spp_master","spp_key.csv")) #key to link phylogenic info

spp_master_ztemp_seus_buoy_scaled <- spp_master_ztemp_seus_buoy_scaled[spp_key, on = "spp", nomatch = NULL] #this drops three species for GMEX that should not be there (see spp_add_genus_etc.Rmd code)

#delete first year each region is sampled (because gains and losses are based on transitions between years, so we need 2 years of observations)

spp_master_ztemp_seus_buoy_scaled[,min_year := min(year),.(reg)]

spp_master_ztemp_seus_buoy_scaled <- spp_master_ztemp_seus_buoy_scaled[year != min_year]

nrow(spp_master_ztemp_seus_buoy_scaled) == 17997    

#instead of col or now_ext =  0.3, add col_03 and now_ext_03  column for plotting
spp_master_ztemp_seus_buoy_scaled[,col_03 := ifelse(col == 1, 0.3, col)][,now_ext_03 := ifelse(now_ext == 1, 0.3, now_ext)]

```


Best Temp Only Model Gains


```{r}


#let's take a look at best performing model
mod_sppregrandom <- glmer(col ~ seas_sst_temp_scaled  + (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled, nAGQ = 0)
                       
```

##Plot Gains versus Seasonality

Here we use this method by [Bolker](https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#predictions-andor-confidence-or-prediction-intervals-on-predictions) to calculate confidence intervals using standard error. 

Steps:
figure out the model matrix X corresponding to the new data;
matrix-multiply X by the parameter vector β to get the predictions (or linear predictor in the case of GLM(M)s);
extract the variance-covariance matrix of the parameters V
compute XVX′
 to get the variance-covariance matrix of the predictions;
extract the diagonal of this matrix to get variances of predictions;
if computing prediction rather than confidence intervals, add the residual variance; (I don't do)
take the square-root of the variances to get the standard deviations (errors) of the predictions;
compute confidence intervals based on a Normal approximation;
for GL(M)Ms, run the confidence interval boundaries (not the standard errors) through the inverse-link function.

```{r}
#best predictor for fish and inverts: seas_sst_temp_scaled as x axis

#create data table of new data with 500 values of seas sst temp scaled in the range we see in the actual data
temp_gains_newdata <- data.table(seas_sst_temp_scaled =  runif(500,min(spp_master_ztemp_seus_buoy_scaled$seas_sst_temp_scaled, na.rm = T), max(spp_master_ztemp_seus_buoy_scaled$seas_sst_temp_scaled, na.rm = T)))


#I want x axis to be meaningful (unscaled)
mod_gain_reverse_scaledtemp <- lm(seas_sst_temp ~ seas_sst_temp_scaled, data = spp_master_ztemp_seus_buoy_scaled)

#add column with unscaled seasonality
temp_gains_newdata[,seas_sst_temp := mod_gain_reverse_scaledtemp$coefficients[2]*seas_sst_temp_scaled + mod_gain_reverse_scaledtemp$coefficients[1] ]

#best predictor for fish and inverts: seas_sst_temp_scaled as x axis

temp_gains_newdata$col <- predict(mod_sppregrandom,temp_gains_newdata, re.form = NA) #includes no random effects

mm <- model.matrix(terms(mod_sppregrandom),temp_gains_newdata) #"figure out the model matrix corresponding to the new data"

#"matrix-multiply X by the parameter vector β to get the predictions" &
#"extract the variance-covariance matrix of the parameters V" &
#"extract the diagonal of this matrix to get variances of predictions"
pvar1 <- diag(mm %*% tcrossprod(vcov(mod_sppregrandom),mm)) #vcov: variance/covariance matrix for fitted model object
#tcrossproduct (matrix cross product)
#diag = matrix diagonal

#tvar1 <- pvar1+VarCorr(mod_sppregrandom)$Subject[1]  ## must be adapted for more complex models
cmult <- 1.96 ## could use 1.96
temp_gains_newdata <- data.table(
    temp_gains_newdata
    , plo = temp_gains_newdata$col-cmult*sqrt(pvar1) #"take the square-root of the variances to get the standard deviations (errors) of the predictions & compute confidence intervals based on a Normal approximation"
    , phi = temp_gains_newdata$col+cmult*sqrt(pvar1)
  #  , tlo = newdat$distance-cmult*sqrt(tvar1)
  #  , thi = newdat$distance+cmult*sqrt(tvar1)
)

#inverse link plo and phi and col 
temp_gains_newdata[,plo_inv_log := exp(plo)/(1+exp(plo))][,phi_inv_log := exp(phi)/(1+exp(phi))][,col_inv_log := exp(col)/(1+exp(col))]


#plot
#NB: note that plo and phi are exp() because "run the confidence interval boundaries (not the standard errors) through the inverse-link function"
(gain_preds_temp_only <- ggplot() +
  geom_ribbon(data = temp_gains_newdata, aes(x = seas_sst_temp, ymin = plo_inv_log, ymax = phi_inv_log), fill = "lightgrey") +
  geom_point(data = spp_master_ztemp_seus_buoy_scaled, aes(x = seas_sst_temp, y = col_03), size = 2, shape = "|") +
  geom_line(data = temp_gains_newdata, aes(x = seas_sst_temp, y= col_inv_log)) +
  labs(x = "Seasonality of Sea Surface Temperature (\u00B0C)", y = "Probability of Gain") +
  theme_classic() +
  theme(text=element_text(size = 10)))


ggsave(gain_preds_temp_only, path = here::here("figures/Figure4"), file = "Figure4_gains_CI.jpg")
ggsave(gain_preds_temp_only, path = here::here("figures/Figure4"), file = "Figure4_gains_CI.eps")

```

Best Temp Only Model Losses

```{r}
#let's take a look at best performing model
mod_sppregrandom_losses <- glmer(now_ext ~ seas_sst_temp_lag3_scaled + (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled, nAGQ = 0)
```

Plot Losses versus Seasonality 
* seas_sst_temp_lag3_scaled
```{r}
#best predictor for fish and inverts: seas_sst_temp_lag3_scaled as x axis
temp_losses_newdata <- data.table(seas_sst_temp_lag3_scaled = runif(500,min(spp_master_ztemp_seus_buoy_scaled$seas_sst_temp_lag3_scaled, na.rm =  T), max(spp_master_ztemp_seus_buoy_scaled$seas_sst_temp_lag3_scaled, na.rm =  T)))

temp_losses_newdata$now_ext <- predict(mod_sppregrandom_losses,temp_losses_newdata, re.form = NA)

mm_loss <- model.matrix(terms(mod_sppregrandom_losses),temp_losses_newdata)

pvar1_loss <- diag(mm_loss %*% tcrossprod(vcov(mod_sppregrandom_losses),mm_loss)) #vcov: variance/covariance matrix for fitted model object
#tcrossproduct (matrix cross product)
#diag = matrix diagonal

#tvar1 <- pvar1+VarCorr(mod_sppregrandom)$Subject[1]  ## must be adapted for more complex models
cmult <- 1.96 ## could use 1.96
temp_losses_newdata <- data.frame(
    temp_losses_newdata
    , plo = temp_losses_newdata$now_ext-cmult*sqrt(pvar1_loss)
    , phi = temp_losses_newdata$now_ext+cmult*sqrt(pvar1_loss)
  #  , tlo = newdat$distance-cmult*sqrt(tvar1)
  #  , thi = newdat$distance+cmult*sqrt(tvar1)
)

#inverse link plo and phi
temp_losses_newdata[,plo_inv_log := exp(plo)/(1+exp(plo))][,phi_inv_log := exp(phi)/(1+exp(phi))][,col_inv_log := exp(col)/(1+exp(col))]

#I want x axis to be meaningful
mod_loss_reverse_scaledtemp <- lm(seas_sst_temp_lag3 ~ seas_sst_temp_lag3_scaled, data = spp_master_ztemp_seus_buoy_scaled)

temp_losses_newdata <- data.table(temp_losses_newdata)

temp_losses_newdata[,seas_sst_temp_lag3 := mod_loss_reverse_scaledtemp$coefficients[2]*seas_sst_temp_lag3_scaled + mod_loss_reverse_scaledtemp$coefficients[1] ]



#1.96 * standard error is 95% confidence interval 
(loss_preds_temp_only <- ggplot() +
  geom_ribbon(data = temp_losses_newdata, aes(x = seas_sst_temp_lag3, ymin = plo_inv_log, ymax = phi_inv_log), fill = "lightgrey") +
  geom_point(data = spp_master_ztemp_seus_buoy_scaled, aes(x = seas_sst_temp_lag3, y = now_ext_03), size = 2, shape = "|") +
  geom_line(data = temp_losses_newdata, aes(x = seas_sst_temp_lag3, y=col_inv_log)) +
  labs(x = "Seasonality of Sea Surface Temperature (\u00B0C)\nLagged Three Years", y = "Probability of Loss") +
  theme_classic() +
  theme(text=element_text(size = 10)))

ggsave(loss_preds_temp_only, path = here::here("figures/Figure4"), file = "Figure4_losses_CI.jpg")
ggsave(loss_preds_temp_only, path = here::here("figures/Figure4"), file = "Figure4_losses_CI.eps")
```

Merge these two together
```{r}
gains_losses_preds_CI <- plot_grid(gain_preds_temp_only, loss_preds_temp_only, ncol = 2, align = "hv", labels = c("a","b"))

ggsave(gains_losses_preds_CI, path = here::here("figures/Figure4"), file = "Figure4_gain_loss_merge.jpg", height = 4, width = 7, units = "in")
ggsave(gains_losses_preds_CI, path = here::here("figures/Figure4"), file = "Figure4_gain_loss_merge.eps", height = 4, width = 7, units = "in")
```

