---
title: "Figure 5"
output: html_notebook
---

```{r setup}
library(data.table)
library(ggplot2)
library(MuMIn)
library(lme4)

arrival.reduced_nosleeper.r <- readRDS(file = here::here("figures","Figure5","arrival.reduced_nosleeper.r.rds"))

arrival.reduced_nosleeper.r_annual <- arrival.reduced_nosleeper.r[!(reg %in% c("ai","goa","wctri"))]

departure.reduced_nosleeper.r <- readRDS(file = here::here("figures","Figure5","departure.reduced_nosleeper.r.rds"))

departure.reduced_nosleeper.r_annual <- departure.reduced_nosleeper.r[!(reg %in% c("ai","goa","wctri"))]

spp_master_ztemp_seus_buoy.traits_scaled.6regions <- readRDS(here::here("figures","Supplemental","annual_sensitivity","spp_master_ztemp_seus_buoy.traits_scaled.6regions.rds"))

```

Link up with spp key to add phylum, order, family, etc.
```{r pull in spp, order, family, etc.}
spp_key <- fread(here::here("Data","Spp_master","spp_key.csv"))

spp_master_ztemp_seus_buoy.traits_scaled.6regions <- spp_master_ztemp_seus_buoy.traits_scaled.6regions[spp_key, on = "spp"]
    
```

What do interactions look like?

##arrival with Traits
Best model: max_sst_temp_change_abs_lag1_scaled	*	age.max_ocean_scaled

All fish no traits: max_sst_temp_change_abs_lag1_scaled (same)


```{r predicted values arrival}
pal <- c("#F9BC77","#AD6D51","#7E7976","#62B3BE","#174A4A")

#best model with fish only 
mod_arrival_trait <- glmer(col ~ max_sst_temp_change_abs_lag1_scaled * age.max_ocean_scaled + (1|reg)  + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = "binomial", data = arrival.reduced_nosleeper.r_annual, nAGQ = 0)

#reverse to age max unscaled
mod_arrival_reverse_scaled_agemax <- lm(age.max_ocean ~ age.max_ocean_scaled, data = spp_master_ztemp_seus_buoy.traits_scaled.6regions) 

#forward to age max unscaled
mod_arrival_forward_scaled_agemax <- lm(age.max_ocean_scaled ~ age.max_ocean, data = spp_master_ztemp_seus_buoy.traits_scaled.6regions) 

#histogram of max age
hist(unique(spp_master_ztemp_seus_buoy.traits_scaled.6regions[,.(spp,age.max_ocean)]$age.max_ocean))

  #age max age match (save for departures)
age_max_match <- data.table(ages_we_want = c(1, 30, 80))
age_max_match[,age.max_ocean_scaled := mod_arrival_forward_scaled_agemax$coefficients[1] + mod_arrival_forward_scaled_agemax$coefficients[2]*ages_we_want]
        

#best predictor for fish only: max_sst_temp_change_abs_lag1_scaled as x axis
temp_arrival_newdata <- runif(500,min(spp_master_ztemp_seus_buoy.traits_scaled.6regions$max_sst_temp_change_abs_lag1_scaled, na.rm = T), max(spp_master_ztemp_seus_buoy.traits_scaled.6regions$max_sst_temp_change_abs_lag1_scaled, na.rm = T))

temp_age_newdata <- age_max_match[,age.max_ocean_scaled] #length

#make all possible temp trait combos
arrival_newdata_traits <- expand.grid("max_sst_temp_change_abs_lag1_scaled" = temp_arrival_newdata, "age.max_ocean_scaled" = temp_age_newdata)


arrival_newdata_traits$col <- predict(mod_arrival_trait, newdata = arrival_newdata_traits, re.form = NA)

preds_predictors_col <- data.table(arrival_newdata_traits)
preds_predictors_col <- preds_predictors_col[age_max_match, on = "age.max_ocean_scaled"]


mm <- model.matrix(terms(mod_arrival_trait), preds_predictors_col)

pvar1 <- diag(mm %*% tcrossprod(vcov(mod_arrival_trait),mm)) #vcov: variance/covariance matrix for fitted model object
#tcrossproduct (matrix cross product)
#diag = matrix diagonal

#tvar1 <- pvar1+VarCorr(mod_sppregrandom)$Subject[1]  ## must be adapted for more complex models
cmult <- 1.96 ## could use 1.96
preds_predictors_col <- data.table(
    preds_predictors_col
    , plo = preds_predictors_col$col-cmult*sqrt(pvar1)
    , phi = preds_predictors_col$col+cmult*sqrt(pvar1)
  #  , tlo = newdat$distance-cmult*sqrt(tvar1)
  #  , thi = newdat$distance+cmult*sqrt(tvar1)
)




#I want x axis to be meaningful
mod_arrival_reverse_scaledtemp <- lm(max_sst_temp_change_abs_lag1 ~ max_sst_temp_change_abs_lag1_scaled, data = spp_master_ztemp_seus_buoy.traits_scaled.6regions)


preds_predictors_col[,max_sst_temp_change_abs_lag1 := mod_arrival_reverse_scaledtemp$coefficients[2]*max_sst_temp_change_abs_lag1_scaled + mod_arrival_reverse_scaledtemp$coefficients[1] ]

(arrival_preds <- ggplot() +
  geom_ribbon(data = preds_predictors_col, aes(x = max_sst_temp_change_abs_lag1, ymin = exp(plo)/(1+exp(plo)), ymax = exp(phi)/(1+exp(phi)), fill = as.factor(ages_we_want)), alpha = 0.2) +
  geom_line(data = preds_predictors_col, aes(x = max_sst_temp_change_abs_lag1, y = exp(col)/(1+exp(col)), color = as.factor(ages_we_want)),size = 0.7) +
  labs(colour="Maximum\nAge", x = "Absolute Change in Maximum Surface\nTemperature Lagged One Year (\u00B0C)", y = "Probability of Gain") +
  theme_classic() +
  scale_color_manual(values = pal) +
    scale_fill_manual(values = pal, guide = 'none') +
  theme(text=element_text(size = 10)))

ggsave(arrival_preds, path = here::here("figures/Supplemental/annual_sensitivity"), file = "predicted_probs_departure_maxlength_nosleeper_annual_only.jpg")
ggsave(arrival_preds, path = here::here("figures/Supplemental/annual_sensitivity"), file = "predicted_probs_departure_maxlength_nosleeper_annual_only.eps")


```

##departure with Traits
Best model: 		max_sst_temp_lag1_scaled	*	age.max_ocean_scaled

Best model from analyses with temperature only: max_sst_temp_scaled

```{r predicted values departure}
#best model with fish only 
mod_departure_trait <- glmer(now_ext ~ max_sst_temp_lag1_scaled * age.max_ocean_scaled + (1|reg)  + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = "binomial", data = departure.reduced_nosleeper.r_annual, nAGQ = 0)

#reverse to length max unscaled
     mod_departure_reverse_scaled_agemax <- lm(age.max_ocean ~ age.max_ocean_scaled, data = spp_master_ztemp_seus_buoy.traits_scaled.6regions) 
      
     mod_departure_forward_scaled_agemax <- lm(age.max_ocean_scaled ~ age.max_ocean, data = spp_master_ztemp_seus_buoy.traits_scaled.6regions)


age_match <- data.table(ages_we_want = c(1,30,80))
age_match[,age.max_ocean_scaled := ages_we_want * coef(mod_departure_forward_scaled_agemax)[2] + coef(mod_departure_forward_scaled_agemax)[1]]


temp_departure_newdata <- runif(500,min(spp_master_ztemp_seus_buoy.traits_scaled.6regions$max_sst_temp_lag1_scaled, na.rm = T), max(spp_master_ztemp_seus_buoy.traits_scaled.6regions$max_sst_temp_lag1_scaled, na.rm = T))

temp_agemax_newdata <- age_match[,age.max_ocean_scaled]

departure_newdata_traits <- expand.grid("max_sst_temp_lag1_scaled" = temp_departure_newdata, "age.max_ocean_scaled" = temp_agemax_newdata)


departure_newdata_traits$now_ext <- predict(mod_departure_trait, newdata = departure_newdata_traits, re.form = NA)

preds_predictors_ext <- data.table(departure_newdata_traits)
preds_predictors_ext <- preds_predictors_ext[age_match, on = "age.max_ocean_scaled"]

mm_ext <- model.matrix(terms(mod_departure_trait), preds_predictors_ext)

pvar1_ext <- diag(mm_ext %*% tcrossprod(vcov(mod_departure_trait),mm_ext)) #vcov: variance/covariance matrix for fitted model object
#tcrossproduct (matrix cross product)
#diag = matrix diagonal

#tvar1 <- pvar1+VarCorr(mod_sppregrandom)$Subject[1]  ## must be adapted for more complex models
cmult <- 1.96 ## could use 1.96
preds_predictors_ext <- data.table(
    preds_predictors_ext
    , plo = preds_predictors_ext$now_ext-cmult*sqrt(pvar1_ext)
    , phi = preds_predictors_ext$now_ext+cmult*sqrt(pvar1_ext)
  #  , tlo = newdat$distance-cmult*sqrt(tvar1)
  #  , thi = newdat$distance+cmult*sqrt(tvar1)
)



#I want x axis to be meaningful
mod_departure_reverse_scaledtemp <- lm(max_sst_temp_lag1 ~ max_sst_temp_lag1_scaled, data = spp_master_ztemp_seus_buoy.traits_scaled.6regions)


preds_predictors_ext[,max_sst_temp_lag1 := mod_departure_reverse_scaledtemp$coefficients[2]*max_sst_temp_lag1_scaled + mod_departure_reverse_scaledtemp$coefficients[1] ]



(departure_preds <- ggplot() +
 geom_ribbon(data = preds_predictors_ext, aes(x = max_sst_temp_lag1, ymin = exp(plo)/(1+exp(plo)), ymax = exp(phi)/(1+exp(phi)), fill = as.factor(ages_we_want)), alpha = 0.2) +
  geom_line(data = preds_predictors_ext, aes(x = max_sst_temp_lag1, y = exp(now_ext)/(1+exp(now_ext)), color = as.factor(ages_we_want)),size = 0.7) +
  labs(colour="Maximum\nAge", x = "Maximum Surface Temperature\n Lagged One Year (\u00B0C)", y = "Probability of Loss") +
  theme_classic() +
  scale_color_manual(values = pal) +
    scale_fill_manual(values = pal, guide = 'none') +
  theme(text=element_text(size = 10)))


ggsave(departure_preds, path = here::here("figures/Supplemental/annual_sensitivity"), file = "predicted_probs_departure_maxlength_nosleeper_annual_only.jpg")
ggsave(departure_preds, path = here::here("figures/Supplemental/annual_sensitivity"), file = "predicted_probs_departure_maxlength_nosleeper_annual_only.eps")


```


Merge marginal effects plots
```{r}
library(cowplot)
mar_effects <- plot_grid(arrival_preds + theme(legend.position = "null"),departure_preds + theme(legend.position = "null"),get_legend(departure_preds),rel_widths = c(10,10,2), labels = c("a.","b.",""), label_size = 10, cols = 3, rows = 1)



ggsave(mar_effects, width = 7, height = 3.3, path = here::here("figures/Supplemental/annual_sensitivity"), file = "Figure5_margeffect_nosleeper_annual_only.jpg")
ggsave(mar_effects, width = 7, height = 3.3, path = here::here("figures/Supplemental/annual_sensitivity"), file = "Figure5_margeffect_nosleeper_annual_only.eps")

```