---
title: "Simple Model Simulations"
output: html_notebook
---

I'm going to try running some simulations to see how models will perform in response to a sudden change in temperature. 

```{r setup}
library(data.table)
library(ggplot2)
library(lme4)

load("spp_master_ztemp_scaled.RData")
load("colonization_model_comparison_spprandom_scaled_nomean.Rdata")
load("extinction_model_comparison_spprandom_scaled_nomean.Rdata")
```

Best colonization model: max_sbt_temp_change_abs_scaled
```{r col mod}
col_mod <- glmer(col~max_sbt_temp_change_abs_scaled + (1|spp) + (1|reg), family = binomial, data = spp_master_ztemp_scaled, nAGQ = 0)
```

Best extinciton model: min_sst_temp_change_lag2_scaled
```{r ext mod}
ext_mod <- glmer(now_ext~min_sst_temp_change_lag2_scaled + (1|spp) + (1|reg), family = binomial, data = spp_master_ztemp_scaled, nAGQ = 0)
```

What's typical for temp data?
```{r temp data summary}
summary(spp_master_ztemp_scaled$max_sbt_temp_change_abs_scaled)
hist(spp_master_ztemp_scaled$max_sbt_temp_change_abs_scaled) #this doesn't look right, look into later
summary(spp_master_ztemp_scaled$min_sst_temp_change_lag2_scaled)
hist(spp_master_ztemp_scaled$min_sst_temp_change_lag2_scaled)
```


Simulate temperature data
```{r sim temp data}

#colonization
year <- seq(1,10, by =1)

#No temp change
col_temps <- rep(0,10)
#Medium temp change
col_temps_2 <- c(rep(0,5), 2, rep(0,4))
#Significant temp change
col_temps_3 <- c(rep(0,5), 5, rep(0,4))

#extinction
#No temp change
ext_temps <- rep(0,10)
#Medium temp change positive
ext_temps_2 <- c(rep(0,5), 2, rep(0,4))
#Medium temp change negative
ext_temps_3 <- c(rep(0,5), -2, rep(0,4))
#Large temp change positive
ext_temps_4 <- c(rep(0,5), 5, rep(0,4))
#Large temp change negative
ext_temps_5 <- c(rep(0,5), -5, rep(0,4))

```

Predictions: No Change for Colonization
```{r predictions no change col}

#NO CHANGE for COLONIZATION
#re.form = NA to ignore random effects

col_temps.dt <- data.table(year = year, max_sbt_temp_change_abs_scaled = col_temps, col = NA)

prob <- predict(col_mod,newdata=col_temps.dt, type="response", re.form=NA)

col_temps.dt[, col := prob]


plot(col_temps.dt$year, col_temps.dt$col)
```
Predictions: Small perturbation col 
```{r predictions small change col}

#NO CHANGE for COLONIZATION
#re.form = NA to ignore random effects

col_temps.dt_2 <- data.table(year = year, max_sbt_temp_change_abs_scaled = col_temps_2, col = NA)

prob2 <- predict(col_mod,newdata=col_temps.dt_2, type="response", re.form=NA)

col_temps.dt_2[, col := prob2]


plot(col_temps.dt_2$year, col_temps.dt_2$col)
```

Predictions: Large perturbation col 
```{r predictions large change col}

#NO CHANGE for COLONIZATION
#re.form = NA to ignore random effects

col_temps.dt_3 <- data.table(year = year, max_sbt_temp_change_abs_scaled = col_temps_3, col = NA)

prob3 <- predict(col_mod,newdata=col_temps.dt_3, type="response", re.form=NA)

col_temps.dt_3[, col := prob3]


plot(col_temps.dt_3$year, col_temps.dt_3$col)
```

Predictions: No perturbation ext
```{r predictions no change ext}

#NO CHANGE for COLONIZATION
#re.form = NA to ignore random effects

ext_temps.dt <- data.table(year = year, min_sst_temp_change_lag2_scaled = ext_temps, ext = NA)

prob <- predict(ext_mod,newdata=ext_temps.dt, type="response", re.form=NA)

ext_temps.dt[, ext := prob]


plot(ext_temps.dt$year, ext_temps.dt$ext)
```
Predictions: Small positive perturbation ext
```{r predictions small + ext}

#NO CHANGE for COLONIZATION
#re.form = NA to ignore random effects

ext_temps.dt_2 <- data.table(year = year, min_sst_temp_change_lag2_scaled = ext_temps_2, ext = NA)

prob_2 <- predict(ext_mod,newdata=ext_temps.dt_2, type="response", re.form=NA)

ext_temps.dt_2[, ext := prob_2]


plot(ext_temps.dt_2$year, ext_temps.dt_2$ext)
```

Predictions: Small negative perturbation ext
```{r predictions small - ext}

ext_temps.dt_3 <- data.table(year = year, min_sst_temp_change_lag2_scaled = ext_temps_3, ext = NA)

prob_3 <- predict(ext_mod,newdata=ext_temps.dt_3, type="response", re.form=NA)

ext_temps.dt_3[, ext := prob_3]


plot(ext_temps.dt_3$year, ext_temps.dt_3$ext)

cor(spp_master_ztemp_scaled$max_sbt_temp_change_abs_scaled, spp_master_ztemp_scaled$max_sst_temp_change_abs_lag1_scaled)

```

Predictions: Large positive perturbation ext
```{r predictions large + ext}

ext_temps.dt_4 <- data.table(year = year, min_sst_temp_change_lag2_scaled = ext_temps_4, ext = NA)

prob_4 <- predict(ext_mod,newdata=ext_temps.dt_4, type="response", re.form=NA)

ext_temps.dt_4[, ext := prob_4]


plot(ext_temps.dt_4$year, ext_temps.dt_4$ext)
```

Predictions: Large negative perturbation ext
```{r predictions large - ext}

ext_temps.dt_5 <- data.table(year = year, min_sst_temp_change_lag2_scaled = ext_temps_5, ext = NA)

prob_5 <- predict(ext_mod,newdata=ext_temps.dt_5, type="response", re.form=NA)

ext_temps.dt_5[, ext := prob_5]


plot(ext_temps.dt_5$year, ext_temps.dt_5$ext)
```

