---
title: "Gain and Loss Final Models: EXCLUDING AI, GOA, WCTRI"
output: html_notebook
---

This is a cleaned version of 04_24_19 Temp link trait fish.Rmd/11_20_19_Temp_link_traits_fish.Rmd.
Changes
* improved formatting
* easier to follow
* species and region as random effects
* scaling temperature values
* setting nACQ to 0, see why [here](https://stats.stackexchange.com/questions/77313/why-cant-i-match-glmmTMB-family-binomial-output-with-manual-implementation-of-g). 

```{r setup}
library(MuMIn)
library(lme4)
library(data.table)
library(ggplot2)
library(here)
library(wesanderson) # to color plots

spp_master_ztemp_seus_buoy_scaled <- readRDS(here::here("Data","Spp_master","spp_master_ztemp_seus_buoy_scaled.rds"))


```

Delete all observations of regions that do not sample every year
```{r delete obs of regs not annually sampled}
levels(spp_master_ztemp_seus_buoy_scaled$reg)

spp_master_ztemp_seus_buoy_scaled_annual_obs <- spp_master_ztemp_seus_buoy_scaled[!(reg %in% c("ai","goa","wctri")),]

nrow(spp_master_ztemp_seus_buoy_scaled_annual_obs) == 16090
```

Link up with spp key to add phylum, order, family, etc.
```{r pull in spp, order, family, etc.}
spp_key <- fread(here::here("Data","Spp_master","spp_key.csv"))

spp_master_ztemp_seus_buoy_scaled_annual_obs <- spp_key[spp_master_ztemp_seus_buoy_scaled_annual_obs, on = "spp", nomatch = NULL]

#delete first year each region is sampled (because gains and losses are based on transitions between years, so we need 2 years of observations)

spp_master_ztemp_seus_buoy_scaled_annual_obs[,min_year := min(year),.(reg)]

spp_master_ztemp_seus_buoy_scaled_annual_obs <- spp_master_ztemp_seus_buoy_scaled_annual_obs[year != min_year]
    
nrow(spp_master_ztemp_seus_buoy_scaled_annual_obs) == 15399
```

Edit column names for understanding
```{r}
spp_master_ztemp_seus_buoy_scaled_annual_obs[,gain :=col][,loss := now_ext]
```


Running, and ranking gain models without traits

##Here, I scaled all temperature values across all regions, and included species as a random effect. In order to do this, I had to change the algorithm slightly--nAGQ = 0.

```{r gain models without traits}
#names of scaled columns
variables_scaled <- grep("*_scaled", names(spp_master_ztemp_seus_buoy_scaled_annual_obs), value = T)

#we don't end up looking at mean, delete here
variables_scaled <- variables_scaled[-grep('mean',variables_scaled)]

#going to make loop to make all models I need to look at with a single temperature variable

gain_model_comparison_annual_obs_taxonomy <- as.data.table(matrix(nrow = length(variables_scaled)))
gain_model_comparison_annual_obs_taxonomy[, variable:=as.factor(V1)][, coef:=as.numeric(V1)][, p_value:=as.numeric(V1)][, AICc:=as.numeric(V1)]
gain_model_comparison_annual_obs_taxonomy[, V1 := NULL]

for (i in 1:length(variables_scaled)){
  mod <- glmer(col ~ get(variables_scaled[i]) + (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)
  gain_model_comparison_annual_obs_taxonomy[i,variable := variables_scaled[i]]
  gain_model_comparison_annual_obs_taxonomy[i,coef := coef(summary(mod))[,"Estimate"][2]]
  gain_model_comparison_annual_obs_taxonomy[i,p_value := coef(summary(mod))[,"Pr(>|z|)"][2]]
  gain_model_comparison_annual_obs_taxonomy[i,AICc := AICc(mod)]
  
  print(paste(i, length(variables_scaled), sep = "/"))
    
}


saveRDS(object = gain_model_comparison_annual_obs_taxonomy, file = here::here("Model_results","gain_model_comparison_annual_obs_taxonomy.rds"))

fwrite(gain_model_comparison_annual_obs_taxonomy, file = here::here("Model_results","gain_model_comparison_annual_obs_taxonomy.csv"))

#saveRDS(object = gain_model_comparison_annual_obs_taxonomy, file ="gain_model_comparison_annual_obs_taxonomy.rds")

#fwrite(gain_model_comparison_annual_obs_taxonomy, file ="gain_model_comparison_annual_obs_taxonomy.csv")


gain_model_comparison_annual_obs_taxonomy <- readRDS(here::here("Model_results","gain_model_comparison_annual_obs_taxonomy.rds"))

```

Loss Models without Traits
```{r loss models without traits}

loss_model_comparison_annual_obs_taxonomy <- as.data.table(matrix(nrow = length(variables_scaled)))
loss_model_comparison_annual_obs_taxonomy[, variable:=as.factor(V1)][, coef:=as.numeric(V1)][, p_value:=as.numeric(V1)][, AICc:=as.numeric(V1)]
loss_model_comparison_annual_obs_taxonomy[, V1 := NULL]

for (i in 1:length(variables_scaled)){
  mod <- glmer(now_ext ~ get(variables_scaled[i]) + (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)
  loss_model_comparison_annual_obs_taxonomy[i,variable := variables_scaled[i]]
  loss_model_comparison_annual_obs_taxonomy[i,coef := coef(summary(mod))[,"Estimate"][2]]
  loss_model_comparison_annual_obs_taxonomy[i,p_value := coef(summary(mod))[,"Pr(>|z|)"][2]]
  loss_model_comparison_annual_obs_taxonomy[i,AICc := AICc(mod)]
  
  print(paste(i, length(variables_scaled), sep = "/"))
    
}

saveRDS(object = loss_model_comparison_annual_obs_taxonomy, file = here::here("Model_results","loss_model_comparison_annual_obs_taxonomy.rds"))
fwrite(loss_model_comparison_annual_obs_taxonomy, file = here::here("Model_results","loss_model_comparison_annual_obs_taxonomy.csv"))

loss_model_comparison_annual_obs_taxonomy <- readRDS(here::here("Model_results","loss_model_comparison_annual_obs_taxonomy.rds"))


```

Relative Variable Importance


How does the null model with only random effects perform?
```{r null models}
#best models
gain_mod_scaled_1 <- glmer(col ~ seas_sst_temp_scaled+ (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)

gain_mod_scaled_1 <- glmer(col ~ seas_sst_temp_scaled+ (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)

gain_mod_scaled_1 <- glmer(col ~ seas_sst_temp_scaled+ (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)

gain_mod_scaled_1 <- glmer(col ~ seas_sst_temp_scaled+ (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)

loss_mod_scaled <- glmer(now_ext ~ seas_sst_temp_lag3_scaled + (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)

#add null to tables
intercept_gain_mod_scaled <- glmer(col ~ (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)

intercept_loss_mod_scaled <- glmer(now_ext ~ (1|reg) + (1|phylum) + (1|class) + (1|order) + (1|family) + (1|genus) + (1|spp), family = binomial, data = spp_master_ztemp_seus_buoy_scaled_annual_obs, nAGQ = 0)

#best model better than null? yes
AICc(intercept_gain_mod_scaled)-min(gain_model_comparison_annual_obs_taxonomy$AICc)
AICc(intercept_loss_mod_scaled)-min(loss_model_comparison_annual_obs_taxonomy$AICc)

#calculate pseudo R squared
r.squaredGLMM(gain_mod_scaled_1)
#                    R2m       R2c
#theoretical 0.006089580 0.4159003
#delta       0.002429494 0.1659273
r.squaredGLMM(loss_mod_scaled)
#                    R2m       R2c
#theoretical 0.013836934 0.4562251
#delta       0.004567733 0.1506052

r.squaredGLMM(intercept_gain_mod_scaled)
#            R2m       R2c
#theoretical   0 0.4175993
#delta         0 0.1668969
            
r.squaredGLMM(intercept_loss_mod_scaled)
 #           R2m       R2c
#theoretical   0 0.4192593
#delta         0 0.1323742


```


Relative Variable Importance for Gain Models
```{r RVI gains}
#add ∆AIC to tables
min_gain_AICc <- min(gain_model_comparison_annual_obs_taxonomy[, AICc])
gain_model_comparison_annual_obs_taxonomy[,"deltaAICc" := (AICc - min_gain_AICc)]

#add relative likelihood exp( -0.5 * ∆AIC score for that model)
gain_model_comparison_annual_obs_taxonomy[,"rel_likelihood" := exp((-0.5 * deltaAICc))]
#sum relative likelihoods across all models
gain_model_comparison_annual_obs_taxonomy.likelihoodsum <- sum(gain_model_comparison_annual_obs_taxonomy[, rel_likelihood])
#Akaike weight for a model is this rel_likelihood divided by the sum of these values across all models
gain_model_comparison_annual_obs_taxonomy[,"akaike_weight" := rel_likelihood/gain_model_comparison_annual_obs_taxonomy.likelihoodsum]

#I want to look at relative importance FOR:

#bottom/surface (bottom = includes sbt, grepl("sbt", variable))
gain_bottom_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("sbt", variable), ]$akaike_weight)
gain_surface_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("sst", variable), ]$akaike_weight)

#lag/not lagged (grepl("lag", variable))
gain_lag_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag", variable), ]$akaike_weight)
gain_nolag_importance <- sum(gain_model_comparison_annual_obs_taxonomy[!grep("lag", variable), ]$akaike_weight)

#absolute (grepl("abs", variable))
gain_abs_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("abs", variable), ]$akaike_weight)

#raw (!grepl("change"))
gain_raw_importance <- sum(gain_model_comparison_annual_obs_taxonomy[!grep("change", variable), ]$akaike_weight)

#change (grepl("change", variable), (!grepl("abs")))
gain_change_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("change", variable), ][!grep("abs", variable),]$akaike_weight)

#type of temp variable (max, min, seas)
gain_max_temp_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("max", variable), ]$akaike_weight)
gain_min_temp_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("min", variable), ]$akaike_weight)
gain_seas_temp_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("seas", variable), ]$akaike_weight)

gain_model_comparison_annual_obs_taxonomy_akaikeweights <- gain_model_comparison_annual_obs_taxonomy

saveRDS(gain_model_comparison_annual_obs_taxonomy_akaikeweights, file = here::here("Model_results","gain_model_comparison_annual_obs_taxonomy_akaikeweights_annual_obs.rds"))
fwrite(gain_model_comparison_annual_obs_taxonomy_akaikeweights, file = here::here("tables","gain_model_comparison_annual_obs_taxonomy_akaikeweights_annual_obs.csv"))

gain_model_comparison_annual_obs_taxonomy_akaikeweights <- readRDS(here::here("Model_results","gain_model_comparison_annual_obs_taxonomy_akaikeweights_annual_obs.rds"))

```

Relative Variable Importance for different Pairs of Transformation and Metric
```{r RVI gain Pairs Transformation and Metric}

#I want to look at relative importance FOR 
#models with raw and min
gain_raw_min_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("min", variable), ][!grep("change", variable)]$akaike_weight)

#models with raw and max
gain_raw_max_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("max", variable), ][!grep("change", variable)]$akaike_weight)

#models with raw and seas
gain_raw_seas_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("seas", variable), ][!grep("change", variable)]$akaike_weight)


#models with change and min
gain_change_min_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("min_s.t_temp_change", variable), ][!grep("abs", variable)]$akaike_weight)

#models with change and max
gain_change_max_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("max_s.t_temp_change", variable), ][!grep("abs", variable)]$akaike_weight)

#models with change and seas
gain_change_seas_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("seas_s.t_temp_change", variable), ][!grep("abs", variable)]$akaike_weight)

#models with abs change and min
gain_abs_change_min_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("min_s.t_temp_change_abs", variable), ]$akaike_weight)

#models with abs change and max
gain_abs_change_max_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("max_s.t_temp_change_abs", variable), ]$akaike_weight)

#models with abs change and seas
gain_abs_change_seas_importance <- sum(gain_model_comparison_annual_obs_taxonomy[grep("seas_s.t_temp_change_abs", variable), ]$akaike_weight)

gain_transform_metric_comparison <- data.table(Transformation =
                                                c("raw", "raw", "raw", "change", "change", "change", "abs_change", "abs_change", "abs_change"), 
                                              Metric = 
                                                rep(c("min", "max", "seas"), 3), 
                                              RVI = 
                                                c(gain_raw_min_importance, gain_raw_max_importance, 
                                                  gain_raw_seas_importance, gain_change_min_importance, 
                                                  gain_change_max_importance, gain_change_seas_importance, 
                                                  gain_abs_change_min_importance, gain_abs_change_max_importance, 
                                                  gain_abs_change_seas_importance))


```


Relative Variable Importance Loss Models
```{r RVI loss}
#add ∆AIC to tables
min_loss_AICc <- min(loss_model_comparison_annual_obs_taxonomy[, AICc])
loss_model_comparison_annual_obs_taxonomy[,"deltaAICc" := (AICc - min_loss_AICc)]

#add relative likelihood exp( -0.5 * ∆AIC score for that model)
loss_model_comparison_annual_obs_taxonomy[,"rel_likelihood" := exp((-0.5 * deltaAICc))]
#sum relative likelihoods across all models
loss_model_comparison_annual_obs_taxonomy.likelihoodsum <- sum(loss_model_comparison_annual_obs_taxonomy[, rel_likelihood])
#Akaike weight for a model is this rel_likelihood devided by the sum of these values across all models
loss_model_comparison_annual_obs_taxonomy[,"akaike_weight" := rel_likelihood/loss_model_comparison_annual_obs_taxonomy.likelihoodsum]

#I want to look at relative importance FOR 
#bottom/surface (bottom = includes sbt, grepl("sbt", variable))
loss_bottom_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("sbt", variable), ]$akaike_weight)
loss_surface_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("sst", variable), ]$akaike_weight)

#lag/not lagged (grepl("lag", variable))
loss_lag_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag", variable), ]$akaike_weight)
loss_nolag_importance <- sum(loss_model_comparison_annual_obs_taxonomy[!grep("lag", variable), ]$akaike_weight)

#absolute (grepl("abs", variable))
loss_abs_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("abs", variable), ]$akaike_weight)

#raw (!grepl("change"))
loss_raw_importance <- sum(loss_model_comparison_annual_obs_taxonomy[!grep("change", variable), ]$akaike_weight)

#change (grepl("change", variable), (!grepl("abs")))
loss_change_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("change", variable), ][!grep("abs", variable),]$akaike_weight)

#type of temp variable (max, min, seas)
loss_max_temp_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("max", variable), ]$akaike_weight)
loss_min_temp_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("min", variable), ]$akaike_weight)
loss_seas_temp_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("seas", variable), ]$akaike_weight)

loss_model_comparison_annual_obs_taxonomy_akaikeweights <- loss_model_comparison_annual_obs_taxonomy

saveRDS(loss_model_comparison_annual_obs_taxonomy_akaikeweights, file = here::here("Model_results","loss_model_comparison_annual_obs_taxonomy_akaikeweights_annual_obs.rds"))
saveRDS(loss_model_comparison_annual_obs_taxonomy_akaikeweights, file = here::here("tables","loss_model_comparison_annual_obs_taxonomy_akaikeweights_annual_obs.rds"))
loss_model_comparison_annual_obs_taxonomy_akaikeweights <- readRDS(file = here::here("Model_results","loss_model_comparison_annual_obs_taxonomy_akaikeweights_annual_obs.rds"))

```

Relative Variable Importance for Pairs of Transformation and Metric Predicting losss
```{r RVI loss Pairs Transformation and Metric}

#I want to look at relative importance FOR 
#models with raw and min
loss_raw_min_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("min", variable), ][!grep("change", variable)]$akaike_weight)

#models with raw and max
loss_raw_max_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("max", variable), ][!grep("change", variable)]$akaike_weight)

#models with raw and seas
loss_raw_seas_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("seas", variable), ][!grep("change", variable)]$akaike_weight)


#models with change and min
loss_change_min_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("min_s.t_temp_change", variable), ][!grep("abs", variable)]$akaike_weight)

#models with change and max
loss_change_max_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("max_s.t_temp_change", variable), ][!grep("abs", variable)]$akaike_weight)

#models with change and seas
loss_change_seas_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("seas_s.t_temp_change", variable), ][!grep("abs", variable)]$akaike_weight)

#models with abs change and min
loss_abs_change_min_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("min_s.t_temp_change_abs", variable), ]$akaike_weight)

#models with abs change and max
loss_abs_change_max_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("max_s.t_temp_change_abs", variable), ]$akaike_weight)

#models with abs change and seas
loss_abs_change_seas_importance <- sum(loss_model_comparison_annual_obs_taxonomy[grep("seas_s.t_temp_change_abs", variable), ]$akaike_weight)

loss_transform_metric_comparison <- data.table(Transformation =
                                                c("raw", "raw", "raw", "change", "change", "change", "abs_change", "abs_change", "abs_change"), 
                                              Metric = 
                                                rep(c("min", "max", "seas"), 3), 
                                              RVI = 
                                                c(loss_raw_min_importance, loss_raw_max_importance, 
                                                  loss_raw_seas_importance, loss_change_min_importance, 
                                                  loss_change_max_importance, loss_change_seas_importance, 
                                                  loss_abs_change_min_importance, loss_abs_change_max_importance, 
                                                  loss_abs_change_seas_importance))

```


Let's see how the RVI importance varies by looking at specific lag values
```{r make a plot here for variability across lag values}

#first, table for lags
lags_RVI_scaled <- data.table(gain_lag = c(0:9), gain_RVI = 0, loss_RVI = 0)

#gains
lags_RVI_scaled[1,2] <- 1-sum(gain_model_comparison_annual_obs_taxonomy[grep("lag", variable), ]$akaike_weight)
lags_RVI_scaled[2,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag1", variable), ]$akaike_weight)
lags_RVI_scaled[3,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag2", variable), ]$akaike_weight)
lags_RVI_scaled[4,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag3", variable), ]$akaike_weight)
lags_RVI_scaled[5,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag4", variable), ]$akaike_weight)
lags_RVI_scaled[6,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag5", variable), ]$akaike_weight)
lags_RVI_scaled[7,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag6", variable), ]$akaike_weight)
lags_RVI_scaled[8,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag7", variable), ]$akaike_weight)
lags_RVI_scaled[9,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag8", variable), ]$akaike_weight)
lags_RVI_scaled[10,2] <- sum(gain_model_comparison_annual_obs_taxonomy[grep("lag9", variable), ]$akaike_weight)

#losss
lags_RVI_scaled[1,3] <- 1-sum(loss_model_comparison_annual_obs_taxonomy[grep("lag", variable), ]$akaike_weight)
lags_RVI_scaled[2,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag1", variable), ]$akaike_weight)
lags_RVI_scaled[3,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag2", variable), ]$akaike_weight)
lags_RVI_scaled[4,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag3", variable), ]$akaike_weight)
lags_RVI_scaled[5,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag4", variable), ]$akaike_weight)
lags_RVI_scaled[6,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag5", variable), ]$akaike_weight)
lags_RVI_scaled[7,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag6", variable), ]$akaike_weight)
lags_RVI_scaled[8,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag7", variable), ]$akaike_weight)
lags_RVI_scaled[9,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag8", variable), ]$akaike_weight)
lags_RVI_scaled[10,3] <- sum(loss_model_comparison_annual_obs_taxonomy[grep("lag9", variable), ]$akaike_weight)

saveRDS(lags_RVI_scaled, here::here("Model_results","lags_RVI_scaled_annual_obs.rds"))
lags_RVI_scaled <- readRDS(here::here("Model_results","lags_RVI_scaled_annual_obs.rds"))



#visualize lags through time
lags_over_time_annual_regs <- ggplot(data = lags_RVI_scaled, aes(x = gain_lag)) +
  geom_line(aes(y = gain_RVI), ) +
  geom_line(aes(y = loss_RVI), linetype = "dashed") +
  labs(x = "Lag (years)", y = "Relative Variable Importance") +
  scale_color_discrete() + #how to make a legend
  theme_classic() +
  theme(text = element_text(size = 13)) +
  scale_x_continuous(breaks = c(0:9))

save(lags_over_time_annual_regs, file = here::here("figures","Supplemental","annual_sensitivity","lags_over_time_annual_regs.RData"))

```

Table for RVI data for gain
```{r put RVIs for gain into data table}
gain_loss <- c("gain","gain","gain","gain","gain", "gain","gain", "gain", "gain","gain","gain","gain","gain","gain", "gain","gain", "gain", "gain","gain")
type <- c("Depth", "Depth", "lag", "lag", "Transformation", "Transformation", "Transformation", "Metric", "Metric", "Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric")
variable <- c("Bottom", "Surface", "no_lag", "lagged", "|Change|", "Change", "Raw", "Min", "Max", "Seas", "Raw x Min", "Raw x Max", "Raw x Seas", "Change x Min", "Change x Max", "Change x Seas", "|Change| x Min", "|Change| x Max", "|Change| x Seas")


value <- c(gain_bottom_importance, gain_surface_importance, gain_nolag_importance, gain_lag_importance, gain_abs_importance, gain_change_importance, gain_raw_importance, gain_min_temp_importance, gain_max_temp_importance, gain_seas_temp_importance, gain_raw_min_importance, gain_raw_max_importance, gain_raw_seas_importance, gain_change_min_importance, gain_change_max_importance, gain_change_seas_importance, gain_abs_change_min_importance, gain_abs_change_max_importance, gain_abs_change_seas_importance)

RVI_gain <- data.table(gain_loss, type, variable, value)
```

Table for RVI data for loss
```{r put RVIs for loss into data table}
gain_loss <- c("loss","loss","loss","loss","loss", "loss","loss", "loss", "loss","loss","loss","loss","loss","loss", "loss","loss", "loss", "loss","loss")
type <- c("Depth", "Depth", "lag", "lag", "Transformation", "Transformation", "Transformation", "Metric", "Metric", "Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric", "Transformation x Metric")
variable <- c("Bottom", "Surface", "no_lag", "lagged", "|Change|", "Change", "Raw", "Min", "Max", "Seas", "Raw x Min", "Raw x Max", "Raw x Seas", "Change x Min", "Change x Max", "Change x Seas", "|Change| x Min", "|Change| x Max", "|Change| x Seas")
value <- c(loss_bottom_importance, loss_surface_importance, loss_nolag_importance, loss_lag_importance, loss_abs_importance, loss_change_importance, loss_raw_importance, loss_min_temp_importance, loss_max_temp_importance, loss_seas_temp_importance, loss_raw_min_importance, loss_raw_max_importance, loss_raw_seas_importance, loss_change_min_importance, loss_change_max_importance, loss_change_seas_importance, loss_abs_change_min_importance, loss_abs_change_max_importance, loss_abs_change_seas_importance)

RVI_loss <- data.table(gain_loss, type, variable, value)
```

Merge RVI tables for gain and loss, and then graph
```{r merge RVI tables and plot}
RVI_table <- as.data.table(rbind(RVI_gain, RVI_loss)) #bind RVI for gains and losss

RVI_table.r <- RVI_table[!grep("Surface", variable)][!grep("lagged", variable)]

#change order of factor levels
RVI_table.r[, variable := factor(variable, levels = c("Bottom", "no_lag", "Raw", "Change", "|Change|", "Min", "Max", "Seas"))]

saveRDS(RVI_table, file = here::here("Model_results", "RVI_table_annual_obs.rds"))

ggplot(data = RVI_table.r, aes(x=variable, y = value, fill = gain_loss)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_discrete(breaks = c("Bottom", "no_lag", "Raw", "Change", "|Change|", "Min", "Max", "Seas"), 
                   labels = c("Bottom Temp", "No Lag", "Raw Temp", "Change in \nTemp", "| Change in\n Temp |", "Min Temp", "Max Temp", "Seas")) +
  labs(x = "Temperature Variables", y = "Relative Variable Importance")  +
  theme_bw()

```

Another way to visualize is stacked bar plot.

```{r RVI stacked plot}
#put levels into correct order for viewing, and add dummy level for legend
RVI_gain[, variable := factor(variable, levels = c("Bottom", "Surface","  ", "Raw", "Change", "|Change|","", "no_lag", "lagged", "Min", "Max", "Seas"))]
RVI_gain[, type := factor(type, levels = c("Depth", "Transformation", "Metric", "lag"))]

RVI_gain_nolag <- RVI_gain[type != "lag",]
RVI_gain_nolag[, type := factor(type, levels = c("Depth", "Transformation", "Metric"))][, variable := factor(variable, levels = c("Bottom", "Surface","  ", "Raw", "Change", "|Change|","", "Min", "Max", "Seas"))]

saveRDS(RVI_gain_nolag, here::here("Model_results", "RVI_gain_nolag_annual_obs.rds"))

#and again for loss
RVI_loss[, variable := factor(variable, levels = c("Bottom", "Surface","  ", "Raw", "Change", "|Change|","", "no_lag", "lagged", "Min", "Max", "Seas"))]
RVI_loss[, type := factor(type, levels = c("Depth", "Transformation", "Metric", "lag"))]

RVI_loss_nolag <- RVI_loss[type != "lag",]
RVI_loss_nolag[, type := factor(type, levels = c("Depth", "Transformation", "Metric"))][, variable := factor(variable, levels = c("Bottom", "Surface","  ", "Raw", "Change", "|Change|","", "Min", "Max", "Seas"))]

saveRDS(RVI_loss_nolag, here::here("Model_results", "RVI_loss_nolag_annual_obs.rds"))


#plotting
pal <- wes_palette("GrandBudapest1", 8, type = "continuous")
colors <- c(pal[1], pal[2], "#FFFFFF", pal[3], pal[4], pal[5],"#FFFFFF", pal[6], pal[7], pal[8])

#gains
RVI_gain_plot <- ggplot(data = RVI_gain_nolag, aes(x=type, y = value, fill = variable)) +
  geom_bar(stat="identity", color = "black", size = 0, width = 0.8) +
  scale_x_discrete(breaks = c("Depth", "Transformation", "Metric"), labels = c("Depth", "Transformation", "Metric")) +
  labs(x = "Temperature Variables", y = "Relative Variable Importance") +
  theme_classic() +
  theme(text = element_text(size = 11.5), 
        legend.position = "right", 
        legend.title = element_blank(), 
        legend.spacing.x = unit(0.2, 'cm'),
        legend.justification = "center",
        aspect.ratio = (1),
        #, legend.position = "none"
        ) +
  guides(fill=guide_legend(ncol=1)) +
  scale_fill_manual(values = colors,
                    drop = F)

#loss
RVI_loss_plot <- ggplot(data = RVI_loss_nolag, aes(x=type, y = value, fill = variable)) +
  geom_bar(stat="identity", color = "black", size = 0, width = 0.8) +
  scale_x_discrete(breaks = c("Depth", "Transformation", "Metric"), labels = c("Depth", "Transformation", "Metric")) +
  labs(x = "Temperature Variables", y = "Relative Variable Importance") +
  theme_classic() +
  theme(text = element_text(size = 11.5), 
        legend.position = "right", 
        legend.title = element_blank(), 
        legend.spacing.x = unit(0.2, 'cm'),
        legend.justification = "center",
        aspect.ratio = (1),
        #, legend.position = "none"
        ) +
  guides(fill=guide_legend(ncol=1)) +
  scale_fill_manual(values = colors,
                    drop = F)

RVI_table
RVI_gain_plot
RVI_loss_plot
```

Next step is to add the coefficients on to the bar chart. To do this, I will need to sum (aikake weight of model including variable x coefficient). I think this will be easy enough Using gain_model_comparison_annual_obs_taxonomy data frame. I will add new column to data frame of aikaike weight x coefficient

```{r avg coefficients for gain models}
#for some reason the coefficient is a factor
gain_model_comparison_annual_obs_taxonomy[,coef_num := as.numeric(as.character(coef))]
gain_model_comparison_annual_obs_taxonomy[,aw_coef := akaike_weight * coef_num]

#bottom temperature avg coefficient
gain_bottom_temp_avg_coef <- sum(gain_model_comparison_annual_obs_taxonomy[grep("sbt", variable)]$coef_num)

#surface temperature avg coefficient
gain_surface_temp_avg_coef <- sum(gain_model_comparison_annual_obs_taxonomy[grep("sst", variable)]$coef_num)

#raw seasonality avg coefficient
gain_seas_raw_avg_coef <- 
  sum(gain_model_comparison_annual_obs_taxonomy[grep("seas", variable), ][!grep("change", variable)]$coef_num)

#max abs change avg coefficient
gain_max_abs_change_avg_coef <- 
  sum(gain_model_comparison_annual_obs_taxonomy[grep("max_s.t_temp_change_abs", variable), ]$coef_num)


```

```{r avg coefficients for loss models}
#for some reason the coefficient is a factor
loss_model_comparison_annual_obs_taxonomy[,coef_num := as.numeric(as.character(coef))]
loss_model_comparison_annual_obs_taxonomy[,aw_coef := akaike_weight * coef_num]

#bottom temperature avg coefficient
loss_bottom_temp_avg_coef <- sum(loss_model_comparison_annual_obs_taxonomy[grep("sbt", variable)]$coef_num)

#surface temperature avg coefficient
loss_surface_temp_avg_coef <- sum(loss_model_comparison_annual_obs_taxonomy[grep("sst", variable)]$coef_num)

#raw seasonality avg coefficient
loss_seas_raw_avg_coef <- 
  sum(loss_model_comparison_annual_obs_taxonomy[grep("seas", variable), ][!grep("change", variable)]$coef_num)

#min change avg coefficient
loss_min_abs_change_avg_coef <- 
  sum(loss_model_comparison_annual_obs_taxonomy[grep("min_s.t_temp_change", variable), ][!grep("abs", variable)]$coef_num)

#seas change avg coefficient
loss_seas_change_avg_coef <- 
  sum(loss_model_comparison_annual_obs_taxonomy[grep("seas_s.t_temp_change", variable), ][!grep("abs", variable)]$coef_num)


```